### 2013-09-09 description/description.tex
Added description

\documentclass[11pt,a4paper,oneside]{article}
\usepackage[english]{babel}

\usepackage[utf8x]{inputenc} 
\usepackage[T1]{fontenc}

\usepackage{geometry} 
\usepackage{color} 
\usepackage{graphicx} 
\usepackage{pifont} 
\usepackage[babel]{csquotes}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{pgf}
\usepackage{nicefrac}
\usepackage{enumerate}

\usepackage{tgpagella}

\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareUnicodeCharacter{10230}{$\rightarrow$}

\newcommand{\wuppdi}[0]{\hfill\ensuremath{\square}}
\newcommand{\qed}[0]{\vspace{-3mm}\begin{flushright}\textit{q.e.d.}\end{flushright}\vspace{3mm}}
\newcommand{\bred}{\ensuremath{\longrightarrow_\beta}}
\newcommand{\acos}{\textrm{arccos}}
\newcommand{\determ}[1]{\textrm{det}(#1)}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\II}{\mathbb{I}}
\newcommand{\kernel}[1]{\textrm{ker}(#1)}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\phi}{\varphi}
\renewcommand{\theta}{\vartheta}
\newcommand{\atan}{\mathrm{arctan}}
\newcommand{\rot}{\mathrm{rot}}
\newcommand{\vdiv}{\mathrm{div}}
\newcommand{\shouldbe}{\stackrel{!}{=}}

\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL}

\geometry{a4paper,left=30mm,right=30mm, top=25mm, bottom=30mm}

\title{A Formalisation of Sturm Chains in \isabellehol}
\author{Manuel Eberl}

\begin{document}
\maketitle
\vskip-3mm

\section{Background}

\emph{\isabellehol} is an interactive theorem proving environment based on 
\emph{Higher Order Logic}. The \isabellehol library already contains a 
theory of univariate polynomials, including, in particular, differentiability. 
Furthermore, all basic analysis, such as limits and the intermediate and mean value 
theorem, are also available. Thus, most of the basic ingredients to 
formalise Sturm's theorem are available.

To explain the statement of Sturm's theorem in its most special case, 
we will first introduce the notion of a Sturm sequence of a polynomial.\\
Let $p\in\RR[X]$ that has no multiple real roots. The \emph{Sturm sequence}, 
or \emph{Sturm chain}, of $p$ is then defined as the sequence 
$p_0,p_1,\ldots p_n$ where:
\begin{align*}
&&\mathrm{where}\quad &p_i = \begin{cases}
\hskip3mm p & \mathrm{for}\ i = 0\\
\hskip3mm p' & \mathrm{for}\ i = 1\\
-p_{i-2}\ \mathrm{mod}\ p_{i-1} &\mathrm{otherwise}
\end{cases}&\\[2mm]
&&\mathrm{and}\quad &n=\mathrm{min} \{i\in\NN\ |\ \mathrm{degree}(p_i) \leq 0\}&
\end{align*}

Further, let $\sigma(x)$ denote the number of sign changes in the sequence 
$p_0(x), p_1(x), \ldots, p_n(x)$ for any $x \in \RR$. 
By \enquote{sign change}, we mean that a sign at one position is $-1$ and at the 
next position, skipping all zeroes, the sign is $1$, or vice versa.

Sturm's theorem now states that for $a,b\in\RR$ with $a\leq b$, the number of 
real roots of $p$ in the interval $(a;b]$ is $\sigma(a)-\sigma(b)$.\footnote{
Since $p$ was assumed to not have multiple roots, multiplicity is not relevant 
for counting here.}

Based on this, a number of generalisations can be used to count roots of 
polynomials with multiple roots as well, and furthermore, the total number of 
real roots can be determined by using $\lim_{x\to\pm\infty} \sigma(x)$, which 
can be determined by looking at the leading coefficient of the polynomials in 
the chain.

\section{Goal}

The goal is to formulise Sturm's theorem and its aforementioned generalisations 
in \emph{\isabellehol}. The result should be an executable function for 
determining the number of roots of a polynomial. Such a function can be 
exported to a programming language such as ML or Haskell using Isabelle's 
code generator, or it can be used to implement a decision procedure fo the 
number of roots of a polynomial directly in Isabelle.

One example for which such a decision procedure would be useful is the 
formalisation of algorithms for numerical approximation; for instance, there 
exists a formalisation of an algorithm for approximating the $\exp$ function 
that requires the fact that the polynomial 
$$\frac{1}{120}x^5+\frac{1}{24}x^4+\frac{1}{6}x^3-\frac{49}{16777216}x^2-\frac{17}{2097152}x$$
has exactly three roots in the interval $(-0.010831,0.010831)$. In Isabelle, this 
is written as:

\vspace*{-5mm}
\begin{align*}
&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"
\end{align*}

This fact is readily checkable with any computer algebra system, but was, so far, 
unfeasible to prove in Isabelle. We aim to be able to prove statements like these 
with our proof method derived from our formalisation of Sturm chains.

\section{Work outline}
We estimate that the work to be done is the following:
\begin{itemize}
\item Define, with enough generality to allow for later generalisations and optimisations, what 
      properties a general Sturm chain must fulfil
\item Prove that a chain fulfilling these requirements can indeed by used to count roots
\item Prove that the various canonical and non-canonical sturm chains fulfil these requirements
\item Prove that one can determine the signs of polynomials for sufficiently large arguments by 
      looking at the leading coefficients (this may require extension of the Polynomial library,
      since no lemmas about limits of polynomials exist so far)
\item Use these definitions and proofs to implement a decision procedure as described above and 
      integrate it with Isabelle
\end{itemize}

\end{document}

### 2013-09-09 description/description.tex
Updated description

@@
 real roots can be determined by using $\lim_{x\to\pm\infty} \sigma(x)$, which 
-can be determined by looking at the leading coefficient of the polynomials in 
+can be determined by looking at the leading coefficients of the polynomials in 
 the chain.
@@
       properties a general Sturm chain must fulfil
-\item Prove that a chain fulfilling these requirements can indeed by used to count roots
+\item Prove that a chain fulfilling these requirements can indeed be used to count roots
 \item Prove that the various canonical and non-canonical sturm chains fulfil these requirements

### 2013-09-11 description/description.tex
Added user guide, updated description

@@
 \usepackage{enumerate}
+\usepackage{xspace}
 
@@
 
-\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL}
+\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL\xspace}
 
@@
 
-\title{A Formalisation of Sturm Chains in \isabellehol}
+\title{A Formalisation of Sturm's theorem in \isabellehol}
 \author{Manuel Eberl}
@@
 
-The goal is to formulise Sturm's theorem and its aforementioned generalisations 
+The goal is to formalise Sturm's theorem and its aforementioned generalisations 
 in \emph{\isabellehol}. The result should be an executable function for 
@@
 code generator, or it can be used to implement a decision procedure fo the 
-number of roots of a polynomial directly in Isabelle.
+number of roots of a polynomial directly in Isabelle.\footnote{This can only work for rational coefficients and interval bounds, of course.}
 
@@
 \begin{align*}
-&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
-&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"
+&\textbf{lemma}\ "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
+&\hskip15mm 1/120*x^4 + 1/24*x^4 + 1/6*x^3 - 49/16777216*x^2 - 17/2097152*x\ =\ 0\} = 3"
 \end{align*}
@@
 \begin{itemize}
+\item Study the mathematical literature on Sturm chains
 \item Define, with enough generality to allow for later generalisations and optimisations, what 

### 2013-09-11 guide/guide.tex
Added user guide, updated description

\documentclass[11pt,a4paper,oneside]{article}
\usepackage[english]{babel}

\usepackage[utf8x]{inputenc} 
\usepackage[T1]{fontenc}

\usepackage{geometry} 
\usepackage{color} 
\usepackage{graphicx} 
\usepackage{pifont} 
\usepackage[babel]{csquotes}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{pgf}
\usepackage{nicefrac}
\usepackage{enumerate}
\usepackage{stmaryrd}

\usepackage{tgpagella}

\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareUnicodeCharacter{10230}{$\rightarrow$}

\newcommand{\ie}{i.\,e.}
\newcommand{\wuppdi}[0]{\hfill\ensuremath{\square}}
\newcommand{\qed}[0]{\vspace{-3mm}\begin{flushright}\textit{q.e.d.}\end{flushright}\vspace{3mm}}
\newcommand{\bred}{\ensuremath{\longrightarrow_\beta}}
\newcommand{\acos}{\textrm{arccos}}
\newcommand{\determ}[1]{\textrm{det}(#1)}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\II}{\mathbb{I}}
\newcommand{\kernel}[1]{\textrm{ker}(#1)}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\phi}{\varphi}
\renewcommand{\theta}{\vartheta}
\newcommand{\atan}{\mathrm{arctan}}
\newcommand{\rot}{\mathrm{rot}}
\newcommand{\vdiv}{\mathrm{div}}
\newcommand{\shouldbe}{\stackrel{!}{=}}
\newcommand{\sturm}{\texttt{sturm}}
\newcommand{\lemma}{\textbf{lemma}}
\newcommand{\card}{\textrm{card}}
\newcommand{\real}{\textrm{real}}

\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL}

\geometry{a4paper,left=30mm,right=30mm, top=25mm, bottom=30mm}

\title{\LARGE User's Guide for the \texttt{sturm} Method\\[4mm]}
\author{\Large Manuel Eberl <eberlm@in.tum.de>\\[1mm]\large Institut für Informatik, Technische Universität München\\[4mm]}

\begin{document}
\begin{center}
\vspace*{20mm}
\includegraphics[width=4cm]{isabelle.pdf}
\end{center}
\vspace*{-5mm}
{\let\newpage\relax\maketitle}
\vspace*{10mm}
\tableofcontents
\newpage

\section{Introduction}

The \sturm\ method uses Sturm's theorem to determine the number of distinct real roots of a polynomial (with rational coefficients) within a certain interval. It also provides some preprocessing to decide a number of statements that can be reduced to real roots of polynomials, such as simple polynomial inequalities and logical combinations of polynomial equations.
\vspace*{10mm}

\section{Usage}

\subsection{Examples}
The following examples should give a good overview of what the \sturm\ method can do:
\begin{align*}
&\lemma\ "\card\ \{x::\real. (x - 1)^2 * (x + 1) = 0\} = 2"\ \textrm{\textbf{by}\ sturm}\\
&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "\card\ \{x::\real. x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\} = 1"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "\card\ \{x::\real. x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\} = 4"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "(x::\real)^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "(x::\real) > 0\ \Longrightarrow\ x^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "\llbracket (x::\real) > 0; x \leq 2/3\rrbracket\ \Longrightarrow\ x*x \neq\ x"\ \textrm{\textbf{by}\ sturm}\\
&\lemma\ "(x::\real) > 1\ \Longrightarrow\ x*x > x"\ \textrm{\textbf{by}\ sturm}\\
\end{align*}

\subsection{Determining the number of real roots}
The \enquote{classical} application of Sturm's theorem is to count the number of real roots of a polynomial in a certain interval. The \sturm\ method supports this for any polynomial with rational coefficients and any real interval, \ie $[a;b]$, $(a;b]$, $[a;b)$, and $(a;b)$ where $a\in\QQ\cup\{-\infty\}$ and $b\in\QQ\cup\{\infty\}$.\footnote{The restriction to rational numbers for the coefficients and interval bounds is to the fact that the code generator is used internally, which, of course, does not support computations on irrational real numbers.} The general form of the theorems the method expects is:
$$\card\ \{x::\real.\ a < x \wedge x < b \wedge p\ x = 0\}\ =\ ?n$$
$?n$ should be replaced by the actual number of such roots and $p$ may be any polynomial real function in $x$ with rational coefficients. The bounds $a < x$ and $x < b$ can be omitted for the \enquote{$\infty$} case.\\

Furthermore, the \sturm\ method can instantiate the number $?n$ on the right-hand side automatically if it is left unspecified (as a schematic variable in a schematic lemma). However, due to technical restrictions this also takes twice as long as simply proving that the specified number is correct.

\newpage
\subsection{Inequalities}

A simple special case of root counting is the statement that a polynomial $p\in\RR[X]$ has no roots in a certain interval, which can be written as:
$$\forall x::\real.\ x > a \wedge x < b \longrightarrow p\ x \neq 0$$
The \sturm\ method can be directly applied to statements such as this and prove them.

\subsection{More complex expressions}

By using some simple preprocessing, the \sturm\ method can also decide more complex statements:
$$\card\ \{x::\real.\ x > a\ \wedge\ x < b\ \wedge\ P\ x\}$$
where $P\ x$ is a \enquote{polynomial expression}, which is defined as:
\begin{enumerate}
\item $p\ x= q\ x$, where $p$ and $q$ are polynomial functions, such as $\lambda x.\ a$, $\lambda x.\ x$, $\lambda x.\ x^2$, $\mathrm{poly}\ p$, and so on
\item $P\ x\ \wedge\ Q\ x$ or $P\ x\ \vee\ Q\ x$, where $P\ x$ and $Q\ x$ are polynomial expressions
\end{enumerate}

Of course, by reduction to the case of zero roots, the following kind of statement is also provable by \sturm\ :
$$\forall x::\real.\ x > a\ \wedge\ x < b\ \longrightarrow\ P\ x$$
where $P\ x$ is a \enquote{negated polynomial expression}, which is defined as:
\begin{enumerate}
\item $p\ x\neq q\ x$, where $p$ and $q$ are polynomial functions
\item $P\ x\ \wedge\ Q\ x$ or $P\ x\ \vee\ Q\ x$, where $P\ x$ and $Q\ x$ are negated polynomial expressions
\end{enumerate}

\subsection{Simple ordered inequalities}
For any polynomial $p\in\RR[X]$, the question whether $p(x) > 0$ for all $x\in I$ for a non-empty real interval $I$ can obviously be reduced to the question of whether $p(x) \neq 0$ for all $x\in I$, \ie $p$ has no roots in $I$, and $p(x) > 0$ for some arbitrary fixed $x\in I$, the first of which can be decided using Sturm's theorem and the second by choosing an arbitrary $x\in I$ and evaluating $p(x)$.\\

Using this reduction, the \sturm\ method can also decide single \enquote{less than}/\enquote{ŋreater than} inequalities of the form 
$$\forall x::\real.\ x > a\ \wedge\ x < b\ \longrightarrow\ p\ x < q\ x$$

\subsection{A note on meta logic versus object logic}

While statements like $\forall x::\real.\ x^2+1>0$ were expressed in their HOL notation in this guide, the \sturm\ method can also prove the meta logic equivalents $\bigwedge x::\real.\ x^2+1>0$ and $(x::\real)^2+1>0$ directly.

\section{Troubleshooting}

Should you find that the \sturm\ method fails to prove a statement that it should, according to the above text, be able to prove, please go through the following steps:
\begin{enumerate}
\item ensure that your function is indeed a \emph{real} polynomial. Add a type annotation to \emph{real} if necessary.
\item use a computer algebra system to ensure that the property is indeed correct
\item if this did not help, send the statement in question to \texttt{eberlm@in.tum.de}; it may be a bug in the preprocessing of the proof method.
\end{enumerate}

\end{document}

### 2013-10-16 paper/paper.tex
Added paper

\documentclass[11pt,a4paper,oneside]{article}
\usepackage[english]{babel}

\usepackage[utf8x]{inputenc} 
\usepackage[T1]{fontenc}

\usepackage{geometry} 
\usepackage{color} 
\usepackage{graphicx} 
\usepackage{pifont} 
\usepackage[babel]{csquotes}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{textcomp}
\usepackage{latexsym}
\usepackage{pgf}
\usepackage{nicefrac}
\usepackage{enumerate}
\usepackage{xspace}

\usepackage{tgpagella}

\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareUnicodeCharacter{10230}{$\rightarrow$}

\newtheorem{definition}{Definition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{theorem}[definition]{Theorem}
\newcommand{\flush}{\leavevmode\newline\vspace*{-1.5em}}

%\newcommand{\qed}[0]{\vspace{-3mm}\begin{flushright}\textit{q.e.d.}\end{flushright}\vspace{3mm}}
\newcommand{\bred}{\ensuremath{\longrightarrow_\beta}}
\newcommand{\acos}{\textit{arccos}}
\newcommand{\determ}[1]{\textit{det}(#1)}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\II}{\mathbb{I}}
\newcommand{\kernel}[1]{\textit{ker}(#1)}
\renewcommand{\epsilon}{\varepsilon}
\renewcommand{\phi}{\varphi}
\renewcommand{\theta}{\vartheta}
\newcommand{\atan}{\mathit{arctan}}
\newcommand{\rot}{\mathit{rot}}
\newcommand{\vdiv}{\mathit{div}}
\newcommand{\shouldbe}{\stackrel{!}{=}}
\newcommand{\sgn}{\mathit{sgn}}
%\renewcommand{\gcd}{\mathit{gcd}}
%\renewcommand{\mod}{\ \mathit{mod}\ }

\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.\xspace}

\newcommand{\isabellehol}{\mbox{Isabelle}\slash HOL\xspace}

\geometry{a4paper,left=30mm,right=30mm, top=25mm, bottom=30mm}

\title{A Formalisation of Sturm's theorem in \isabellehol}
\author{Manuel Eberl}

\begin{document}
\maketitle
\vskip-3mm
\parindent0mm

\section{Outline}

This paper will give a detailed proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.

\section{Notation}
In this paper, $\RR[X]$ will be used to denote the set of all polynomials with real coefficients. For a polynomial $P\in\RR[X]$, we denote its derivative by $P'$. We say that some property holds in a neighbourhood of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.

\section{Proof of Sturm's theorem}

\subsection{Definitions}

First, we define three notions that will be important in the proof of Sturm's theorem:

A \emph{quasi-Sturm sequence} is a list of polynomials $P_0,P_1,\ldots,P_{n-1} \in \RR[X]$ that fulfils the following properties:
\begin{itemize}
\item $n \geq 1$, i.e. the sequence is not empty
\item $P_n$ does not change its sign, i.e. $\sgn(P_n(x))=\sgn(P_n(y))$ for all $x,y\in\RR$
\item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds
\end{itemize}
It can easily be seen that any nonempty sequence obtained by dropping a number of elements from the beginning of a quasi-Sturm sequence is again a quasi-Sturm sequence.\\

A \emph{Sturm sequence} is a quasi-Sturm sequence that fulfils the following additional properties:
\begin{itemize}
\item $n \geq 2$, i.e. the sequence contains at least two polynomials
\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$\footnote{
In \isabellehol, this is expressed as \enquote{eventually, the property $\sgn(P_0P_1(x))\ =\ \mathsf{if}\ x>x_0\ \mathsf{then}\ 1\ \mathsf{else}\ -1$ holds at $x_0$}.}
\item $P_0$ and $P_1$ have no common roots.
\end{itemize}

In Isabelle, these three concepts are captured in \emph{locales} of the same name.
\vskip5mm

Next, we define the notion of \emph{sign changes}. Let $P_0,\ldots,P_{n-1}\in\RR[X]$ be a sequence of polynomials and $x\in\RR$. By evaluating the $P_i$ at $x$, we obtain a sequence of real numbers $y_0,\ldots,y_{n-1}$. We now traverse this sequence from left to right and count how often the sign changes, \ie how often we see a positive number and the last nonzero number in the sequence was negative (or vice versa). This is called the number of \emph{sign changes} of the sequence $P_0,\ldots,P_{n-1}$ at the position $x$ and is denoted as $\sigma(P_0,\ldots,P_{n-1}; x)$.\\
In Isabelle, this is realised by evaluating the $P_i$ at $x$, deleting all zeros from the list, applying the \emph{rem\_adj\_dups} operation on the list and taking the length of the remaining list minus one as the result. The \emph{rem\_adj\_dups} function simply deletes equal adjacent elements in the list, \eg it turns $[1,2,2,1,1,1]$ into $[1,2,1]$.

A related notion are the sign changes \enquote{at infinity}. For this, we compute the sequences
$$\left(\sgn\left(\lim\limits_{x\to -\infty} P_i(x)\right)\right)_{0\leq i < n}\hskip6mm\textit{and}\hskip10mm
\left(\sgn\left(\lim\limits_{x\to \infty} P_i(x)\right)\right)_{0\leq i < n}$$
and count the sign changes in these sequences in the same way as before. The number of sign changes is called $\sigma_{-\infty}(P_0,\ldots,P_{n-1})$ and $\sigma_{\infty}(P_0,\ldots,P_{n-1})$, respectively. Note that the signs of these limits for a $P_i$ can be determined by taking the sign of the leading coefficient of $P_i$.

\subsection{Important auxiliary lemmas}

Since Sturm's theorem is based on counting sign changes, a crucial part in proving it is to formalise the counting of sign changes in a way that is convenient for subsequent proofs. In this proof, this was done by proving that under certain conditions, the computation of the number of sign changes of a sequence can be \enquote{decomposed}.

\begin{lemma}\label{thm:limits}
Let $P\in\RR[X]\setminus\{0\}$. Then there exist $l,u\in\RR$ such that $l<u$, all roots of $P$ are in $(l;u)$ and
\begin{align*}
&\forall x\leq l.\ \sgn(P(x)) = \sgn\left(\lim\limits_{x\to -\infty} P(x)\right)\\
&\forall x\geq u.\ \sgn(P(x)) = \sgn\left(\lim\limits_{x\to \infty} P(x)\right)
\end{align*}
\end{lemma}
\begin{proof}
Obviously, the limit of $P(x)$ at $-\infty$ (resp. $\infty$) is either $\infty$ or $-\infty$. Therefore, there exists an $l$ (resp. a $u$) such that for all $x\leq l$ (resp. $x\geq u$), $P(x)$ increases above every bound in the case of the limit $\infty$ or falls below every bound in case of the limit $-\infty$. In particular, if we choose 0 as the bound, we see that $P(x)$ will be positive for all values beyond some point in the case of $\infty$ or negative in the case of $-\infty$, which implies $\sgn(P(x)) = \sgn(\lim_{x\to -\infty} P(x))$ (resp. $\sgn(P(x)) = \sgn(\lim_{x\to \infty} P(x))$) for all $x$ beyond that point. Therefore, bounds $l$ and $u$ that fulfil the last two requirements exist.\\
Also, since the $P(x)$ is positive or negative for all $x$ below $l$ (resp. above $u$), it has no roots outside the interval $(l;u)$. Lastly, should $l<u$ not hold already, we can simply increase $u$ until it does.
\end{proof}

\begin{lemma}\label{thm:sign_changes_distrib}
Let $P_0, \ldots, P_{n-1}$ be a sequence of polynomials and $i\in\{0,\ldots,n-1\}$. Furthermore, assume $P_i(x) \neq 0$. Then $\sigma(P_0,\ldots,P_{n-1}; x) = \sigma(P_0,\ldots,P_i; x)+\sigma(P_i,\ldots,P_{n-1}; x)$
\end{lemma}
\begin{proof}
According to the informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{rem\_adj\_dups} and \emph{filter} functions.
\end{proof}

\begin{lemma}\label{thm:sign_changes_triple}
Assume $P(x)\neq 0$ and $\sgn(R(x)) = -\sgn(P(x))$. Then $\sigma(P,R,Q; x) = 1$
\end{lemma}
\begin{proof}
By case distinction
\end{proof}

\begin{lemma}\label{thm:same_signs}
Let $P_0,\ldots,P_{n-1}\in\RR[X]$, $Q_0\ldots,Q_{n-1}\in\RR[X]$ be two sequences of polynomials. Assume $\mathit{sgn}(P_i(x_1))=\mathit{sgn}(Q_i(x_2))$ for all $i$. Then $\sigma(P_0,\ldots,P_{n-1};x_1)=\sigma(Q_0,\ldots,Q_{n-1};x_2)$.
\end{lemma}
\begin{proof}
Trivial.
\end{proof}

%\begin{lemma}
%Let $P_0,\ldots,P_{n-1}\in\RR[X]$ be a quasi-Sturm sequence of length $\geq 3$. Then no two adjacent polynomials $P_i, P_{i+1}$ have a common root.
%\end{lemma}
%\begin{proof}
%Suppose $P_i$ and $P_{i+1}$ had a common root $x$. Since $n \geq 3$, at least one of $P_{i-1}$ or $P_{i+2}$ exists, and with the definition of a quasi-Sturm sequence and the fact that $P_{i-1}(x)=0$ and $P_i(x)=0$, we can conclude $0=P_{i-1}P_{i+1}<0$ resp. $0=P_iP_{i+2}<0$, which is obviously false.
%\end{proof}

\subsection{Sign changes of Sturm sequences}

\begin{lemma}\label{thm:nonzero}
Let $P_0,\ldots,P_{n-1}$ be a quasi-Sturm sequence. Assume $P_0(x_0)\neq 0$. Then $\sigma(P_0,\ldots,P_n; x_0)$ is constant in the neighbourhood of $x_0$.\\
\end{lemma}
\begin{proof}
By induction over the length of the sequence
\begin{itemize}
\item if the sequence has length 1, the number of sign changes is, of course, $0$ at any position.
\item if the sequence has length 2, we know that $P_1$ does not change its sign anywhere by definition of a quasi-Sturm chain. Furthermore, since $P_0(x_0)\neq 0$, the sign of $P_0$ does not change in the neighbourhood of $x_0$. By Lemma \ref{thm:same_signs}, the number of sign changes of the sequence $P_0,P_1$ also remains constant in the neighbourhood of $x_0$.
\item if the sequence has length $\geq 3$ and $P_1(x_0)\neq 0$, we have $\sigma(P_0,\ldots,P_{n-1}; x_0)=\sigma(P_0,P_1;x_0)+\sigma(P_1,\ldots,P_{n-1};x_0)$ from Lemma \ref{thm:sign_changes_distrib}. Since $P_0(x_0)\neq 0$ and $P_1(x_0)\neq 0$, the signs of $P_0$ and $P_1$ do not change in the neighbourhood of $x_0$ and therefore $\sigma(P_1,\ldots,P_{n-1};x_0)$ is constant in the neighbourhood of $x_0$. Also, the sequence $P_1,\ldots,P_{n-1}$ is again a quasi-Sturm sequence with $P_1(x_0)\neq 0$, \ie the induction hypothesis can be applied and we can conclude that $\sigma(P_1,\ldots,P_{n-1};x_0)$ is also constant in the neighbourhood of $x_0$.
\item if the sequence has length $\geq 3$ and $P_1(x_0)=0$, we have $P_0(x_0)P_2(x_0)<0$ by the definition of a quasi-Sturm chain and thus $P_0(x_0)\neq 0$, $P_2(x_0)\neq 0$, and $\mathit{sgn}(P_2(x_0))=-\mathit{sgn}(P_0(x_0))$. With Lemma \ref{thm:sign_changes_distrib}, we have $\sigma(P_0,\ldots,P_{n-1}; x_0) = \sigma(P_0,P_1,P_2;x_0) + \sigma(P_2,\ldots,P_{n-1}; x_0)$. Furthermore, since $P_0(x_0)\neq 0$ and $P_2(x_0)\neq 0$, $\sgn(P_2(x))=-\sgn(P_0(x))$ holds in the entire neighbourhood of $x_0$ as the signs of $P_0$ and $P_2$ do not change in the neighbourhood of $x_0$, and with Lemma \ref{thm:sign_changes_triple}, we then have $\sigma(P_0,P_1,P_2;x)=2$ in the entire neighbourhood of $x_0$. As for the second summand, we know that $P_2,\ldots,P_{n-1}$ is again a quasi-Sturm sequence and $P_2(x_0)\neq 0$, so we can apply the induction hypothesis and obtain that $\sigma(P_2,\ldots,P_{n-1};x_0)$, too, is constant in the neighbourhood of $x_0$.
\end{itemize}
\end{proof}

\label{thm:zero}
\begin{lemma}
Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence. Assume $P_0(x_0)=0$. Then for $x$ in the neighbourhood of $x_0$, we have $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)+1$ if $x<x_0$ and $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)$ if $x\geq x_0$.\\
\end{lemma}
\begin{proof}
From the definition of a Sturm chain, we know that $P_0(x_0)=0$ implies $P_1(x_0)\neq 0$ and that in the neighbourhood of $x_0$, we have $P_0(x)P_1(x)<0$ if $x<x_0$ and $P_0(x)P_1(x)>0$ if $x>x_0$. Therefore, in the neighbourhood of $x_0$, we have $P_1(x)\neq 0$ and, with Lemma \ref{thm:nonzero}, $\sigma(P_1,\ldots,P_{n-1};x) = \sigma(P_1,\ldots,P_{n-1};x_0)$.\\
Fix some $x$ in that neighbourhood of $x_0$. By Lemma \ref{thm:sign_changes_distrib}, we then have $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x)+\sigma(P_1,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x)+\sigma(P_1,\ldots,P_{n-1};x_0)$. What remains to be shown is that $\sigma(P_0,P_1;x)=1$ for $x<x_0$ and $\sigma(P_0,P_1;x)=0$ for $x>x_0$. This is a simple consequence of $P_0(x)P_1(x)<0$ for $x<0$ and $P_0(x)P_1(x)>0$ for $x>0$.
\end{proof}

\label{thm:count_roots_between}
\begin{lemma}
Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$. Fix $a,b\in\RR$ with $a \leq b$. Then $\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = |\{x\in (a;b]\ |\ P_0(x)=0\}|$\\
\end{lemma}
\begin{proof}
By induction over $|\{x\in (a;b].\ P_0(x)=0\}|$ for arbitrary $a, b$.
\begin{itemize}
\item if $|\{x\in (a;b].\ P_0(x)=0\}|=0$, we have $P_0(x)\neq 0$ for all $x\in (a;b]$. We now distinguish two cases:
\begin{itemize}
\item if $P_0(a)\neq 0$, we have $P_0(x)\neq 0$ for all $x\in[a;b]$. Lemma \ref{thm:nonzero} then implies that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of all $x\in[a;b]$ and must therefore be constant in the entire interval $[a;b]$.
\item if $P_0(a)=0$, we know from Lemma \ref{thm:zero} that $\sigma(P_0,\ldots,P_{n-1}; x)=\sigma(P_0,\ldots,P_{n-1}; a)$ for all $x\in(a;a+\epsilon)$ for some $\epsilon > 0$, \ie $\sigma(P_0,\ldots,P_{n-1})$ is constant in the right neighbourhood of $a$. Furthermore, using $P_0(x)\neq 0$ for all $x\in(a;b]$ and Lemma \ref{thm:nonzero}, that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of every $x\in(a;b]$. Therefore, $\sigma(P_0,\ldots,P_{n-1})$ is constant on the entire interval $[a;b]$.
\end{itemize}
Since $\sigma(P_0,\ldots,P_{n-1})$ is constant on the entire interval $[a;b]$, we have $\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = 0$, which is exactly the number of roots of $P_0$ in $(a;b]$.

\item if $|\{x\in (a;b].\ P_0(x)=0\}|=n>0$, we take the smallest root of $P_0$ in the interval $(a;b]$ and call it $x_2$. With Lemma \ref{thm:zero}, we know that $\sigma(P_0,\ldots,P_{n-1};x) = \sigma(P_0,\ldots,P_{n-1};x_2)+1$ for all $x$ in a sufficiently small left neighbourhood of $x_2$. Let $x_1$ then be some value in $(a;x_2)$ that is in this neighbourhood. We know that $P_0$ has no roots in the interval $(a;x_2]$ and thus, by induction hypothesis, $\sigma(P_0,\ldots,P_{n-1};a)=\sigma(P_0,\ldots,P_{n-1};x_1)=\sigma(P_0,\ldots,P_{n-1};x_2)+1$. Furthermore, the number of roots of $P_0$ in the interval $(x_2;b]$ is exactly $n-1$, so by induction hypothesis, $\sigma(P_0,\ldots,P_{n-1};x_2)-\sigma(P_0,\ldots,P_{n-1};b) = n-1$. Adding these two equations yields $\sigma(P_0,\ldots,P_{n-1};a)-\sigma(P_0,\ldots,P_{n-1};b) = n$.
\end{itemize}
\end{proof}

\begin{lemma}\label{thm:count_roots}
Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$. Then $\sigma_{-\infty}(P_0,\ldots,P_{n-1})-\sigma_\infty(P_0,\ldots,P_{n-1}) = |\{x\ |\ P_0(x)=0\}|$
\end{lemma}
\begin{proof}
Using Lemma \ref{thm:limits} for every $P_i$, we obtain $l,u\in\RR$ with $l<u$ such that $P_0$ has no roots outside the interval $(l;u)$ and for all $P_i$, $\sgn(P_i(l))=\sgn(\lim_{x\to -\infty} P_i(x))$ and $\sgn(P_i(u))=\sgn(\lim_{x\to \infty} P_i(x))$. It is then easy to see that the number of sign changes of the sequence $P_0,\ldots,P_{n-1}$ can be determined by considering the signs of the limits of the $P_i$ at $\pm\infty$ instead of $P_i(u)$ and $P_i(l)$. Since all roots of $P_0$ lie in the interval $(l;u)$, we have thus shown that the total number of roots of $P_0$ can be determined in the way described above.
\end{proof}

Note: similar statements for the usage of $\sigma_{-\infty}$ and $\sigma_\infty$ to compute the number of roots $>a$ or $\leq a$ for some fixed $a\in\RR$ can be proven analogously.

\section{Constructing Sturm chains}

\subsection{The canonical Sturm chain}

The canonical Sturm chain $P_0,\ldots,P_{n-1}$ of a polynomial $P\in\RR[X]$ is constructed as follows:
\begin{align*}
&&\mathit{where}\quad &P_i = \begin{cases}
\hskip3mm P & \mathit{for}\ i = 0\\
\hskip3mm P' & \mathit{for}\ i = 1\\
-P_{i-2}\ \mathit{mod}\ P_{i-1} &\mathit{otherwise}
\end{cases}&\\[2mm]
\end{align*}

$n$ is chosen such that the $n\geq 2$ and $P_{n-1}$ is constant. The details of this are not important, since choosing a higher value for $n$ only results in a number of zero polynomials at the end of the sequence, which do not change the result in any way. Note that this construction always terminates as the degree of the polynomials involved strictly decreases with every step, so that one reaches a constant polynomial in finitely many steps.\\

\begin{lemma}\label{thm:sturm_gcd}
For all $i\in\{0,n-1\}$, we have $\gcd(P_i, P_{i+1}) = \gcd(P,P')$\\
\end{lemma}
\begin{proof}
By induction on $i$ using the fact that $\gcd(R,S)=\gcd(S,R\mod S)=\gcd(S,-R\mod S)$.
\end{proof}

\begin{lemma}\label{thm:sturm_seq_canonical}
Let $P\in\RR[X]$ with no multiple roots. Then the canonical Sturm chain construction of $P$ yields an actual Sturm chain.
\end{lemma}
\begin{proof}
We simply prove the five conditions a Sturm chain has to satisfy:
\begin{itemize}
\item the sequence has at least length 2: obvious by definition
\item $P_{n-1}$ does not change its sign: obvious, since $P_{n-1}$ is constant
\item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds:\\
By construction, we have $P_{i+2} = -P_{i}\ \textit{mod}\ P_{i+1}$. With $P_{i+1}(x_0)=0$, this implies that $P_{i+2}(x) = -P_i(x)$, and because of Lemma \ref{thm:sturm_gcd} and the fact that $P$ has no multiple roots, we also have $P_i(x)\neq 0$ and thus $P_i(x)P_{i+2}(x)<0$.
\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
since $P_0=P$ and $P_1=P'$ are polynomials, there exists some neighbourhood of $x_0$ that does not contain any roots of either $P_0$ or $P_1$. For any $x<x_0$ in that neighbourhood, we can then apply the mean value theorem to obtain some $\xi\in[x;x_0]$ with $P(x)-P(x_0) = P'(\xi)(x-x_0)$, and since $P(x_0)=0$ and $x-x_0<0$, $P(x)=P'(\xi)$. Furthermore, $P(x)\neq 0$ and $P'(\xi)$ has the same sign as $P'(x)$, since we chose a neighbourhood without roots; therefore we have $P(x)P'(x)<0$. For any $x>x_0$, we can use the same argument to find that $P(x)P'(x)>0$.
\item $P_0$ and $P_1$ have no common roots:\\
$P_0=P$ and $P_1=P'$, so if $P_0$ and $P_1$ had a common root, it would be a multiple root of $P$, which contradicts our assumption.
\end{itemize}
\end{proof}
%First of all, it can easily be seen that some $x_0\in\RR$ is a multiple root of $P$ iff it is a root of both $P$ and $P'$ and thus a root of $\gcd(P,P')$:
%\begin{itemize}
%\item if $x_0$ is a root of $P$ with order $k\geq 2$, we can write $P$ as $(x-x_0)^kQ$ for some $Q\in\RR[X]$, and thus $P'=((x-x_0)^kQ)' = (x-x_0)^{k-1}Q + (x-x_0)^kQ' = (x-x_0)^{k-1}(Q + (x-x_0)Q')$, \ie $x_0$ is also a root of order $k-1\geq 1$ of $P'$. Consequently, $x-x_0$ divides both $P$ and $P'$ and thus also $\gcd(P,P')$, so that $x_0$ is a root of $\gcd(P,P')$.
%\item if $x_0$ is a root of $\gcd(P,P')$, $x-x_0$ divides $\gcd(P,P')$ and thus divides $P$ and $P'$. Therefore, $x_0$ is a root of both $P$ and $P'$. We can then write $P$ as $(x-x_0)Q$ for some $Q\in\RR[X]$ and then have $P'=((x-x_0)Q)'= Q + (x-x_0)Q'$, and since $x-x_0$ divides $P'=Q+(x-x_0)Q'$, it must also divide $Q$, which implies that $x_0$ is a root of $P$ of at least order 2.
%\end{itemize}

\subsection{The case of multiple roots}

Of course, one way to handle a polynomial with multiple roots is to \enquote{divide away} the \enquote{excess roots} by dividing the polynomial $P$ by $\gcd(P,P')$. However, surprisingly, it turns out that as long as the interval bounds $a$ and $b$ are not multiple roots of $P$ themselves, i.e. at least one of $P(a)\neq 0$ and $P'(a)\neq 0$ and at least one of $P(b)\neq 0$ and $P'(b)\neq 0$ holds, one can simply use the canonical Sturm chain without any modification. To prove this, we will first prove the following auxiliary result:

%Lemma: Let $P\in\RR[X]\setminus\{0\}$. Let $P_0,\ldots,P_{n-1}$ be the result of the canonical Sturm chain construction of $P$ and define $Q_i := P_i / \gcd(P,P')$. Then $Q_0,\ldots,Q_{n-1}$ is a Sturm chain and $Q_0$ has the same roots, disregarding multiplicity, as $P_0$.\\
%Proof:
%\begin{itemize}
%\item since $Q_0,\ldots,Q_{n-1}$ has the same length as $P_0,\ldots,P_{n-1}$, which is the result of the canonical Sturm chain construction, the length of $Q_0,\ldots,Q_{n-1}$ is $\geq 2$.

%\end{itemize}

\end{document}

### 2013-11-18 output/document/Sturm.tex
Lots of stuff

%
\begin{isabellebody}%
\def\isabellecontext{Sturm}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
\isacommand{theory}\isamarkupfalse%
\ Sturm\isanewline
\isakeyword{imports}\ {\isachardoublequoteopen}{\isachartilde}{\isachartilde}{\isacharslash}src{\isacharslash}HOL{\isacharslash}Library{\isacharslash}Poly{\isacharunderscore}Deriv{\isachardoublequoteclose}\ Sturm{\isacharunderscore}Library\isanewline
\isakeyword{begin}%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{\isa{sign{\isacharunderscore}changes} function%
}
\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ sign{\isacharunderscore}changes\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ {\isacharparenleft}x{\isacharcolon}{\isacharcolon}real{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ length\ {\isacharparenleft}group\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}{\isacharparenright}\ ps{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}ps\isactrlsub {\isadigit{1}}\ {\isacharat}\ {\isacharbrackleft}p{\isacharbrackright}\ {\isacharat}\ ps\isactrlsub {\isadigit{2}}{\isacharparenright}\ x\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}ps\isactrlsub {\isadigit{1}}\ {\isacharat}\ {\isacharbrackleft}p{\isacharbrackright}{\isacharparenright}\ x\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}{\isacharbrackleft}p{\isacharbrackright}\ {\isacharat}\ ps\isactrlsub {\isadigit{2}}{\isacharparenright}\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharcomma}\ subst\ group{\isacharunderscore}append{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ same{\isacharunderscore}signs{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}length\ ps\ {\isacharequal}\ length\ ps{\isacharprime}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}i\ {\isacharless}\ length\ ps{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps{\isacharbang}i{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps{\isacharprime}{\isacharbang}i{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \isacommand{from}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ y{\isacharparenright}{\isacharparenright}\ ps{\isacharprime}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ rule{\isacharcolon}\ list{\isacharunderscore}induct{\isadigit{2}}{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackright}{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ p\ ps\ p{\isacharprime}\ ps{\isacharprime}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}i{\isacharless}length\ ps{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps{\isacharprime}\ {\isacharbang}\ i{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ this{\isacharbrackright}\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ A{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ same{\isacharunderscore}signs{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharprime}{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}p\ {\isasymin}\ set\ ps{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ y{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ y{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ same{\isacharunderscore}signs{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}sturm{\isacharunderscore}triple{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ r\ x{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharbrackleft}p{\isacharcomma}q{\isacharcomma}r{\isacharbrackright}\ x\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}insert\ assms{\isacharcomma}\ auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}real{\isacharunderscore}def{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps\ {\isacharequal}\ \isanewline
\ \ \ \ length\ {\isacharparenleft}group\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ poly{\isacharunderscore}inf\ ps{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharequal}\ \isanewline
\ \ \ \ length\ {\isacharparenleft}group\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ poly{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}%
\isamarkupsection{Definition of Sturm sequences locale%
}
\isamarkuptrue%
\isacommand{locale}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharequal}\isanewline
\ \ \isakeyword{fixes}\ ps\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharparenleft}real\ poly{\isacharparenright}\ list{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ last{\isacharunderscore}ps{\isacharunderscore}sgn{\isacharunderscore}const{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ ps{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ ps{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ ps{\isacharunderscore}not{\isacharunderscore}Nil{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}ps\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ signs{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}i\ x{\isachardot}\ {\isasymlbrakk}i\ {\isacharless}\ length\ ps\ {\isacharminus}\ {\isadigit{2}}{\isacharsemicolon}\ poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x{\isacharparenright}\ {\isacharasterisk}\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x{\isacharparenright}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ sturm{\isacharunderscore}seq\ {\isacharequal}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharplus}\ \isanewline
\ \ \isakeyword{fixes}\ p\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ hd{\isacharunderscore}ps{\isacharunderscore}p{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}hd\ ps\ {\isacharequal}\ p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ length{\isacharunderscore}ps{\isacharunderscore}ge{\isacharunderscore}{\isadigit{2}}{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}length\ ps\ {\isasymge}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ deriv{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x\isactrlsub {\isadigit{0}}{\isachardot}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p\ {\isacharasterisk}\ ps{\isacharbang}{\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isakeyword{begin}\isanewline
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharcolon}\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ ps{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ ps{\isacharunderscore}first{\isacharunderscore}two{\isacharcolon}\isanewline
\ \ \ \ \isakeyword{obtains}\ q\ ps{\isacharprime}\ \isakeyword{where}\ {\isachardoublequoteopen}ps\ {\isacharequal}\ p\ {\isacharhash}\ q\ {\isacharhash}\ ps{\isacharprime}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}ps{\isacharunderscore}p\ length{\isacharunderscore}ps{\isacharunderscore}ge{\isacharunderscore}{\isadigit{2}}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ ps{\isacharcomma}\ simp{\isacharcomma}\ clarsimp{\isacharcomma}\ rename{\isacharunderscore}tac\ ps{\isacharprime}{\isacharcomma}\ case{\isacharunderscore}tac\ ps{\isacharprime}{\isacharcomma}\ auto{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ ps{\isacharunderscore}first{\isacharcolon}\ {\isachardoublequoteopen}ps\ {\isacharbang}\ {\isadigit{0}}\ {\isacharequal}\ p{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ ps{\isacharunderscore}first{\isacharunderscore}two{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\ \ \isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isasymin}\ set\ ps{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}in{\isacharunderscore}set{\isacharbrackleft}OF\ ps{\isacharunderscore}not{\isacharunderscore}Nil{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{end}\isamarkupfalse%
\isanewline
\isanewline
\isacommand{locale}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isacharequal}\ sturm{\isacharunderscore}seq\ {\isacharplus}\isanewline
\ \ \isakeyword{assumes}\ p{\isacharunderscore}squarefree{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ {\isasymnot}{\isacharparenleft}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymand}\ poly\ {\isacharparenleft}ps{\isacharbang}{\isadigit{1}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymnot}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}def{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}Cons{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharparenleft}p{\isacharhash}ps{\isacharparenright}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}ps\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ ps{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}unfold{\isacharunderscore}locales{\isacharparenright}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}ps\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ fact\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}p{\isacharhash}ps{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ y\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ last{\isacharunderscore}ps{\isacharunderscore}sgn{\isacharunderscore}const\ \isakeyword{and}\ {\isacharbackquoteopen}ps\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ ps{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ ps{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}p{\isacharhash}ps{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ i\ x\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ ps\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ signs{\isacharbrackleft}of\ {\isachardoublequoteopen}i{\isacharplus}{\isadigit{1}}{\isachardoublequoteclose}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Auxiliary lemmas about roots and sign changes%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ {\isacharminus}{\isacharparenright}\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}aux{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}ps\ {\isacharcolon}{\isacharcolon}\ real\ poly\ list{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymAnd}i\ x{\isachardot}\ {\isasymlbrakk}i\ {\isacharless}\ length\ ps\ {\isacharminus}\ {\isadigit{2}}{\isacharsemicolon}\ poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}j{\isasymle}i{\isacharplus}{\isadigit{1}}{\isachardot}\ poly\ {\isacharparenleft}ps\ {\isacharbang}\ j{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ i{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isadigit{0}}\ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}clarsimp{\isacharcomma}\ rename{\isacharunderscore}tac\ j{\isacharcomma}\ case{\isacharunderscore}tac\ j{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}Suc\ i{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ Suc{\isachardot}prems{\isacharparenleft}{\isadigit{1}}{\isacharcomma}{\isadigit{2}}{\isacharparenright}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ assms{\isacharparenleft}{\isadigit{4}}{\isacharparenright}{\isacharparenright}\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Suc{\isachardot}prems{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}ps\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Suc{\isachardot}prems\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}j{\isasymle}i{\isacharplus}{\isadigit{1}}{\isachardot}\ poly\ {\isacharparenleft}ps\ {\isacharbang}\ j{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ Suc{\isachardot}IH{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ Suc{\isachardot}prems{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}clarsimp{\isacharcomma}\ rename{\isacharunderscore}tac\ j{\isacharcomma}\ case{\isacharunderscore}tac\ {\isachardoublequoteopen}j\ {\isacharequal}\ Suc\ {\isacharparenleft}Suc\ i{\isacharparenright}{\isachardoublequoteclose}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
This function splits the sign list of a Sturm sequence at a 
  position \isa{x} that is not a root of \isa{p} into a 
  list of sublists such that the number of sign changes within 
  every sublist is constant in the neighbourhood of \isa{x},
  thus proving that the total number is also constant.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{fun}\isamarkupfalse%
\ split{\isacharunderscore}sign{\isacharunderscore}changes\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharbrackleft}p{\isacharbrackright}\ {\isacharparenleft}x\ {\isacharcolon}{\isacharcolon}\ real{\isacharparenright}\ {\isacharequal}\ {\isacharbrackleft}{\isacharbrackleft}p{\isacharbrackright}{\isacharbrackright}{\isachardoublequoteclose}\ {\isacharbar}\isanewline
{\isachardoublequoteopen}split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ x\ {\isacharequal}\ {\isacharbrackleft}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isacharbrackright}{\isachardoublequoteclose}\ {\isacharbar}\isanewline
{\isachardoublequoteopen}split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\ {\isacharequal}\isanewline
\ \ \ \ {\isacharparenleft}if\ poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ poly\ q\ x\ {\isacharequal}\ {\isadigit{0}}\ then\ \isanewline
\ \ \ \ \ \ \ {\isacharbrackleft}p{\isacharcomma}q{\isacharcomma}r{\isacharbrackright}\ {\isacharhash}\ split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x\isanewline
\ \ \ \ \ else\isanewline
\ \ \ \ \ \ \ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ {\isacharhash}\ split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}subset{\isacharbrackleft}dest{\isacharbrackright}{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}ps{\isacharprime}\ {\isasymin}\ set\ {\isacharparenleft}split{\isacharunderscore}sign{\isacharunderscore}changes\ ps\ x{\isacharparenright}\ {\isasymLongrightarrow}\ set\ ps{\isacharprime}\ {\isasymsubseteq}\ set\ ps{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}insert\ ps{\isacharunderscore}not{\isacharunderscore}Nil{\isacharparenright}\isanewline
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}induction\ ps\ x\ rule{\isacharcolon}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isachardot}induct{\isacharparenright}\isanewline
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharcomma}\ simp{\isacharcomma}\ rename{\isacharunderscore}tac\ p\ q\ r\ ps\ x{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ case{\isacharunderscore}tac\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ poly\ q\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}\ auto{\isacharparenright}\isanewline
\isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
A custom induction rule for \isa{split{\isacharunderscore}sign{\isacharunderscore}changes} that 
  uses the fact that all the intermediate parameters in calls 
  of \isa{split{\isacharunderscore}sign{\isacharunderscore}changes} are quasi-Sturm sequences.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}induct{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}{\isasymlbrakk}{\isasymAnd}p\ x{\isachardot}\ P\ {\isacharbrackleft}p{\isacharbrackright}\ x{\isacharsemicolon}\ {\isasymAnd}p\ q\ x{\isachardot}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ {\isasymLongrightarrow}\ P\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ x{\isacharsemicolon}\isanewline
\ \ \ \ {\isasymAnd}p\ q\ r\ ps\ x{\isachardot}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ {\isasymlbrakk}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ poly\ q\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ P\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x{\isacharsemicolon}\ \isanewline
\ \ \ \ \ \ \ \ poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ P\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isacharsemicolon}\isanewline
\ \ \ \ \ \ \ \ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ P\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isasymrbrakk}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ P\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isasymrbrakk}\ {\isasymLongrightarrow}\ P\ ps\ x{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ goal{\isadigit{1}}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ ps{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ ps\ x\ rule{\isacharcolon}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isachardot}induct{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{3}}\ p\ q\ r\ ps\ x{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}{\isacharbrackleft}OF\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{6}}{\isacharparenright}{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ q\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{6}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ dest{\isacharcolon}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}Cons{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{3}}\ A\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P\ {\isacharparenleft}r\ {\isacharhash}\ ps{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{6}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ dest{\isacharcolon}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}Cons{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{3}}\ A\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P\ {\isacharparenleft}q\ {\isacharhash}\ r\ {\isacharhash}\ ps{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{assume}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{6}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ dest{\isacharcolon}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}Cons{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{3}}\ A\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}P\ {\isacharparenleft}q\ {\isacharhash}\ r\ {\isacharhash}\ ps{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\ simp{\isacharunderscore}all\ \ \isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The total number of sign changes in the split list is the same
  as the number of sign changes in the original list.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}correct{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}hd\ ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharprime}\ {\isasymequiv}\ {\isasymlambda}ps\ x{\isachardot}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymSum}ps{\isacharprime}{\isasymleftarrow}split{\isacharunderscore}sign{\isacharunderscore}changes\ ps\ x{\isachardot}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharprime}\ ps\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ x\isactrlsub {\isadigit{0}}\ rule{\isacharcolon}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}induct{\isacharparenright}\isanewline
\isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{3}}\ p\ q\ r\ ps\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ IH\ {\isacharequal}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{2}}{\isacharcomma}{\isadigit{3}}{\isacharcomma}{\isadigit{4}}{\isacharparenright}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{3}}\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ signs{\isacharbrackleft}of\ {\isadigit{0}}{\isacharbrackright}\ \isakeyword{and}\ True\ \isacommand{have}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ sgn{\isacharunderscore}r{\isacharunderscore}x{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isacharasterisk}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{3}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharbrackleft}OF\ this{\isacharcomma}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isachardoublequoteclose}\ ps{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}{\isacharbrackleft}p{\isacharcomma}\ q{\isacharcomma}\ r{\isacharbrackright}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}r\ {\isacharhash}\ ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharprime}\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}{\isacharbackquoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ IH{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ True\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharbrackleft}OF\ this{\isacharcomma}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isacharhash}ps{\isachardoublequoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharprime}\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ IH{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ False\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}correct{\isacharunderscore}nbh{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}hd\ ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharprime}\ {\isasymequiv}\ {\isasymlambda}x\isactrlsub {\isadigit{0}}\ ps\ x{\isachardot}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymSum}ps{\isacharprime}{\isasymleftarrow}split{\isacharunderscore}sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isachardot}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ goal{\isadigit{1}}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}ps{\isacharunderscore}nz\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharbraceleft}p\ {\isasymin}\ set\ ps{\isachardot}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p{\isasymin}{\isacharquery}ps{\isacharunderscore}nz{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}Ball{\isacharunderscore}finite{\isacharcomma}\ auto\ intro{\isacharcolon}\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}same{\isacharunderscore}sign{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ {\isacharparenleft}{\isasymforall}p{\isasymin}{\isacharbraceleft}p\ {\isasymin}\ set\ ps{\isachardot}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbraceright}{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isasymlongrightarrow}\isanewline
\ \ \ \ \ \ \ \ sign{\isacharunderscore}changes{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}clarify{\isacharparenright}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ nbh{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}{\isacharquery}ps{\isacharunderscore}nz{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ x\isactrlsub {\isadigit{0}}\ rule{\isacharcolon}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}induct{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{3}}\ p\ q\ r\ ps\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ IH\ {\isacharequal}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{2}}{\isacharcomma}{\isadigit{3}}{\isacharcomma}{\isadigit{4}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{3}}\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ signs{\isacharbrackleft}of\ {\isadigit{0}}{\isacharbrackright}\ \isakeyword{and}\ True\ \isacommand{have}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn{\isacharunderscore}r{\isacharunderscore}x{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isacharasterisk}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{3}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ nbh\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ r\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharbrackleft}OF\ this{\isacharcomma}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isachardoublequoteclose}\ ps{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}{\isacharbrackleft}p{\isacharcomma}\ q{\isacharcomma}\ r{\isacharbrackright}{\isacharparenright}\ x\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}r\ {\isacharhash}\ ps{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ {\isacharparenleft}r{\isacharhash}ps{\isacharparenright}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ nbh\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}{\isacharbackquoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ IH{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ True\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ nbh\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharbrackleft}OF\ this{\isacharcomma}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}{\isachardoublequoteclose}\ {\isachardoublequoteopen}r{\isacharhash}ps{\isachardoublequoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isacharparenright}\ x\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ {\isacharparenleft}q{\isacharhash}r{\isacharhash}ps{\isacharparenright}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ nbh\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{5}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ IH{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ False\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sign{\isacharunderscore}changes{\isacharprime}{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharunderscore}aux{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}hd\ ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isasymin}\ set\ {\isacharparenleft}split{\isacharunderscore}sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ x\isactrlsub {\isadigit{0}}\ rule{\isacharcolon}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}induct{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ p\ x{\isacharparenright}\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{2}}\ p\ q\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isacharequal}\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{2}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{2}}{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}same{\isacharunderscore}sign{\isacharbrackleft}OF\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ last{\isacharunderscore}ps{\isacharunderscore}sgn{\isacharunderscore}const\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ sgn{\isacharunderscore}q{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ q\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ \ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p{\isasymin}set{\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ intro{\isacharcolon}\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ {\isacharunderscore}\ A{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ same{\isacharunderscore}signs{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharprime}{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{3}}\ p\ q\ r\ ps{\isacharprime}{\isacharprime}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ sturm\ {\isacharequal}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ IH\ {\isacharequal}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{2}}{\isacharcomma}{\isadigit{3}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ ps{\isacharprime}{\isacharprime}{\isacharunderscore}props\ {\isacharequal}\ goal{\isadigit{3}}{\isacharparenleft}{\isadigit{6}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ q{\isacharunderscore}{\isadigit{0}}\ {\isacharequal}\ this\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm\ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}p{\isacharhash}q{\isacharhash}r{\isacharhash}ps{\isacharprime}{\isacharprime}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ signs{\isacharbrackleft}of\ {\isadigit{0}}{\isacharbrackright}\ \isakeyword{and}\ q{\isacharunderscore}{\isadigit{0}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ signs{\isacharprime}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isacharasterisk}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ r{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isasymin}\ set\ {\isacharparenleft}split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharparenleft}r\ {\isacharhash}\ ps{\isacharprime}{\isacharprime}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ IH{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharcomma}\ fact{\isacharcomma}\ fact{\isacharcomma}\ simp\ add{\isacharcolon}\ r{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}{\isacharcomma}\ fact{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ ps{\isacharprime}{\isacharprime}{\isacharunderscore}props\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ q{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ ps{\isacharprime}{\isacharunderscore}props{\isacharcolon}\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isacharequal}\ {\isacharbrackleft}p{\isacharcomma}q{\isacharcomma}r{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ signs{\isacharbrackleft}of\ {\isadigit{0}}{\isacharbrackright}\ \isakeyword{and}\ q{\isacharunderscore}{\isadigit{0}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ sgn{\isacharunderscore}r{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\isactrlsub {\isadigit{0}}\ {\isacharasterisk}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ sgn{\isacharunderscore}r\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ r\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ r\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ eventually{\isacharunderscore}conj\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}same{\isacharunderscore}sign{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ simp{\isacharunderscore}all\ add{\isacharcolon}\ r{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ {\isacharunderscore}\ A{\isacharbrackright}{\isacharcomma}\ clarify{\isacharcomma}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ subst\ ps{\isacharprime}{\isacharunderscore}props{\isacharcomma}\ subst\ sign{\isacharunderscore}changes{\isacharunderscore}sturm{\isacharunderscore}triple{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isakeyword{and}\ B{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ r\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ r\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ prod{\isacharunderscore}neg{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}a\ {\isacharparenleft}b{\isacharcolon}{\isacharcolon}real{\isacharparenright}{\isachardot}\ {\isasymlbrakk}a{\isachargreater}{\isadigit{0}}{\isacharsemicolon}\ b{\isachargreater}{\isadigit{0}}{\isacharsemicolon}\ a{\isacharasterisk}b{\isacharless}{\isadigit{0}}{\isasymrbrakk}\ {\isasymLongrightarrow}\ False{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a\ {\isacharparenleft}b{\isacharcolon}{\isacharcolon}real{\isacharparenright}{\isachardot}\ {\isasymlbrakk}a{\isacharless}{\isadigit{0}}{\isacharsemicolon}\ b{\isacharless}{\isadigit{0}}{\isacharsemicolon}\ a{\isacharasterisk}b{\isacharless}{\isadigit{0}}{\isasymrbrakk}\ {\isasymLongrightarrow}\ False{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}drule\ mult{\isacharunderscore}pos{\isacharunderscore}pos{\isacharcomma}\ simp{\isacharcomma}\ simp{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ drule\ mult{\isacharunderscore}neg{\isacharunderscore}neg{\isacharcomma}\ simp{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ A\ \isakeyword{and}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ simp{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ sgn{\isacharunderscore}r\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ r{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ A\ B\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ r\ x\ {\isacharasterisk}\ poly\ p\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ r\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ sgn{\isacharunderscore}less\ sgn{\isacharunderscore}times{\isacharcomma}\ metis\ sgn{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}{\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ sgn{\isacharunderscore}r\ \isacommand{show}\isamarkupfalse%
\ sgn{\isacharunderscore}r{\isacharprime}{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ r\ x{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sgn{\isacharunderscore}real{\isacharunderscore}def\ not{\isacharunderscore}le\ not{\isacharunderscore}less\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharcomma}\ intro\ conjI\ impI{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ prod{\isacharunderscore}neg{\isacharbrackleft}of\ {\isachardoublequoteopen}poly\ r\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ p\ x{\isachardoublequoteclose}{\isacharbrackright}\ \isacommand{apply}\isamarkupfalse%
\ force{\isacharplus}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isadigit{1}}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ ps{\isacharprime}{\isacharunderscore}props{\isacharcomma}\ subst\ sign{\isacharunderscore}changes{\isacharunderscore}sturm{\isacharunderscore}triple{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ fact{\isacharcomma}\ metis\ A\ B\ sgn{\isacharunderscore}r{\isacharprime}{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ q{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ {\isacharequal}\ this\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isasymin}\ set\ {\isacharparenleft}split{\isacharunderscore}sign{\isacharunderscore}changes\ {\isacharparenleft}q\ {\isacharhash}\ r\ {\isacharhash}\ ps{\isacharprime}{\isacharprime}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ IH{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharcomma}\ fact{\isacharcomma}\ simp\ add{\isacharcolon}\ q{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}{\isacharcomma}\ fact{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ ps{\isacharprime}{\isacharprime}{\isacharunderscore}props\ \isakeyword{and}\ q{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isacharequal}\ {\isacharbrackleft}p{\isacharcomma}\ q{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}set\ ps{\isacharprime}{\isachardot}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ q{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ p{\isacharunderscore}not{\isacharunderscore}{\isadigit{0}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}p{\isasymin}set\ ps{\isacharprime}{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ same{\isacharunderscore}signs{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharprime}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}p{\isasymin}set\ ps{\isacharprime}{\isachardot}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ p\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ intro{\isacharcolon}\ eventually{\isacharunderscore}Ball{\isacharunderscore}finite\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ poly{\isacharunderscore}neighbourhood{\isacharunderscore}same{\isacharunderscore}sign{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharparenright}\ hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}hd\ ps{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}pss\ {\isacharequal}\ {\isachardoublequoteopen}split{\isacharunderscore}sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}f\ {\isacharequal}\ {\isachardoublequoteopen}{\isasymlambda}pss\ x{\isachardot}\ {\isasymSum}ps{\isacharprime}{\isasymleftarrow}pss{\isachardot}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ pss\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}ps{\isacharprime}{\isachardot}\ ps{\isacharprime}{\isasymin}set\ pss\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharquery}f\ pss\ x\ {\isacharequal}\ {\isacharquery}f\ pss\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ pss{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}Cons\ ps{\isacharprime}\ pss{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ {\isacharquery}f\ pss\ x\ {\isacharequal}\ {\isacharquery}f\ pss\ x\isactrlsub {\isadigit{0}}\ {\isasymand}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymlongrightarrow}\ {\isacharquery}f\ {\isacharparenleft}ps{\isacharprime}{\isacharhash}pss{\isacharparenright}\ x\ {\isacharequal}\ {\isacharquery}f\ {\isacharparenleft}ps{\isacharprime}{\isacharhash}pss{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ A\ {\isacharequal}\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ this\ eventually{\isacharunderscore}conj{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ A{\isacharcomma}\ simp{\isacharunderscore}all\ add{\isacharcolon}\ Cons{\isacharparenright}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ A\ {\isacharequal}\ this{\isacharbrackleft}of\ {\isacharquery}pss{\isacharbrackright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ B{\isacharcolon}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharquery}f\ {\isacharquery}pss\ x\ {\isacharequal}\ {\isacharquery}f\ {\isacharquery}pss\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ A{\isacharcomma}\ rule\ hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharunderscore}aux{\isacharbrackleft}OF\ assms{\isacharbrackright}{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ C\ {\isacharequal}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}correct{\isacharunderscore}nbh{\isacharbrackleft}OF\ assms{\isacharbrackright}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ D\ {\isacharequal}\ split{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}correct{\isacharbrackleft}OF\ assms{\isacharbrackright}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ E\ {\isacharequal}\ eventually{\isacharunderscore}conj{\isacharbrackleft}OF\ B\ C{\isacharbrackright}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ {\isacharunderscore}\ E{\isacharbrackright}{\isacharcomma}\ auto\ simp{\isacharcolon}\ D{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{hide{\isacharunderscore}fact}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isachardot}hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharunderscore}aux%
\begin{isamarkuptext}%
If x is not a root of p, the number of sign changes of the sequence 
  remains constant in the neighbourhood of x.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharparenright}\ p{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ \isanewline
\ \ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
If \isa{x} is a root of \isa{p} and \isa{p} is not the 
  zero polynomial, the number of sign changes decreases by 1 at 
  \isa{x}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharparenright}\ p{\isacharunderscore}zero{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}\ {\isacharplus}\ {\isacharparenleft}if\ x{\isacharless}x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isadigit{0}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ ps{\isacharunderscore}first{\isacharunderscore}two\ \isacommand{obtain}\isamarkupfalse%
\ q\ ps{\isacharprime}\ \isakeyword{where}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}ps\ {\isacharequal}\ p{\isacharhash}q{\isacharhash}ps{\isacharprime}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}ps{\isacharbang}{\isadigit{1}}\ {\isacharequal}\ q{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ eventually{\isacharunderscore}at{\isacharcomma}\ rule\ exI{\isacharbrackleft}of\ {\isacharunderscore}\ {\isadigit{1}}{\isacharbrackright}{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ p{\isacharunderscore}squarefree\ \isakeyword{and}\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ ps{\isachardoublequoteclose}\ \isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq{\isacharunderscore}Cons{\isacharbrackleft}of\ p\ {\isachardoublequoteopen}q{\isacharhash}ps{\isacharprime}{\isachardoublequoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ quasi{\isacharunderscore}sturm{\isacharunderscore}seq\ {\isachardoublequoteopen}q{\isacharhash}ps{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}poly\ q\ x\isactrlsub {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharbrackleft}\isakeyword{where}\ x\isactrlsub {\isadigit{0}}{\isacharequal}x\isactrlsub {\isadigit{0}}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\ \ \ \isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}without{\isacharunderscore}roots{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackright}\ deriv{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackright}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}\ {\isasymand}\ poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p{\isacharasterisk}ps{\isacharbang}{\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ only{\isacharcolon}\ {\isacharbackquoteopen}ps{\isacharbang}{\isadigit{1}}\ {\isacharequal}\ q{\isacharbackquoteclose}{\isacharcomma}\ intro\ eventually{\isacharunderscore}conj{\isacharparenright}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ {\isacharunderscore}\ A{\isacharbrackright}{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ x{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ zero{\isacharunderscore}less{\isacharunderscore}mult{\isacharunderscore}pos\ \isacommand{have}\isamarkupfalse%
\ zero{\isacharunderscore}less{\isacharunderscore}mult{\isacharunderscore}pos{\isacharprime}{\isacharcolon}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a\ b{\isachardot}\ {\isasymlbrakk}{\isacharparenleft}{\isadigit{0}}{\isacharcolon}{\isacharcolon}real{\isacharparenright}\ {\isacharless}\ a{\isacharasterisk}b{\isacharsemicolon}\ {\isadigit{0}}\ {\isacharless}\ b{\isasymrbrakk}\ {\isasymLongrightarrow}\ {\isadigit{0}}\ {\isacharless}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subgoal{\isacharunderscore}tac\ {\isachardoublequoteopen}a{\isacharasterisk}b\ {\isacharequal}\ b{\isacharasterisk}a{\isachardoublequoteclose}{\isacharcomma}\ auto{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ q{\isacharunderscore}sgn{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\ {\isacharless}\ x\isactrlsub {\isadigit{0}}\ then\ {\isacharminus}sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ else\ sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ sgn{\isacharunderscore}real{\isacharunderscore}def\ elim{\isacharcolon}\ linorder{\isacharunderscore}neqE{\isacharunderscore}linordered{\isacharunderscore}idom\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ dest{\isacharcolon}\ mult{\isacharunderscore}pos{\isacharunderscore}pos\ mult{\isacharunderscore}neg{\isacharunderscore}neg\ zero{\isacharunderscore}less{\isacharunderscore}mult{\isacharunderscore}pos\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ zero{\isacharunderscore}less{\isacharunderscore}mult{\isacharunderscore}pos{\isacharprime}\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}distrib{\isacharbrackleft}OF\ {\isacharbackquoteopen}poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ of\ {\isachardoublequoteopen}{\isacharbrackleft}p{\isacharbrackright}{\isachardoublequoteclose}\ ps{\isacharprime}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ x\ {\isacharplus}\ sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ q{\isacharunderscore}sgn\ \isakeyword{and}\ {\isacharbackquoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ x\ {\isacharequal}\ {\isacharparenleft}if\ x{\isacharless}x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sgn{\isacharunderscore}zero{\isacharunderscore}iff\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{4}}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ {\isacharparenleft}q{\isacharhash}ps{\isacharprime}{\isacharparenright}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\ \ \ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}aux{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}a\ {\isasymle}\ b{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}x{\isacharcolon}{\isacharcolon}real{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymlongrightarrow}\ eventually\ {\isacharparenleft}{\isasymlambda}{\isasymxi}{\isachardot}\ f\ {\isasymxi}\ {\isacharequal}\ {\isacharparenleft}f\ x{\isacharcolon}{\isacharcolon}nat{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymlongrightarrow}\ f\ x\ {\isacharequal}\ f\ b{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}clarify{\isacharparenright}\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isachargreater}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymle}\ b{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}x{\isacharprime}{\isachardot}\ x\ {\isasymle}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ b\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}{\isasymxi}{\isachardot}\ f\ {\isasymxi}\ {\isacharequal}\ f\ x{\isacharprime}{\isacharparenright}\ {\isacharparenleft}at\ x{\isacharprime}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ natfun{\isacharunderscore}eq{\isacharunderscore}in{\isacharunderscore}ivl{\isacharbrackleft}OF\ {\isacharbackquoteopen}x\ {\isasymle}\ b{\isacharbackquoteclose}\ this{\isacharbrackright}\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ x\ {\isacharequal}\ f\ b{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
If \isa{p} is non-constant, the number of roots in an interval can 
  be computed by the number of sign changes of the sequence at 
  the border of the interval.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharparenright}\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymle}\ b{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ b\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ int\ {\isacharparenleft}sign{\isacharunderscore}changes\ ps\ b{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}a\ {\isasymle}\ b{\isacharbackquoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ arbitrary{\isacharcolon}\ a\ b\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ rule{\isacharcolon}\ less{\isacharunderscore}induct{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}less\ a\ b{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}{\isasymexists}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ no{\isacharunderscore}roots{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}roots{\isacharequal}{\isacharunderscore}{\isachardoublequoteclose}{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ card{\isacharunderscore}roots{\isacharcolon}\ {\isachardoublequoteopen}card\ {\isacharquery}roots\ {\isacharequal}\ {\isacharparenleft}{\isadigit{0}}{\isacharcolon}{\isacharcolon}int{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ no{\isacharunderscore}roots{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}simp\ only{\isacharcolon}\ card{\isacharunderscore}roots\ eq{\isacharunderscore}iff{\isacharunderscore}diff{\isacharunderscore}eq{\isacharunderscore}{\isadigit{0}}{\isacharbrackleft}symmetric{\isacharbrackright}\ int{\isacharunderscore}int{\isacharunderscore}eq{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cases\ {\isachardoublequoteopen}poly\ p\ a\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ no{\isacharunderscore}roots\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ intro{\isacharcolon}\ natfun{\isacharunderscore}eq{\isacharunderscore}in{\isacharunderscore}ivl\ {\isacharbackquoteopen}a\ {\isasymle}\ b{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ p{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymlongrightarrow}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}aux{\isacharcomma}\ fact{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ p{\isacharunderscore}nonzero{\isacharunderscore}imp{\isacharunderscore}sign{\isacharunderscore}changes{\isacharunderscore}const{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}insert\ False{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ a{\isacharparenright}\ {\isacharparenleft}at\ a{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharparenright}\ \isacommand{defer}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}rule\ p{\isacharunderscore}zero{\isacharbrackleft}OF\ {\isacharbackquoteopen}poly\ p\ a\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}{\isacharcomma}\ force{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{done}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymdelta}\ \isakeyword{where}\ {\isasymdelta}{\isacharunderscore}props{\isacharcolon}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymdelta}\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isacharless}\ a{\isacharplus}{\isasymdelta}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ eventually{\isacharunderscore}at\ dist{\isacharunderscore}real{\isacharunderscore}def{\isacharparenright}\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}a\ {\isacharequal}\ b{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ x\ {\isasymequiv}\ {\isachardoublequoteopen}min\ {\isacharparenleft}a{\isacharplus}{\isasymdelta}{\isacharslash}{\isadigit{2}}{\isacharparenright}\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ False\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isacharless}\ x{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isacharless}\ a{\isacharplus}{\isasymdelta}{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymle}\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}{\isasymdelta}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}a\ {\isasymle}\ b{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isasymdelta}{\isacharunderscore}props\ {\isacharbackquoteopen}a\ {\isacharless}\ x{\isacharbackquoteclose}\ {\isacharbackquoteopen}x\ {\isacharless}\ a{\isacharplus}{\isasymdelta}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ A\ {\isacharbackquoteopen}a\ {\isacharless}\ x{\isacharbackquoteclose}\ {\isacharbackquoteopen}x\ {\isasymle}\ b{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ b{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}roots{\isacharunderscore}finite{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ fin{\isacharcolon}\ {\isachardoublequoteopen}finite\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ intro{\isacharcolon}\ finite{\isacharunderscore}subset{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ True\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isasymnoteq}\ {\isacharbraceleft}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ fin\ \isacommand{have}\isamarkupfalse%
\ card{\isacharunderscore}greater{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ fastforce\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ x\ {\isasymequiv}\ {\isachardoublequoteopen}Min\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ Min{\isacharunderscore}in{\isacharbrackleft}OF\ fin{\isacharbrackright}\ \isakeyword{and}\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ x{\isacharunderscore}props{\isacharcolon}\ {\isachardoublequoteopen}x\ {\isachargreater}\ a{\isachardoublequoteclose}\ {\isachardoublequoteopen}x\ {\isasymle}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ x{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ blast{\isacharplus}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ Min{\isacharunderscore}le{\isacharbrackleft}OF\ fin{\isacharbrackright}\ x{\isacharunderscore}props\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ x{\isacharunderscore}le{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isacharprime}{\isachardot}\ {\isasymlbrakk}x{\isacharprime}\ {\isachargreater}\ a{\isacharsemicolon}\ x{\isacharprime}\ {\isasymle}\ b{\isacharsemicolon}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\ {\isasymLongrightarrow}\ x\ {\isasymle}\ x{\isacharprime}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ x{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ left{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isacharprime}{\isachardot}\ a\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ x\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ x{\isacharunderscore}props\ x{\isacharunderscore}le\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ a\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ x\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ p{\isacharunderscore}zero{\isacharbrackleft}OF\ {\isacharbackquoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ unfolded\ eventually{\isacharunderscore}at\ dist{\isacharunderscore}real{\isacharunderscore}def{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ {\isasymepsilon}\ \isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isasymepsilon}{\isacharunderscore}props{\isacharcolon}\ {\isachardoublequoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}x{\isacharprime}{\isachardot}\ x{\isacharprime}\ {\isasymnoteq}\ x\ {\isasymand}\ {\isasymbar}x{\isacharprime}\ {\isacharminus}\ x{\isasymbar}\ {\isacharless}\ {\isasymepsilon}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ x{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharplus}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x{\isacharprime}\ {\isacharless}\ x\ then\ {\isadigit{1}}\ else\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ x{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}max\ {\isacharparenleft}x\ {\isacharminus}\ {\isasymepsilon}\ {\isacharslash}\ {\isadigit{2}}{\isacharparenright}\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymbar}x{\isacharprime}\ {\isacharminus}\ x{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}\ x{\isacharunderscore}props\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ x{\isacharprime}{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x{\isacharprime}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x{\isacharprime}\ {\isacharless}\ x\ then\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharplus}\ {\isadigit{1}}\ else\ sign{\isacharunderscore}changes\ ps\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymepsilon}{\isacharunderscore}props{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}x{\isacharprime}\ {\isacharequal}\ x{\isachardoublequoteclose}{\isacharcomma}\ simp{\isacharcomma}\ force{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x{\isacharprime}\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ x\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ x{\isacharprime}{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ x{\isacharunderscore}props\ {\isacharbackquoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymnotin}\ {\isacharbraceleft}x{\isacharprime}{\isacharprime}{\isachardot}\ a\ {\isacharless}\ x{\isacharprime}{\isacharprime}\ {\isasymand}\ x{\isacharprime}{\isacharprime}\ {\isasymle}\ x{\isacharprime}\ {\isasymand}\ poly\ p\ x{\isacharprime}{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ x{\isacharprime}{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ left\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isacharprime}{\isacharprime}{\isachardot}\ a\ {\isacharless}\ x{\isacharprime}{\isacharprime}\ {\isasymand}\ x{\isacharprime}{\isacharprime}\ {\isasymle}\ x{\isacharprime}\ {\isasymand}\ poly\ p\ x{\isacharprime}{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ less{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}of\ a\ x{\isacharprime}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ x{\isacharprime}{\isacharunderscore}def\ {\isacharbackquoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ simp{\isacharcolon}\ card{\isacharunderscore}greater{\isacharunderscore}{\isadigit{0}}{\isacharparenright}\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ signs{\isacharunderscore}left{\isacharcolon}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ int\ {\isacharparenleft}sign{\isacharunderscore}changes\ ps\ x{\isacharparenright}\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ a\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ x\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isasymunion}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ x\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ x{\isacharunderscore}props\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ left\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ x\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharplus}\ {\isadigit{1}}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ fin\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ x\ {\isacharless}\ x{\isacharprime}\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharless}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ less{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ this\ x{\isacharunderscore}props{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackright}\ \isakeyword{and}\ A\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ signs{\isacharunderscore}right{\isacharcolon}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ x\ {\isacharminus}\ int\ {\isacharparenleft}sign{\isacharunderscore}changes\ ps\ b{\isacharparenright}\ {\isacharplus}\ {\isadigit{1}}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isacharprime}{\isachardot}\ x{\isacharprime}\ {\isachargreater}\ a\ {\isasymand}\ x{\isacharprime}\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ signs{\isacharunderscore}left\ \isakeyword{and}\ signs{\isacharunderscore}right\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The number of sign changes in the limits of the polynomials to 
  positive (resp. negative) infinity can be used to compute the 
  number of roots above or below a certain number, or the total number.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharparenright}\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ set\ ps{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}in{\isacharunderscore}set{\isacharbrackleft}OF\ ps{\isacharunderscore}not{\isacharunderscore}Nil{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite\ {\isacharparenleft}set\ ps{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ polys{\isacharunderscore}inf{\isacharunderscore}sign{\isacharunderscore}thresholds{\isacharbrackleft}OF\ this{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ l\ u\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ lu{\isacharunderscore}props\ {\isacharequal}\ this\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}u\ {\isacharequal}\ {\isachardoublequoteopen}max\ a\ u{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymle}\ {\isacharquery}u{\isachardoublequoteclose}\isanewline
\ \ \ \isacommand{using}\isamarkupfalse%
\ lu{\isacharunderscore}props{\isacharparenleft}{\isadigit{3}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymin}\ set\ ps{\isacharbackquoteclose}{\isacharcomma}\ of\ x{\isacharbrackright}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}u\ {\isasymle}\ x{\isachardoublequoteclose}{\isacharcomma}\ auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}\ {\isacharequal}\ this\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ {\isacharquery}u{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ map\ poly{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ {\isacharquery}u{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ only{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sign{\isacharunderscore}changes{\isacharunderscore}inf{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}OF\ assms{\isacharbrackright}\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ \ card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ {\isacharquery}u\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ {\isacharquery}u\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ lu{\isacharunderscore}props\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharparenright}\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymin}\ set\ ps{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}in{\isacharunderscore}set{\isacharbrackleft}OF\ ps{\isacharunderscore}not{\isacharunderscore}Nil{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite\ {\isacharparenleft}set\ ps{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ polys{\isacharunderscore}inf{\isacharunderscore}sign{\isacharunderscore}thresholds{\isacharbrackleft}OF\ this{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ l\ u\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ lu{\isacharunderscore}props\ {\isacharequal}\ this\isanewline
\ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}l\ {\isacharequal}\ {\isachardoublequoteopen}min\ a\ l{\isachardoublequoteclose}\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isachargreater}\ {\isacharquery}l{\isachardoublequoteclose}\isanewline
\ \ \ \isacommand{using}\isamarkupfalse%
\ lu{\isacharunderscore}props{\isacharparenleft}{\isadigit{4}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymin}\ set\ ps{\isacharbackquoteclose}{\isacharcomma}\ of\ x{\isacharbrackright}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}l\ {\isacharless}\ x{\isachardoublequoteclose}{\isacharcomma}\ auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}\ {\isacharequal}\ this\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ {\isacharquery}l{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ map\ poly{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ {\isacharquery}l\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ only{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}OF\ assms{\isacharbrackright}\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ \ card\ {\isacharbraceleft}x{\isachardot}\ {\isacharquery}l\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ {\isacharquery}l\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ a\ {\isasymge}\ x\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ lu{\isacharunderscore}props\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharparenleft}\isakeyword{in}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharparenright}\ count{\isacharunderscore}roots{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ card\ {\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite\ {\isacharparenleft}set\ ps{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ polys{\isacharunderscore}inf{\isacharunderscore}sign{\isacharunderscore}thresholds{\isacharbrackleft}OF\ this{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ l\ u\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ lu{\isacharunderscore}props\ {\isacharequal}\ this\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ l{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ map\ poly{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}p{\isachardot}\ sgn\ {\isacharparenleft}poly\ p\ u{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ map\ poly{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps\ l\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ u{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ only{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ sign{\isacharunderscore}changes{\isacharunderscore}inf{\isacharunderscore}def\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}OF\ assms{\isacharbrackright}\ lu{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ \ card\ {\isacharbraceleft}x{\isachardot}\ l\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ u\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ l\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ u\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ lu{\isacharunderscore}props\ assms\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Canonical Sturm sequence%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ degree{\isacharunderscore}mod{\isacharunderscore}less{\isacharprime}{\isacharcolon}\ {\isachardoublequoteopen}degree\ q\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ degree\ {\isacharparenleft}p\ mod\ q{\isacharparenright}\ {\isacharless}\ degree\ q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms\ degree{\isacharunderscore}mod{\isacharunderscore}less\ \isacommand{by}\isamarkupfalse%
\ force%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{function}\isamarkupfalse%
\ sturm{\isacharunderscore}aux\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}sturm{\isacharunderscore}aux\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}\ q\ {\isacharequal}\ \isanewline
\ \ \ \ {\isacharparenleft}if\ degree\ q\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isacharbrackleft}p{\isacharcomma}q{\isacharbrackright}\ else\ p\ {\isacharhash}\ sturm{\isacharunderscore}aux\ q\ {\isacharparenleft}{\isacharminus}{\isacharparenleft}p\ mod\ q{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}pat{\isacharunderscore}completeness{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{termination}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}relation\ {\isachardoublequoteopen}measure\ {\isacharparenleft}degree\ {\isasymcirc}\ snd{\isacharparenright}{\isachardoublequoteclose}{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ simp{\isacharunderscore}all\ add{\isacharcolon}\ o{\isacharunderscore}def\ degree{\isacharunderscore}mod{\isacharunderscore}less{\isacharprime}{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{declare}\isamarkupfalse%
\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharbrackleft}simp\ del{\isacharbrackright}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ sturm\ \isakeyword{where}\ {\isachardoublequoteopen}sturm\ p\ {\isacharequal}\ sturm{\isacharunderscore}aux\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}{\isadigit{0}}{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm\ {\isadigit{0}}\ {\isacharequal}\ {\isacharbrackleft}{\isadigit{0}}{\isacharcomma}{\isadigit{0}}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}unfold\ sturm{\isacharunderscore}def{\isacharcomma}\ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}aux\ p\ q\ {\isacharequal}\ {\isacharbrackleft}{\isacharbrackright}\ {\isasymlongleftrightarrow}\ False{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}induction\ p\ q\ rule{\isacharcolon}\ sturm{\isacharunderscore}aux{\isachardot}induct{\isacharcomma}\ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ auto{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}neq{\isacharunderscore}Nil{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm\ p\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}hd\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharequal}\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isasymin}\ set\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ hd{\isacharunderscore}in{\isacharunderscore}set{\isacharbrackleft}OF\ sturm{\isacharunderscore}neq{\isacharunderscore}Nil{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isasymge}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ q\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ {\isacharparenleft}sturm{\isacharunderscore}aux\ p\ q{\isacharparenright}\ {\isasymge}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}induction\ p\ q\ rule{\isacharcolon}\ sturm{\isacharunderscore}aux{\isachardot}induct{\isacharcomma}\ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}degree\ {\isacharparenleft}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ q\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}degree\ {\isacharparenleft}last\ {\isacharparenleft}sturm{\isacharunderscore}aux\ p\ q{\isacharparenright}{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}induction\ p\ q\ rule{\isacharcolon}\ sturm{\isacharunderscore}aux{\isachardot}induct{\isacharcomma}\ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}aux\ p\ q\ {\isacharbang}\ {\isadigit{0}}\ {\isacharequal}\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}aux\ p\ q\ {\isacharbang}\ Suc\ {\isadigit{0}}\ {\isacharequal}\ q{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isadigit{0}}\ {\isacharequal}\ p{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isacommand{lemma}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ Suc\ {\isadigit{0}}\ {\isacharequal}\ pderiv\ p{\isachardoublequoteclose}\ \isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}indices{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sturm\ p{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}{\isacharparenleft}sturm\ p{\isacharbang}i\ mod\ sturm\ p{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ ps\ q\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlbrakk}ps\ {\isacharequal}\ sturm{\isacharunderscore}aux\ p\ q{\isacharsemicolon}\ i\ {\isacharless}\ length\ ps\ {\isacharminus}\ {\isadigit{2}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ ps{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}{\isacharparenleft}ps{\isacharbang}i\ mod\ ps{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ p\ q\ arbitrary{\isacharcolon}\ ps\ i\ rule{\isacharcolon}\ sturm{\isacharunderscore}aux{\isachardot}induct{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ p\ q{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}i\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ i{\isacharprime}\ \isakeyword{where}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isacharequal}\ Suc\ i{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ i{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ ps\ {\isasymge}\ {\isadigit{4}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ goal{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ deg{\isacharcolon}\ {\isachardoublequoteopen}degree\ q\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{obtain}\isamarkupfalse%
\ ps{\isacharprime}\ \isakeyword{where}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}ps\ {\isacharequal}\ p\ {\isacharhash}\ ps{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ deg\ \isacommand{have}\isamarkupfalse%
\ ps{\isacharprime}{\isacharcolon}\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isacharequal}\ sturm{\isacharunderscore}aux\ q\ {\isacharparenleft}{\isacharminus}{\isacharparenleft}p\ mod\ q{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}length\ ps\ {\isasymge}\ {\isadigit{4}}{\isacharbackquoteclose}\ \isakeyword{and}\ {\isacharbackquoteopen}ps\ {\isacharequal}\ p\ {\isacharhash}\ ps{\isacharprime}{\isacharbackquoteclose}goal{\isadigit{1}}{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}i\ {\isacharminus}\ {\isadigit{1}}\ {\isacharless}\ length\ ps{\isacharprime}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ deg\ ps{\isacharprime}\ this{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ ps\ {\isasymge}\ {\isadigit{3}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}degree\ q\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}aux\ p\ q\ {\isacharbang}\ Suc\ {\isacharparenleft}Suc\ {\isadigit{0}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}{\isacharparenleft}p\ mod\ q{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ True\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}ps{\isacharbang}i\ {\isacharequal}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}ps{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}\ {\isacharequal}\ q{\isachardoublequoteclose}\ {\isachardoublequoteopen}ps{\isacharbang}{\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}{\isacharparenleft}p\ mod\ q{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ this{\isacharbrackleft}OF\ sturm{\isacharunderscore}def\ assms{\isacharbrackright}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}aux{\isacharunderscore}gcd{\isacharcolon}\ {\isachardoublequoteopen}r\ {\isasymin}\ set\ {\isacharparenleft}sturm{\isacharunderscore}aux\ p\ q{\isacharparenright}\ {\isasymLongrightarrow}\ gcd\ p\ q\ dvd\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}induction\ p\ q\ rule{\isacharcolon}\ sturm{\isacharunderscore}aux{\isachardot}induct{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ p\ q{\isacharparenright}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}r\ {\isacharequal}\ p{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ r{\isacharcolon}\ {\isachardoublequoteopen}r\ {\isasymin}\ set\ {\isacharparenleft}sturm{\isacharunderscore}aux\ q\ {\isacharparenleft}{\isacharminus}{\isacharparenleft}p\ mod\ q{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharcomma}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}degree\ q\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ False\ r{\isacharbrackright}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ gcd{\isacharunderscore}poly{\isachardot}simps{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}q\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ \isakeyword{and}\ {\isacharbackquoteopen}r\ {\isasymnoteq}\ p{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}r\ {\isacharequal}\ q{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ {\isacharparenleft}asm{\isacharparenright}\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}gcd{\isacharcolon}\ {\isachardoublequoteopen}r\ {\isasymin}\ set\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isasymLongrightarrow}\ gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ dvd\ r{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ \ \ %
\endisadelimproof
%
\isatagproof
\isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ sturm{\isacharunderscore}aux{\isacharunderscore}gcd{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}propagate{\isacharunderscore}left{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}j{\isasymle}i{\isacharplus}{\isadigit{1}}{\isachardot}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ j{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\isanewline
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}intro\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}aux{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{1}}{\isacharcomma}{\isadigit{2}}{\isacharcomma}{\isadigit{3}}{\isacharparenright}{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ i\ x{\isacharparenright}\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}p\ {\isacharequal}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ i{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}q\ {\isacharequal}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{let}\isamarkupfalse%
\ {\isacharquery}r\ {\isacharequal}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{2}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}indices{\isacharbrackleft}OF\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}p\ {\isacharequal}\ {\isacharquery}p\ div\ {\isacharquery}q\ {\isacharasterisk}\ {\isacharquery}q\ {\isacharminus}\ {\isacharquery}r{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ mod{\isacharunderscore}div{\isacharunderscore}equality{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}p\ x\ {\isacharequal}\ poly\ {\isacharparenleft}{\isacharquery}p\ div\ {\isacharquery}q\ {\isacharasterisk}\ {\isacharquery}q\ {\isacharminus}\ {\isacharquery}r{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}p\ x\ {\isacharequal}\ {\isacharminus}poly\ {\isacharquery}r\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sgn{\isacharunderscore}minus{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}not{\isacharunderscore}squarefree{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymnot}rsquarefree\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}propagate{\isacharunderscore}left{\isacharbrackleft}OF\ assms{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}firsttwo{\isacharunderscore}signs{\isacharunderscore}aux{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}q\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ q{\isacharunderscore}pderiv{\isacharcolon}\ \isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ p{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ {\isacharparenleft}x\isactrlsub {\isadigit{0}}{\isacharcolon}{\isacharcolon}real{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p{\isacharasterisk}q{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \ {\isacharbackquoteopen}q\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}same{\isacharunderscore}sign\ q{\isacharunderscore}pderiv\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ poly{\isacharunderscore}neighbourhood{\isacharunderscore}without{\isacharunderscore}roots\ eventually{\isacharunderscore}conj{\isacharparenright}\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ {\isasymepsilon}\ \isakeyword{where}\ {\isasymepsilon}{\isacharunderscore}props{\isacharcolon}\ {\isachardoublequoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymforall}x{\isachardot}\ x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}\ {\isasymand}\ {\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}\ {\isasymlongrightarrow}\ \isanewline
\ \ \ \ \ \ poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ eventually{\isacharunderscore}at\ dist{\isacharunderscore}real{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ sqr{\isacharunderscore}pos{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isacharcolon}{\isacharcolon}real{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ sgn\ x\ {\isacharasterisk}\ sgn\ x\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}real{\isacharunderscore}def{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}simp\ only{\isacharcolon}\ eventually{\isacharunderscore}at\ dist{\isacharunderscore}real{\isacharunderscore}def{\isacharcomma}\ rule\ exI{\isacharbrackleft}of\ {\isacharunderscore}\ {\isasymepsilon}{\isacharbrackright}{\isacharcomma}\isanewline
\ \ \ \ \ \ \ \ \ intro\ conjI{\isacharcomma}\ fact\ {\isacharbackquoteopen}{\isasymepsilon}\ {\isachargreater}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isasymepsilon}{\isacharunderscore}props\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ q\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p{\isacharasterisk}q{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}x\ {\isasymge}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacharbackquoteopen}x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}MVT{\isacharbrackleft}OF\ this{\isacharcomma}\ of\ p{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ {\isasymxi}\ \isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isacharequal}\ this\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacharbackquoteopen}{\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isacharbackquoteclose}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}\ {\isasymepsilon}{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymbar}{\isasymxi}\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isachardoublequoteclose}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharasterisk}\ sgn\ {\isacharparenleft}poly\ q\ {\isasymxi}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ q{\isacharunderscore}pderiv\ sgn{\isacharunderscore}mult{\isacharparenright}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isasymepsilon}{\isacharunderscore}props\ {\isacharbackquoteopen}{\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}t{\isachardot}\ {\isasymxi}\ {\isasymle}\ t\ {\isasymand}\ t\ {\isasymle}\ x\ {\isasymlongrightarrow}\ poly\ q\ t\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ q\ {\isasymxi}{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isasymepsilon}{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ no{\isacharunderscore}roots{\isacharunderscore}inbetween{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ True\ {\isacharbackquoteopen}x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}\ {\isasymepsilon}{\isacharunderscore}props\ {\isasymxi}{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}mult\ sqr{\isacharunderscore}pos{\isacharparenright}\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}x\ {\isacharless}\ x\isactrlsub {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ sgn{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}{\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}MVT{\isacharbrackleft}OF\ {\isacharbackquoteopen}x\ {\isacharless}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ of\ p{\isacharbrackright}\ \isacommand{guess}\isamarkupfalse%
\ {\isasymxi}\ \isacommand{{\isachardot}{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isacharequal}\ this\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ {\isacharbackquoteopen}{\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isacharbackquoteclose}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}x\ {\isacharless}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}\ {\isasymepsilon}{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymbar}{\isasymxi}\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ p\ x\ {\isacharequal}\ {\isacharparenleft}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharasterisk}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ {\isasymxi}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}poly\ p\ {\isasymxi}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ field{\isacharunderscore}simps{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ p\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isacharparenright}\ {\isacharasterisk}\ sgn\ {\isacharparenleft}poly\ q\ {\isasymxi}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isasymepsilon}{\isacharunderscore}props\ {\isasymxi}{\isacharunderscore}props\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ q{\isacharunderscore}pderiv\ sgn{\isacharunderscore}mult{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isasymepsilon}{\isacharunderscore}props\ {\isacharbackquoteopen}{\isasymbar}x\ {\isacharminus}\ x\isactrlsub {\isadigit{0}}{\isasymbar}\ {\isacharless}\ {\isasymepsilon}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}t{\isachardot}\ x\ {\isasymle}\ t\ {\isasymand}\ t\ {\isasymle}\ {\isasymxi}\ {\isasymlongrightarrow}\ poly\ q\ t\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ q\ {\isasymxi}{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isasymxi}{\isacharunderscore}props\ {\isasymepsilon}{\isacharunderscore}props\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule{\isacharunderscore}tac\ sym{\isacharcomma}\ intro\ no{\isacharunderscore}roots{\isacharunderscore}inbetween{\isacharunderscore}imp{\isacharunderscore}same{\isacharunderscore}sign{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ False\ {\isacharbackquoteopen}x\ {\isasymnoteq}\ x\isactrlsub {\isadigit{0}}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}mult\ sqr{\isacharunderscore}pos{\isacharparenright}\ \isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}firsttwo{\isacharunderscore}signs{\isacharcolon}\isanewline
\ \ \isakeyword{fixes}\ ps\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly\ list{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ squarefree{\isacharcolon}\ {\isachardoublequoteopen}rsquarefree\ p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ p{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ {\isacharparenleft}x\isactrlsub {\isadigit{0}}{\isacharcolon}{\isacharcolon}real{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p\ {\isacharasterisk}\ sturm\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ squarefree\ p{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}pderiv\ p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ \ add{\isacharcolon}rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ sturm{\isacharunderscore}firsttwo{\isacharunderscore}signs{\isacharunderscore}aux{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ simp{\isacharunderscore}all\ add{\isacharcolon}\ rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}signs{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ squarefree{\isacharcolon}\ {\isachardoublequoteopen}rsquarefree\ p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{assumes}\ q{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}poly\ {\isacharquery}q\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}poly\ {\isacharquery}p\ x\ {\isacharasterisk}\ poly\ {\isacharquery}r\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}indices{\isacharbrackleft}OF\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i\ mod\ sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharquery}r\ {\isacharequal}\ {\isacharminus}\ {\isacharparenleft}{\isacharquery}p\ mod\ {\isacharquery}q{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharminus}{\isacharquery}r\ {\isacharequal}\ {\isacharquery}p\ mod\ {\isacharquery}q{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ mod{\isacharunderscore}div{\isacharunderscore}equality{\isacharbrackleft}of\ {\isacharquery}p\ {\isacharquery}q{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}p\ div\ {\isacharquery}q\ {\isacharasterisk}\ {\isacharquery}q\ {\isacharminus}\ {\isacharquery}r\ {\isacharequal}\ {\isacharquery}p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}{\isacharquery}p\ div\ {\isacharquery}q{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharquery}q\ x\ {\isacharminus}\ poly\ {\isacharquery}r\ x\ {\isacharequal}\ poly\ {\isacharquery}p\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ poly{\isacharunderscore}diff\ poly{\isacharunderscore}mult{\isacharparenright}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ q{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ r{\isacharunderscore}x{\isacharcolon}\ {\isachardoublequoteopen}poly\ {\isacharquery}r\ x\ {\isacharequal}\ {\isacharminus}poly\ {\isacharquery}p\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ sqr{\isacharunderscore}pos{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isacharcolon}{\isacharcolon}real{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ x\ {\isacharasterisk}\ x\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}case{\isacharunderscore}tac\ {\isachardoublequoteopen}x\ {\isasymge}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ mult{\isacharunderscore}pos{\isacharunderscore}pos\ mult{\isacharunderscore}neg{\isacharunderscore}neg{\isacharparenright}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}not{\isacharunderscore}squarefree{\isacharbrackleft}of\ i\ p{\isacharbrackright}\ assms\ r{\isacharunderscore}x\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}p\ x\ {\isacharasterisk}\ poly\ {\isacharquery}p\ x\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}force\ intro{\isacharcolon}\ sqr{\isacharunderscore}pos{\isacharparenright}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}r\ x\ {\isacharasterisk}\ poly\ {\isacharquery}p\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
If \isa{p} contains no multiple roots, \isa{sturm\ p}, i.e. 
  the canonical Sturm sequence for p, is a squarefree Sturm sequence 
  that can be used to determine the number of roots of \isa{p}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}sturm{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ \isanewline
\ \ \ \isakeyword{assumes}\ {\isachardoublequoteopen}rsquarefree\ p{\isachardoublequoteclose}\isanewline
\ \ \ \isakeyword{shows}\ {\isachardoublequoteopen}sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isacharparenleft}sturm\ p{\isacharparenright}\ p{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sturm\ p\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}hd\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharequal}\ p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isasymge}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ {\isasymnot}{\isacharparenleft}poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymand}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\ {\isacharcolon}{\isacharcolon}\ real\ \isakeyword{and}\ y\ {\isacharcolon}{\isacharcolon}\ real\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}degree\ {\isacharparenleft}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ {\isachardoublequoteopen}last\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharequal}\ {\isacharbrackleft}{\isacharcolon}c{\isacharcolon}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isachardoublequoteclose}{\isacharcomma}\ simp\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}firsttwo{\isacharunderscore}signs{\isacharbrackleft}OF\ assms{\isacharbrackright}\ \isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\isactrlsub {\isadigit{0}}{\isachardot}\ poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p{\isacharasterisk}sturm\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}signs{\isacharbrackleft}OF\ assms{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}i\ x{\isachardot}\ {\isasymlbrakk}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isacharsemicolon}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Canonical squarefree Sturm sequence%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
This removes multiple roots from \isa{p} by dividing it by 
  its gcd with its derivative. The resulting polynomials has the same 
  roots as \isa{p}, but with multiplicity 1.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree\ \isakeyword{where}\isanewline
\ \ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p\ {\isacharequal}\ sturm\ {\isacharparenleft}p\ div\ {\isacharparenleft}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharunderscore}not{\isacharunderscore}Nil{\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharunderscore}def{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{defines}\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p{\isacharprime}\ {\isasymequiv}\ p\ div\ gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}\ p{\isacharprime}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}rsquarefree\ p{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}subst\ rsquarefree{\isacharunderscore}roots{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}pderiv\ p{\isacharprime}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharcolon}{\isacharminus}x{\isacharcomma}{\isadigit{1}}{\isacharcolon}{\isacharbrackright}\ dvd\ gcd\ p{\isacharprime}\ {\isacharparenleft}pderiv\ p{\isacharprime}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ poly{\isacharunderscore}eq{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}iff{\isacharunderscore}dvd{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}gcd\ p{\isacharprime}\ {\isacharparenleft}pderiv\ p{\isacharprime}{\isacharparenright}\ {\isacharequal}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ False\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ poly{\isacharunderscore}eq{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}iff{\isacharunderscore}dvd{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}sturm{\isacharbrackleft}OF\ {\isacharbackquoteopen}rsquarefree\ p{\isacharprime}{\isacharbackquoteclose}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharcolon}\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p{\isachardoublequoteclose}\ p{\isacharprime}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharunderscore}def{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}hd\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}\ {\isacharequal}\ p{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}\ {\isasymge}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ {\isadigit{0}}\ {\isacharequal}\ p{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ Suc\ {\isadigit{0}}\ {\isacharequal}\ pderiv\ p{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharunderscore}def{\isacharparenright}\ \isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}rsquarefree\ p{\isacharprime}{\isacharbackquoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ {\isasymnot}\ {\isacharparenleft}poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymand}\ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ rsquarefree{\isacharunderscore}roots{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isachardot}signs\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}i\ x{\isachardot}\ {\isasymlbrakk}i\ {\isacharless}\ length\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isacharsemicolon}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isachardot}deriv\ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\isactrlsub {\isadigit{0}}{\isachardot}\ poly\ p{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p{\isacharprime}\ {\isacharasterisk}\ sturm{\isacharunderscore}squarefree\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\ {\isachargreater}\ x\isactrlsub {\isadigit{0}}\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Optimisation for non-multiple roots%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
We can also define the following non-canonical Sturm sequence that 
  is obtained by taking the canonical Sturm sequence of \isa{p}
  (possibly with multiple roots) and then dividing the entire 
  sequence by the gcd of p and its derivative.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharequal}\ {\isacharparenleft}let\ d\ {\isacharequal}\ gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ map\ {\isacharparenleft}{\isasymlambda}p{\isacharprime}{\isachardot}\ p{\isacharprime}\ div\ d{\isacharparenright}\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}propagate{\isacharunderscore}left{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isakeyword{and}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymforall}j{\isasymle}i{\isacharplus}{\isadigit{1}}{\isachardot}\ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ j{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}intro\ sturm{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}aux{\isacharbrackleft}OF\ assms{\isacharparenleft}{\isadigit{2}}{\isacharcomma}{\isadigit{3}}{\isacharcomma}{\isadigit{4}}{\isacharparenright}{\isacharbrackright}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ i\ x{\isacharparenright}\isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ q\ {\isasymequiv}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ i{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ r\ {\isasymequiv}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}Suc\ i{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ s\ {\isasymequiv}\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}Suc\ {\isacharparenleft}Suc\ i{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ q{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}q\ div\ d{\isachardoublequoteclose}\ \isakeyword{and}\ r{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}r\ div\ d{\isachardoublequoteclose}\ \isakeyword{and}\ s{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}s\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isacommand{have}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}d\ dvd\ q{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ dvd\ r{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ dvd\ s{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ q{\isacharunderscore}def\ r{\isacharunderscore}def\ s{\isacharunderscore}def\ d{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ intro{\isacharcolon}\ sturm{\isacharunderscore}gcd{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ qrs{\isacharunderscore}simps{\isacharcolon}\ {\isachardoublequoteopen}q\ {\isacharequal}\ q{\isacharprime}\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ {\isachardoublequoteopen}r\ {\isacharequal}\ r{\isacharprime}\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ {\isachardoublequoteopen}s\ {\isacharequal}\ s{\isacharprime}\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ q{\isacharprime}{\isacharunderscore}def\ r{\isacharprime}{\isacharunderscore}def\ s{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ dvd{\isacharunderscore}div{\isacharunderscore}mult{\isacharunderscore}self{\isacharparenright}\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ goal{\isadigit{1}}{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ i{\isacharunderscore}in{\isacharunderscore}range\ \isacommand{have}\isamarkupfalse%
\ r{\isacharprime}{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ r{\isacharprime}{\isacharunderscore}def\ r{\isacharunderscore}def\ d{\isacharunderscore}def\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ r{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ r\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ {\isacharbackquoteopen}r\ {\isacharequal}\ r{\isacharprime}\ {\isacharasterisk}\ d{\isacharbackquoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}indices{\isacharbrackleft}OF\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isacharequal}\ q\ div\ r\ {\isacharasterisk}\ r\ {\isacharminus}\ s{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ q{\isacharunderscore}def\ r{\isacharunderscore}def\ s{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ mod{\isacharunderscore}div{\isacharunderscore}equality{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}q{\isacharprime}\ {\isacharequal}\ {\isacharparenleft}q\ div\ r\ {\isacharasterisk}\ r\ {\isacharminus}\ s{\isacharparenright}\ div\ d{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ q{\isacharprime}{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ {\isacharparenleft}q\ div\ r\ {\isacharasterisk}\ r{\isacharparenright}\ div\ d\ {\isacharminus}\ s{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ s{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ div{\isacharunderscore}diff{\isacharbrackleft}symmetric{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ q\ div\ r\ {\isacharasterisk}\ r{\isacharprime}\ {\isacharminus}\ s{\isacharprime}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ dvd{\isacharunderscore}div{\isacharunderscore}mult{\isacharbrackleft}OF\ {\isacharbackquoteopen}d\ dvd\ r{\isacharbackquoteclose}{\isacharcomma}\ of\ {\isachardoublequoteopen}q\ div\ r{\isachardoublequoteclose}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ algebra{\isacharunderscore}simps\ r{\isacharprime}{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ div\ r\ {\isacharequal}\ q{\isacharprime}\ div\ r{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ qrs{\isacharunderscore}simps\ {\isacharbackquoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ q{\isacharprime}\ x\ {\isacharequal}\ poly\ {\isacharparenleft}q{\isacharprime}\ div\ r{\isacharprime}\ {\isacharasterisk}\ r{\isacharprime}\ {\isacharminus}\ s{\isacharprime}{\isacharparenright}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ r{\isacharprime}{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isachardot}{\isachardot}{\isachardot}\ {\isacharequal}\ {\isacharminus}poly\ s{\isacharprime}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ s{\isacharprime}\ x\ {\isacharequal}\ {\isacharminus}poly\ q{\isacharprime}\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{using}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ q{\isacharprime}{\isacharunderscore}def\ s{\isacharprime}{\isacharunderscore}def\ q{\isacharunderscore}def\ s{\isacharunderscore}def\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ d{\isacharunderscore}def\ sgn{\isacharunderscore}minus{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}adjacent{\isacharunderscore}roots{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ False\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}adjacent{\isacharunderscore}root{\isacharunderscore}propagate{\isacharunderscore}left{\isacharbrackleft}OF\ assms{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isadigit{0}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}pderiv\ p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ div{\isacharunderscore}gcd{\isacharunderscore}coprime{\isacharunderscore}poly\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}coprime\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ {\isacharparenleft}pderiv\ p\ div\ d{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ False\ \isacommand{using}\isamarkupfalse%
\ coprime{\isacharunderscore}imp{\isacharunderscore}no{\isacharunderscore}common{\isacharunderscore}roots\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}signs{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ {\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\ \isanewline
\ \ \isakeyword{assumes}\ q{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}poly\ {\isacharquery}q\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\ \isanewline
\ \ \ \ \ \ \ \ \ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}poly\ {\isacharquery}r\ x\ {\isacharasterisk}\ poly\ {\isacharquery}p\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ coprime{\isacharunderscore}imp{\isacharunderscore}no{\isacharunderscore}common{\isacharunderscore}roots\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ rsquarefree{\isacharcolon}\ {\isachardoublequoteopen}rsquarefree\ {\isacharparenleft}p\ div\ d{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ rsquarefree{\isacharunderscore}roots\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range\ \isacommand{have}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharprime}{\isacharcolon}\ {\isachardoublequoteopen}i\ {\isacharless}\ length\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharminus}\ {\isadigit{2}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ dvd\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i{\isacharparenright}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}d\ dvd\ {\isacharquery}p{\isacharprime}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}d\ dvd\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}Suc\ i{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}d\ dvd\ {\isacharquery}q{\isacharprime}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}d\ dvd\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ {\isacharparenleft}Suc\ {\isacharparenleft}Suc\ i{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}d\ dvd\ {\isacharquery}r{\isacharprime}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ intro{\isacharcolon}\ sturm{\isacharunderscore}gcd{\isacharparenright}\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ pqr{\isacharunderscore}simps{\isacharcolon}\ {\isachardoublequoteopen}{\isacharquery}p{\isacharprime}\ {\isacharequal}\ {\isacharquery}p\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}q{\isacharprime}\ {\isacharequal}\ {\isacharquery}q\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isacharquery}r{\isacharprime}\ {\isacharequal}\ {\isacharquery}r\ {\isacharasterisk}\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharprime}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ dvd{\isacharunderscore}div{\isacharunderscore}mult{\isacharunderscore}self{\isacharparenright}\ \isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ q{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ q{\isacharprime}{\isacharunderscore}{\isadigit{0}}{\isacharcolon}\ {\isachardoublequoteopen}poly\ {\isacharquery}q{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}indices{\isacharbrackleft}OF\ i{\isacharunderscore}in{\isacharunderscore}range{\isacharprime}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{2}}{\isacharparenright}\ {\isacharequal}\ {\isacharminus}\ {\isacharparenleft}sturm\ p\ {\isacharbang}\ i\ mod\ sturm\ p\ {\isacharbang}\ {\isacharparenleft}i{\isacharplus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharminus}{\isacharquery}r{\isacharprime}\ {\isacharequal}\ {\isacharquery}p{\isacharprime}\ mod\ {\isacharquery}q{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ mod{\isacharunderscore}div{\isacharunderscore}equality{\isacharbrackleft}of\ {\isacharquery}p{\isacharprime}\ {\isacharquery}q{\isacharprime}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}p{\isacharprime}\ div\ {\isacharquery}q{\isacharprime}\ {\isacharasterisk}\ {\isacharquery}q{\isacharprime}\ {\isacharminus}\ {\isacharquery}r{\isacharprime}\ {\isacharequal}\ {\isacharquery}p{\isacharprime}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}d{\isacharasterisk}{\isacharparenleft}{\isacharquery}p\ div\ {\isacharquery}q\ {\isacharasterisk}\ {\isacharquery}q\ {\isacharminus}\ {\isacharquery}r{\isacharparenright}\ {\isacharequal}\ d{\isacharasterisk}\ {\isacharquery}p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ pqr{\isacharunderscore}simps\ algebra{\isacharunderscore}simps{\isacharparenright}\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}p\ div\ {\isacharquery}q\ {\isacharasterisk}\ {\isacharquery}q\ {\isacharminus}\ {\isacharquery}r\ {\isacharequal}\ {\isacharquery}p{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}{\isacharquery}p\ div\ {\isacharquery}q{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharquery}q\ x\ {\isacharminus}\ poly\ {\isacharquery}r\ x\ {\isacharequal}\ poly\ {\isacharquery}p\ x{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}metis\ poly{\isacharunderscore}diff\ poly{\isacharunderscore}mult{\isacharparenright}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ q{\isacharunderscore}{\isadigit{0}}\ \isacommand{have}\isamarkupfalse%
\ r{\isacharunderscore}x{\isacharcolon}\ {\isachardoublequoteopen}poly\ {\isacharquery}r\ x\ {\isacharequal}\ {\isacharminus}poly\ {\isacharquery}p\ x{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}adjacent{\isacharunderscore}roots{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ i{\isacharunderscore}in{\isacharunderscore}range\ q{\isacharunderscore}{\isadigit{0}}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}p\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ sqr{\isacharunderscore}pos{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}x{\isacharcolon}{\isacharcolon}real{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ x\ {\isacharasterisk}\ x\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}case{\isacharunderscore}tac\ {\isachardoublequoteopen}x\ {\isasymge}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ mult{\isacharunderscore}pos{\isacharunderscore}pos\ mult{\isacharunderscore}neg{\isacharunderscore}neg{\isacharparenright}\isanewline
\ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ r{\isacharunderscore}x\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
This approach indeed also yields a valid squarefree Sturm sequence
  for the polynomial \isa{p\ {\isacharslash}\ q}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharprime}{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isacharparenleft}p\ {\isacharcolon}{\isacharcolon}\ real\ poly{\isacharparenright}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}d\ {\isasymequiv}\ gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isacharparenright}\ {\isacharparenleft}p\ div\ d{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isacharquery}ps{\isacharprime}\ {\isacharquery}p{\isacharprime}{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharquery}ps{\isacharprime}\ {\isasymnoteq}\ {\isacharbrackleft}{\isacharbrackright}{\isachardoublequoteclose}\ {\isachardoublequoteopen}hd\ {\isacharquery}ps{\isacharprime}\ {\isacharequal}\ {\isacharquery}p{\isacharprime}{\isachardoublequoteclose}\ {\isachardoublequoteopen}{\isadigit{2}}\ {\isasymle}\ length\ {\isacharquery}ps{\isacharprime}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ d{\isacharunderscore}def\ hd{\isacharunderscore}map{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ dvd\ last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}rule\ sturm{\isacharunderscore}gcd{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}last\ {\isacharparenleft}sturm\ p{\isacharparenright}\ {\isacharequal}\ last\ {\isacharquery}ps{\isacharprime}\ {\isacharasterisk}\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ last{\isacharunderscore}map\ d{\isacharunderscore}def\ dvd{\isacharunderscore}div{\isacharunderscore}mult{\isacharunderscore}self{\isacharparenright}\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ this\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}last\ {\isacharquery}ps{\isacharprime}\ dvd\ last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ dvd{\isacharunderscore}imp{\isacharunderscore}degree{\isacharunderscore}le{\isacharbrackleft}OF\ this{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}degree\ {\isacharparenleft}last\ {\isacharquery}ps{\isacharprime}{\isacharparenright}\ {\isasymle}\ degree\ {\isacharparenleft}last\ {\isacharparenleft}sturm\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}last\ {\isacharquery}ps{\isacharprime}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}\ auto{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}degree\ {\isacharparenleft}last\ {\isacharquery}ps{\isacharprime}{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{then}\isamarkupfalse%
\ \isacommand{obtain}\isamarkupfalse%
\ c\ \isakeyword{where}\ {\isachardoublequoteopen}last\ {\isacharquery}ps{\isacharprime}\ {\isacharequal}\ {\isacharbrackleft}{\isacharcolon}c{\isacharcolon}{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}last\ {\isacharquery}ps{\isacharprime}{\isachardoublequoteclose}{\isacharcomma}\ simp\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}x\ y{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharquery}ps{\isacharprime}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}last\ {\isacharquery}ps{\isacharprime}{\isacharparenright}\ y{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ squarefree{\isacharcolon}\ {\isachardoublequoteopen}rsquarefree\ {\isacharquery}p{\isacharprime}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ rsquarefree{\isacharunderscore}roots{\isacharcomma}\ unfold\ d{\isacharunderscore}def{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ intro\ allI\ coprime{\isacharunderscore}imp{\isacharunderscore}no{\isacharunderscore}common{\isacharunderscore}roots\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ Suc\ {\isadigit{0}}\ {\isacharequal}\ pderiv\ p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ sturm{\isacharunderscore}def\ d{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ coprime{\isacharcolon}\ {\isachardoublequoteopen}coprime\ {\isacharquery}p{\isacharprime}\ {\isacharparenleft}pderiv\ p\ div\ d{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ div{\isacharunderscore}gcd{\isacharunderscore}coprime{\isacharunderscore}poly\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ squarefree{\isacharprime}{\isacharcolon}\isanewline
\ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}x{\isachardot}\ {\isasymnot}\ {\isacharparenleft}poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymand}\ poly\ {\isacharparenleft}sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ coprime{\isacharunderscore}imp{\isacharunderscore}no{\isacharunderscore}common{\isacharunderscore}roots\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}signs{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymAnd}i\ x{\isachardot}\ {\isasymlbrakk}i\ {\isacharless}\ length\ {\isacharquery}ps{\isacharprime}\ {\isacharminus}\ {\isadigit{2}}{\isacharsemicolon}\ poly\ {\isacharparenleft}{\isacharquery}ps{\isacharprime}\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{1}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isasymrbrakk}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isasymLongrightarrow}\ poly\ {\isacharparenleft}{\isacharquery}ps{\isacharprime}\ {\isacharbang}\ {\isacharparenleft}i\ {\isacharplus}\ {\isadigit{2}}{\isacharparenright}{\isacharparenright}\ x\ {\isacharasterisk}\ poly\ {\isacharparenleft}{\isacharquery}ps{\isacharprime}\ {\isacharbang}\ i{\isacharparenright}\ x\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharquery}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ squarefree\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ rsquarefree{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}{\isacharquery}p{\isacharprime}\ {\isacharequal}\ {\isacharquery}ps{\isacharprime}\ {\isacharbang}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}pderiv\ p\ div\ d\ {\isacharequal}\ {\isacharquery}ps{\isacharprime}\ {\isacharbang}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ sturm{\isacharunderscore}def{\isacharcomma}\isanewline
\ \ \ \ \ \ \ \ \ \ subst\ sturm{\isacharunderscore}aux{\isachardot}simps{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharquery}ps{\isacharprime}\ {\isacharbang}\ {\isadigit{0}}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ squarefree\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ A\ rsquarefree{\isacharunderscore}def{\isacharparenright}\isanewline
\isanewline
\ \ \isacommand{fix}\isamarkupfalse%
\ x\isactrlsub {\isadigit{0}}\ {\isacharcolon}{\isacharcolon}\ real\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharquery}p{\isacharprime}\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}pderiv\ p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ dest{\isacharcolon}\ pderiv{\isacharunderscore}iszero{\isacharparenright}\isanewline
\ \ \isacommand{with}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}poly\ p\ x\isactrlsub {\isadigit{0}}\ {\isacharequal}\ {\isadigit{0}}{\isacharbackquoteclose}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p\ {\isacharasterisk}\ pderiv\ p{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ x\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ sturm{\isacharunderscore}firsttwo{\isacharunderscore}signs{\isacharunderscore}aux{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ ev\ {\isacharequal}\ eventually{\isacharunderscore}conj{\isacharbrackleft}OF\ A\ poly{\isacharunderscore}neighbourhood{\isacharunderscore}without{\isacharunderscore}roots{\isacharbrackleft}OF\ {\isacharbackquoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}{\isacharbrackright}\isanewline
\isanewline
\ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}eventually\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}p\ div\ d\ {\isacharasterisk}\ sturm{\isacharunderscore}squarefree{\isacharprime}\ p\ {\isacharbang}\ {\isadigit{1}}{\isacharparenright}\ x{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}if\ x\isactrlsub {\isadigit{0}}\ {\isacharless}\ x\ then\ {\isadigit{1}}\ else\ {\isacharminus}{\isadigit{1}}{\isacharparenright}{\isacharparenright}\ {\isacharparenleft}at\ x\isactrlsub {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}rule\ eventually{\isacharunderscore}mono{\isacharbrackleft}OF\ {\isacharunderscore}\ ev{\isacharbrackright}{\isacharcomma}\ clarify{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}intro{\isacharbrackright}{\isacharcolon}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a\ {\isacharparenleft}b{\isacharcolon}{\isacharcolon}real{\isacharparenright}{\isachardot}\ b\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ a\ {\isacharless}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ a\ {\isacharslash}\ {\isacharparenleft}b\ {\isacharasterisk}\ b{\isacharparenright}\ {\isacharless}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}{\isasymAnd}a\ {\isacharparenleft}b{\isacharcolon}{\isacharcolon}real{\isacharparenright}{\isachardot}\ b\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ a\ {\isachargreater}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ a\ {\isacharslash}\ {\isacharparenleft}b\ {\isacharasterisk}\ b{\isacharparenright}\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}{\isacharparenleft}case{\isacharunderscore}tac\ {\isachardoublequoteopen}b\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ auto\ simp{\isacharcolon}\ mult{\isacharunderscore}pos{\isacharunderscore}pos\ mult{\isacharunderscore}neg{\isacharunderscore}neg\ field{\isacharunderscore}simps{\isacharparenright}\ {\isacharbrackleft}{\isacharbrackright}{\isacharparenright}{\isacharplus}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ {\isacharparenleft}goal{\isadigit{1}}\ x{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ \ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}poly\ d\ x\ {\isacharasterisk}\ poly\ d\ x\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}poly\ d\ x\ {\isachargreater}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}\ auto\ simp{\isacharcolon}\ mult{\isacharunderscore}pos{\isacharunderscore}pos\ mult{\isacharunderscore}neg{\isacharunderscore}neg{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharunderscore}aux{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}pderiv\ p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}\ {\isasymlongleftrightarrow}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{moreover}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ dvd\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ dvd\ pderiv\ p{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \ \ \isacommand{ultimately}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}case\ \isacommand{using}\isamarkupfalse%
\ goal{\isadigit{1}}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ sgn{\isacharunderscore}real{\isacharunderscore}def\ poly{\isacharunderscore}div\ not{\isacharunderscore}less{\isacharbrackleft}symmetric{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ zero{\isacharunderscore}less{\isacharunderscore}divide{\isacharunderscore}iff\ split{\isacharcolon}\ split{\isacharunderscore}if{\isacharunderscore}asm{\isacharparenright}\isanewline
\ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Critically, unless \isa{x} is a multiple root of \isa{p} 
  (i.e. a root of both \isa{p} and its derivative), the number 
  of sign changes in the non-canonical Sturm sequence we defined 
  is the same as the number of sign changes in the canonical Sturm 
  sequence. Therefore we can use the canonical Sturm sequence 
  even in the non-squarefree case if the borders of the interval 
  we are interested in are not multiple roots of the polynomial.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}mult{\isacharunderscore}aux{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}d\ {\isasymnoteq}\ {\isacharparenleft}{\isadigit{0}}{\isacharcolon}{\isacharcolon}real{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}length\ {\isacharparenleft}group\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ {\isacharparenleft}op\ {\isacharasterisk}d\ {\isasymcirc}\ f{\isacharparenright}\ xs{\isacharparenright}{\isacharparenright}{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ length\ {\isacharparenleft}group\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ f\ xs{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ inj{\isacharcolon}\ {\isachardoublequoteopen}inj\ {\isacharparenleft}op\ {\isacharasterisk}d{\isacharparenright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ intro{\isacharcolon}\ injI{\isacharparenright}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ assms\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharparenleft}op{\isacharasterisk}\ d\ {\isasymcirc}\ f{\isacharparenright}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharequal}\ filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ f\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isachardoublequoteopen}filter\ {\isacharparenleft}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isasymcirc}\ f{\isacharparenright}\ {\isacharequal}\ filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ f\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharunderscore}all\ add{\isacharcolon}\ o{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isacharparenleft}map\ {\isacharparenleft}op{\isacharasterisk}\ d\ {\isasymcirc}\ f{\isacharparenright}\ xs{\isacharparenright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ map\ {\isacharparenleft}op{\isacharasterisk}\ d\ {\isasymcirc}\ f{\isacharparenright}\ {\isacharparenleft}filter\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharparenleft}op{\isacharasterisk}\ d\ {\isasymcirc}\ f{\isacharparenright}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ xs{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ filter{\isacharunderscore}map\ o{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ group{\isacharunderscore}map{\isacharunderscore}injective{\isacharbrackleft}OF\ inj{\isacharbrackright}\ assms\isanewline
\ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ filter{\isacharunderscore}map\ map{\isacharunderscore}map{\isacharbrackleft}symmetric{\isacharbrackright}\ del{\isacharcolon}\ map{\isacharunderscore}map{\isacharparenright}\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharcolon}\isanewline
\ \ \isakeyword{fixes}\ p\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{defines}\ {\isachardoublequoteopen}ps\ {\isasymequiv}\ sturm\ p{\isachardoublequoteclose}\ \isakeyword{and}\ {\isachardoublequoteopen}ps{\isacharprime}\ {\isasymequiv}\ sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymLongrightarrow}\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
{\isacharminus}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ p{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ s{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}poly{\isacharunderscore}inf\ d{\isachardoublequoteclose}\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ s{\isacharprime}{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}poly{\isacharunderscore}neg{\isacharunderscore}inf\ d{\isachardoublequoteclose}\isanewline
\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ {\isacharcolon}{\isacharcolon}\ real\ \isakeyword{and}\ q\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymin}\ set\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ dvd\ q{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ ps{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ sturm{\isacharunderscore}gcd\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ q{\isacharunderscore}prod{\isacharcolon}\ {\isachardoublequoteopen}q\ {\isacharequal}\ {\isacharparenleft}q\ div\ d{\isacharparenright}\ {\isacharasterisk}\ d{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ d{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ algebra{\isacharunderscore}simps\ dvd{\isacharunderscore}mult{\isacharunderscore}div{\isacharunderscore}cancel{\isacharparenright}\isanewline
\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}poly\ q\ x\ {\isacharequal}\ poly\ d\ x\ {\isacharasterisk}\ poly\ {\isacharparenleft}q\ div\ d{\isacharparenright}\ x{\isachardoublequoteclose}\ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ q{\isacharunderscore}prod{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ s{\isadigit{1}}{\isacharcolon}\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}\ {\isacharequal}\ sgn\ {\isacharparenleft}poly\ d\ x{\isacharparenright}\ {\isacharasterisk}\ sgn\ {\isacharparenleft}poly\ {\isacharparenleft}q\ div\ d{\isacharparenright}\ x{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ q{\isacharunderscore}prod{\isacharcomma}\ simp\ add{\isacharcolon}\ sgn{\isacharunderscore}mult{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}inf{\isacharunderscore}mult\ \isacommand{have}\isamarkupfalse%
\ s{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}poly{\isacharunderscore}inf\ q\ {\isacharequal}\ s{\isacharprime}\ {\isacharasterisk}\ poly{\isacharunderscore}inf\ {\isacharparenleft}q\ div\ d{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ s{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ q{\isacharunderscore}prod{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}inf{\isacharunderscore}mult\ \isacommand{have}\isamarkupfalse%
\ s{\isadigit{3}}{\isacharcolon}\ {\isachardoublequoteopen}poly{\isacharunderscore}neg{\isacharunderscore}inf\ q\ {\isacharequal}\ s{\isacharprime}{\isacharprime}\ {\isacharasterisk}\ poly{\isacharunderscore}neg{\isacharunderscore}inf\ {\isacharparenleft}q\ div\ d{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ s{\isacharprime}{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ q{\isacharunderscore}prod{\isacharcomma}\ simp{\isacharparenright}\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ s{\isadigit{1}}\ s{\isadigit{2}}\ s{\isadigit{3}}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ signs\ {\isacharequal}\ this\isanewline
\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ f\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly\ {\isasymRightarrow}\ real{\isachardoublequoteclose}\ \isakeyword{and}\ s\ {\isacharcolon}{\isacharcolon}\ real\isanewline
\ \ \ \ \isacommand{assume}\isamarkupfalse%
\ f{\isacharcolon}\ {\isachardoublequoteopen}{\isasymAnd}q{\isachardot}\ q\ {\isasymin}\ set\ ps\ {\isasymLongrightarrow}\ f\ q\ {\isacharequal}\ s\ {\isacharasterisk}\ f\ {\isacharparenleft}q\ div\ d{\isacharparenright}{\isachardoublequoteclose}\ \isakeyword{and}\ s{\isacharcolon}\ {\isachardoublequoteopen}s\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}inverse\ s\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isacommand{fix}\isamarkupfalse%
\ q\ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}q\ {\isasymin}\ set\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}f\ {\isacharparenleft}q\ div\ d{\isacharparenright}\ {\isacharequal}\ inverse\ s\ {\isacharasterisk}\ f\ q{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}subst\ f{\isacharbrackleft}of\ q{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all\ add{\isacharcolon}\ s{\isacharparenright}\isanewline
\ \ \ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ f{\isacharprime}\ {\isacharequal}\ this\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}length\ {\isacharparenleft}group\ {\isacharbrackleft}x{\isasymleftarrow}map\ f\ {\isacharparenleft}map\ {\isacharparenleft}{\isasymlambda}q{\isachardot}\ q\ div\ d{\isacharparenright}\ ps{\isacharparenright}{\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbrackright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ length\ {\isacharparenleft}group\ {\isacharbrackleft}x{\isasymleftarrow}map\ {\isacharparenleft}{\isasymlambda}q{\isachardot}\ f\ {\isacharparenleft}q\ div\ d{\isacharparenright}{\isacharparenright}\ ps\ {\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbrackright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ only{\isacharcolon}\ sign{\isacharunderscore}changes{\isacharunderscore}def\ o{\isacharunderscore}def\ map{\isacharunderscore}map{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}q{\isachardot}\ q\ div\ d{\isacharparenright}\ ps\ {\isacharequal}\ ps{\isacharprime}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ ps{\isacharunderscore}def\ ps{\isacharprime}{\isacharunderscore}def\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ f{\isacharprime}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}map\ {\isacharparenleft}{\isasymlambda}q{\isachardot}\ f\ {\isacharparenleft}q\ div\ d{\isacharparenright}{\isacharparenright}\ ps\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ map\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharparenleft}op{\isacharasterisk}{\isacharparenleft}inverse\ s{\isacharparenright}\ {\isasymcirc}\ f{\isacharparenright}\ x{\isacharparenright}\ ps{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ o{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}mult{\isacharunderscore}aux{\isacharbrackleft}OF\ {\isacharbackquoteopen}inverse\ s\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ of\ f\ ps{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ \isanewline
\ \ \ \ \ \ \ \ {\isachardoublequoteopen}length\ {\isacharparenleft}group\ {\isacharbrackleft}x{\isasymleftarrow}map\ f\ ps{\isacharprime}\ {\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbrackright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}\ {\isacharequal}\isanewline
\ \ \ \ \ \ \ \ \ length\ {\isacharparenleft}group\ {\isacharbrackleft}x{\isasymleftarrow}map\ f\ ps\ {\isachardot}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbrackright}{\isacharparenright}\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\ \ \isacommand{note}\isamarkupfalse%
\ length{\isacharunderscore}group\ {\isacharequal}\ this\isanewline
\isanewline
\ \ \isacommand{{\isacharbraceleft}}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{fix}\isamarkupfalse%
\ x\ \isacommand{assume}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ x\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ x\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ dvd\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}d\ dvd\ pderiv\ p{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ A\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}sgn\ {\isacharparenleft}poly\ d\ x{\isacharparenright}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ sgn{\isacharunderscore}zero{\isacharunderscore}iff\ elim{\isacharcolon}\ dvdE{\isacharparenright}\ \isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes\ ps{\isacharprime}\ x\ {\isacharequal}\ sign{\isacharunderscore}changes\ ps\ x{\isachardoublequoteclose}\ \isacommand{using}\isamarkupfalse%
\ signs{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}intro\ length{\isacharunderscore}group{\isacharbrackleft}of\ {\isachardoublequoteopen}{\isasymlambda}q{\isachardot}\ sgn\ {\isacharparenleft}poly\ q\ x{\isacharparenright}{\isachardoublequoteclose}{\isacharbrackright}{\isacharcomma}\ simp{\isacharunderscore}all{\isacharparenright}\isanewline
\ \ \isacommand{{\isacharbraceright}}\isamarkupfalse%
\isanewline
\isanewline
\ \ \isacommand{assume}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}s{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}s{\isacharprime}{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ s{\isacharprime}{\isacharunderscore}def\ s{\isacharprime}{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ length{\isacharunderscore}group{\isacharbrackleft}of\ poly{\isacharunderscore}inf\ s{\isacharprime}{\isacharcomma}\ OF\ signs{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ {\isacharbackquoteopen}s{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}inf{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ length{\isacharunderscore}group{\isacharbrackleft}of\ poly{\isacharunderscore}neg{\isacharunderscore}inf\ s{\isacharprime}{\isacharprime}{\isacharcomma}\ OF\ signs{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ {\isacharbackquoteopen}s{\isacharprime}{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{show}\isamarkupfalse%
\ {\isachardoublequoteopen}sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isacharprime}\ {\isacharequal}\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Root-counting functions%
}
\isamarkuptrue%
\isacommand{definition}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}between\ p\ a\ b\ {\isacharequal}\ {\isacharparenleft}if\ a\ {\isasymle}\ b\ {\isasymand}\ p\ {\isasymnoteq}\ {\isadigit{0}}\ then\ \isanewline
\ \ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ b{\isacharparenright}\ else\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ count{\isacharunderscore}roots\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}count{\isacharunderscore}roots\ p\ {\isacharequal}\ {\isacharparenleft}if\ {\isacharparenleft}p{\isacharcolon}{\isacharcolon}real\ poly{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\ else\isanewline
\ \ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ in\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}above\ p\ a\ {\isacharequal}\ {\isacharparenleft}if\ {\isacharparenleft}p{\isacharcolon}{\isacharcolon}real\ poly{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\ else\isanewline
\ \ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{definition}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below\ \isakeyword{where}\isanewline
{\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}below\ p\ a\ {\isacharequal}\ {\isacharparenleft}if\ {\isacharparenleft}p{\isacharcolon}{\isacharcolon}real\ poly{\isacharparenright}\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\ else\isanewline
\ \ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ in\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}correct{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}between\ p\ a\ b\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymand}\ a\ {\isasymle}\ b{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ False{\isacharprime}\ {\isacharequal}\ this\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}a\ {\isacharless}\ b{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ False\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ subset{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbraceleft}a{\isacharless}{\isachardot}{\isachardot}{\isacharless}b{\isacharbraceright}\ {\isasymsubseteq}\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ real{\isacharunderscore}infinite{\isacharunderscore}interval{\isacharbrackleft}OF\ True{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}finite\ {\isacharbraceleft}a{\isacharless}{\isachardot}{\isachardot}{\isacharless}b{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}finite\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ finite{\isacharunderscore}subset{\isacharbrackleft}OF\ subset{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \ \ \isacommand{with}\isamarkupfalse%
\ False{\isacharprime}\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp{\isacharcolon}\ not{\isacharunderscore}less\ card{\isacharunderscore}eq{\isacharunderscore}{\isadigit{0}}{\isacharunderscore}iff{\isacharparenright}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{unfolding}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ False\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ {\isachardoublequoteopen}a\ {\isasymle}\ b{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ p{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}p\ div\ {\isacharparenleft}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ clarsimp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p{\isachardoublequoteclose}\ p{\isacharprime}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}roots{\isacharunderscore}finite{\isacharbrackleft}OF\ {\isacharbackquoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}finite\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ fast\isanewline
\ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}between\ p\ a\ b\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}def\ Let{\isacharunderscore}def\isanewline
\ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ True\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}OF\ {\isacharbackquoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}a\ {\isasymle}\ b{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}correct{\isacharcolon}\isanewline
\ \ \isakeyword{fixes}\ p\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}count{\isacharunderscore}roots\ p\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}\ {\isacharequal}\ card\ {\isacharquery}S{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ real{\isacharunderscore}infinite{\isacharunderscore}interval{\isacharbrackleft}of\ {\isadigit{0}}\ {\isadigit{1}}{\isacharbrackright}\ finite{\isacharunderscore}subset{\isacharbrackleft}of\ {\isachardoublequoteopen}{\isacharbraceleft}{\isadigit{0}}{\isacharless}{\isachardot}{\isachardot}{\isacharless}{\isadigit{1}}{\isacharbraceright}{\isachardoublequoteclose}\ {\isacharquery}S{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}finite\ {\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}def\ True{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ p{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}p\ div\ {\isacharparenleft}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ clarsimp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p{\isachardoublequoteclose}\ p{\isacharprime}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharbrackleft}OF\ {\isacharbackquoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}count{\isacharunderscore}roots\ p\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}correct{\isacharcolon}\isanewline
\ \ \isakeyword{fixes}\ p\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}above\ p\ a\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}\ {\isacharequal}\ card\ {\isacharquery}S{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ real{\isacharunderscore}infinite{\isacharunderscore}interval{\isacharbrackleft}of\ a\ {\isachardoublequoteopen}a{\isacharplus}{\isadigit{1}}{\isachardoublequoteclose}{\isacharbrackright}\ finite{\isacharunderscore}subset{\isacharbrackleft}of\ {\isachardoublequoteopen}{\isacharbraceleft}a{\isacharless}{\isachardot}{\isachardot}{\isacharless}a{\isacharplus}{\isadigit{1}}{\isacharbraceright}{\isachardoublequoteclose}\ {\isacharquery}S{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}finite\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}def\ True{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ p{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}p\ div\ {\isacharparenleft}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ clarsimp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p{\isachardoublequoteclose}\ p{\isacharprime}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharbrackleft}OF\ {\isacharbackquoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}above\ p\ a\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ x\ {\isachargreater}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}correct{\isacharcolon}\isanewline
\ \ \isakeyword{fixes}\ p\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}real\ poly{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}below\ p\ a\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ {\isacharparenleft}\isakeyword{is}\ {\isachardoublequoteopen}{\isacharunderscore}\ {\isacharequal}\ card\ {\isacharquery}S{\isachardoublequoteclose}{\isacharparenright}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{with}\isamarkupfalse%
\ real{\isacharunderscore}infinite{\isacharunderscore}interval{\isacharbrackleft}of\ {\isachardoublequoteopen}a\ {\isacharminus}\ {\isadigit{1}}{\isachardoublequoteclose}\ a{\isacharbrackright}\ \isanewline
\ \ \ \ \ \ \ \ \ finite{\isacharunderscore}subset{\isacharbrackleft}of\ {\isachardoublequoteopen}{\isacharbraceleft}a\ {\isacharminus}\ {\isadigit{1}}{\isacharless}{\isachardot}{\isachardot}{\isacharless}a{\isacharbraceright}{\isachardoublequoteclose}\ {\isacharquery}S{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}finite\ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ force\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}def\ True{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \isacommand{def}\isamarkupfalse%
\ p{\isacharprime}\ {\isasymequiv}\ {\isachardoublequoteopen}p\ div\ {\isacharparenleft}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ clarsimp\isanewline
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree\ p{\isachardoublequoteclose}\ p{\isacharprime}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \isacommand{from}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharbrackleft}OF\ {\isacharbackquoteopen}p{\isacharprime}\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}below\ p\ a\ {\isacharequal}\ card\ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}def\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{from}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p{\isacharprime}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ p{\isacharprime}{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ blast\isanewline
\ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}code{\isacharbrackright}{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}between\ p\ a\ b\ {\isacharequal}\isanewline
\ \ \ \ \ {\isacharparenleft}let\ q\ {\isacharequal}\ pderiv\ p\isanewline
\ \ \ \ \ \ \ in\ if\ a\ {\isachargreater}\ b\ {\isasymor}\ p\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\isanewline
\ \ \ \ \ \ \ else\ if\ {\isacharparenleft}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ q\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isasymand}\ {\isacharparenleft}poly\ p\ b\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ q\ b\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ then\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm\ p\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ b{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ else\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ b{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}a\ {\isachargreater}\ b\ {\isasymor}\ p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}def\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ False{\isadigit{1}}\ {\isacharequal}\ this\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}a\ {\isasymle}\ b{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}{\isacharparenleft}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}\ {\isasymand}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharparenleft}poly\ p\ b\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ b\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ False{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ Let{\isacharunderscore}def\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isakeyword{and}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ B{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ b\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ b\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
\ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharprime}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharunderscore}correct\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ x\ {\isasymle}\ b\ {\isasymand}\ poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}between{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ {\isacharbackquoteopen}a\ {\isasymle}\ b{\isacharbackquoteclose}{\isacharcomma}\ symmetric{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ A{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ B{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ True\ False\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}code{\isacharbrackleft}code{\isacharbrackright}{\isacharcolon}\ \isanewline
\ \ {\isachardoublequoteopen}count{\isacharunderscore}roots\ {\isacharparenleft}p{\isacharcolon}{\isacharcolon}real\ poly{\isacharparenright}\ {\isacharequal}\isanewline
\ \ \ \ {\isacharparenleft}if\ p\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\ \isanewline
\ \ \ \ \ else\ let\ ps\ {\isacharequal}\ sturm\ p\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharcomma}\ simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharprime}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}correct\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ {\isacharbraceleft}x{\isachardot}\ poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ symmetric{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{3}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ False\ \isacommand{unfolding}\isamarkupfalse%
\ Let{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}code{\isacharbrackleft}code{\isacharbrackright}{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}above\ p\ a\ {\isacharequal}\isanewline
\ \ \ \ \ {\isacharparenleft}let\ q\ {\isacharequal}\ pderiv\ p\isanewline
\ \ \ \ \ \ \ in\ if\ p\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\isanewline
\ \ \ \ \ \ \ else\ if\ poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ q\ a\ {\isasymnoteq}\ {\isadigit{0}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ then\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm\ p\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ else\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes\ ps\ a\ {\isacharminus}\ sign{\isacharunderscore}changes{\isacharunderscore}inf\ ps{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}def\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ False{\isadigit{1}}\ {\isacharequal}\ this\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}{\isacharparenleft}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ False{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ Let{\isacharunderscore}def\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharprime}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharunderscore}correct\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isachardot}\ a\ {\isacharless}\ x\ {\isasymand}\ poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}above{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ symmetric{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ A{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ True\ False\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}code{\isacharbrackleft}code{\isacharbrackright}{\isacharcolon}\isanewline
\ \ {\isachardoublequoteopen}count{\isacharunderscore}roots{\isacharunderscore}below\ p\ a\ {\isacharequal}\isanewline
\ \ \ \ \ {\isacharparenleft}let\ q\ {\isacharequal}\ pderiv\ p\isanewline
\ \ \ \ \ \ \ in\ if\ p\ {\isacharequal}\ {\isadigit{0}}\ then\ {\isadigit{0}}\isanewline
\ \ \ \ \ \ \ else\ if\ poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ q\ a\ {\isasymnoteq}\ {\isadigit{0}}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ then\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm\ p\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a{\isacharparenright}\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ else\ {\isacharparenleft}let\ ps\ {\isacharequal}\ sturm{\isacharunderscore}squarefree\ p\isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ in\ sign{\isacharunderscore}changes{\isacharunderscore}neg{\isacharunderscore}inf\ ps\ {\isacharminus}\ sign{\isacharunderscore}changes\ ps\ a{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
\isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}p\ {\isacharequal}\ {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}def\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\isacommand{next}\isamarkupfalse%
\isanewline
\ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \isacommand{note}\isamarkupfalse%
\ False{\isadigit{1}}\ {\isacharequal}\ this\isanewline
\ \ \ \ \isacommand{hence}\isamarkupfalse%
\ {\isachardoublequoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp{\isacharunderscore}all\isanewline
\ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\isanewline
\ \ \ \ \isacommand{proof}\isamarkupfalse%
\ {\isacharparenleft}cases\ {\isachardoublequoteopen}{\isacharparenleft}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isacharparenright}{\isachardoublequoteclose}{\isacharparenright}\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ False\isanewline
\ \ \ \ \ \ \isacommand{thus}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ False{\isadigit{1}}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ Let{\isacharunderscore}def\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{next}\isamarkupfalse%
\isanewline
\ \ \ \ \isacommand{case}\isamarkupfalse%
\ True\isanewline
\ \ \ \ \ \ \isacommand{hence}\isamarkupfalse%
\ A{\isacharcolon}\ {\isachardoublequoteopen}poly\ p\ a\ {\isasymnoteq}\ {\isadigit{0}}\ {\isasymor}\ poly\ {\isacharparenleft}pderiv\ p{\isacharparenright}\ a\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{def}\isamarkupfalse%
\ d\ {\isasymequiv}\ {\isachardoublequoteopen}gcd\ p\ {\isacharparenleft}pderiv\ p{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}\ \isacommand{have}\isamarkupfalse%
\ {\isacharbrackleft}simp{\isacharbrackright}{\isacharcolon}\ {\isachardoublequoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isachardoublequoteclose}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}auto\ simp\ add{\isacharcolon}\ d{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \ \ \isacommand{from}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree{\isacharprime}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{interpret}\isamarkupfalse%
\ sturm{\isacharunderscore}seq{\isacharunderscore}squarefree\ {\isachardoublequoteopen}sturm{\isacharunderscore}squarefree{\isacharprime}\ p{\isachardoublequoteclose}\ {\isachardoublequoteopen}p\ div\ d{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}def\ Let{\isacharunderscore}def\ d{\isacharunderscore}def\ \isacommand{{\isachardot}}\isamarkupfalse%
\isanewline
\ \ \ \ \ \ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharunderscore}correct\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{have}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ p\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}\ {\isacharequal}\ \isanewline
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\isacharbraceleft}x{\isachardot}\ x\ {\isasymle}\ a\ {\isasymand}\ poly\ {\isacharparenleft}p\ div\ d{\isacharparenright}\ x\ {\isacharequal}\ {\isadigit{0}}{\isacharbraceright}{\isachardoublequoteclose}\isanewline
\ \ \ \ \ \ \ \ \ \ \isacommand{unfolding}\isamarkupfalse%
\ d{\isacharunderscore}def\ \isacommand{using}\isamarkupfalse%
\ poly{\isacharunderscore}div{\isacharunderscore}gcd{\isacharunderscore}squarefree{\isacharparenleft}{\isadigit{2}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\ \isacommand{by}\isamarkupfalse%
\ simp\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ count{\isacharunderscore}roots{\isacharunderscore}below{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ div\ d\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharcomma}\ symmetric{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{1}}{\isacharparenright}{\isacharbrackleft}OF\ A{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{also}\isamarkupfalse%
\ \isacommand{note}\isamarkupfalse%
\ sturm{\isacharunderscore}sturm{\isacharunderscore}squarefree{\isacharprime}{\isacharunderscore}same{\isacharunderscore}sign{\isacharunderscore}changes{\isacharparenleft}{\isadigit{3}}{\isacharparenright}{\isacharbrackleft}OF\ {\isacharbackquoteopen}p\ {\isasymnoteq}\ {\isadigit{0}}{\isacharbackquoteclose}{\isacharbrackright}\isanewline
\ \ \ \ \ \ \isacommand{finally}\isamarkupfalse%
\ \isacommand{show}\isamarkupfalse%
\ {\isacharquery}thesis\ \isacommand{using}\isamarkupfalse%
\ True\ False\ \isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}simp\ add{\isacharcolon}\ Let{\isacharunderscore}def{\isacharparenright}\isanewline
\ \ \ \ \isacommand{qed}\isamarkupfalse%
\isanewline
\isacommand{qed}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
\isanewline
%
\endisadelimproof
%
\isadelimtheory
\isanewline
%
\endisadelimtheory
%
\isatagtheory
\isacommand{end}\isamarkupfalse%
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End:

### 2013-11-18 paper/paper.tex
Lots of stuff

@@
 \renewcommand{\theta}{\vartheta}
-\newcommand{\atan}{\mathit{arctan}}
-\newcommand{\rot}{\mathit{rot}}
-\newcommand{\vdiv}{\mathit{div}}
+\newcommand{\atan}{\mathrm{arctan}}
+\newcommand{\rot}{\mathrm{rot}}
+\newcommand{\vdiv}{\mathrm{div}}
 \newcommand{\shouldbe}{\stackrel{!}{=}}
-\newcommand{\sgn}{\mathit{sgn}}
-%\renewcommand{\gcd}{\mathit{gcd}}
-%\renewcommand{\mod}{\ \mathit{mod}\ }
+\newcommand{\sgn}{\sgn}
+\renewcommand{\mod}{\ \ \textrm{mod}\ \ }
 
@@
 \item $n \geq 1$, i.e. the sequence is not empty
-\item $P_n$ does not change its sign, i.e. $\sgn(P_n(x))=\sgn(P_n(y))$ for all $x,y\in\RR$
+\item $P_{n-1}$ does not change its sign, i.e. $\sgn(P_{n-1}(x))=\sgn(P_{n-1}(y))$ for all $x,y\in\RR$
 \item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds
@@
 
-In Isabelle, these three concepts are captured in \emph{locales} of the same name.
+In Isabelle, these two concepts are captured in \emph{locales} of the same name.
 \vskip5mm
@@
 This is called the number of \emph{sign changes} of the sequence $P_0,\ldots,P_{n-1}$ at the position $x$ and is denoted as $\sigma(P_0,\ldots,P_{n-1}; x)$.\\
-In Isabelle, this is realised by evaluating the $P_i$ at $x$, deleting all zeros from the list, applying the \emph{rem\_adj\_dups} operation on the list and taking the length of the remaining list minus one as the result.
-The \emph{rem\_adj\_dups} function simply deletes equal adjacent elements in the list, \eg it turns $[1,2,2,1,1,1]$ into $[1,2,1]$.
+In Isabelle, this is realised by evaluating the $P_i$ at $x$, deleting all zeros from the list, applying the \emph{remdups\_adj} operation on the list and taking the length of the remaining list minus one as the result.
+The \emph{remdups\_adj} function simply deletes equal adjacent elements in the list, \eg it turns $[1,2,2,1,1,1]$ into $[1,2,1]$.
 
@@
 \begin{proof}
-According to the informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{rem\_adj\_dups} and \emph{filter} functions.
+According to the informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{remdups\_adj} and \emph{filter} functions.
 \end{proof}
@@
 Let $P_0,\ldots,P_{n-1}\in\RR[X]$, $Q_0\ldots,Q_{n-1}\in\RR[X]$ be two sequences of polynomials.
-Assume $\mathit{sgn}(P_i(x_1))=\mathit{sgn}(Q_i(x_2))$ for all $i$.
+Assume $\sgn(P_i(x_1))=\sgn(Q_i(x_2))$ for all $i$.
 Then $\sigma(P_0,\ldots,P_{n-1};x_1)=\sigma(Q_0,\ldots,Q_{n-1};x_2)$.
@@
 \end{proof}
+
+Later, we will also require some algebraic properties of polynomials, such as this:
+\begin{lemma}\label{thm:squarefree}
+Let $P \in \RR[X]\setminus \RR$, i.e. a nonconstant real polynomial. Let $D:=\gcd(P,P')$. Then the following holds:
+\begin{enumerate}
+\item $\gcd(P/D, P'/D) = 1$
+\item $\forall x\in\RR.\ (P/D)(x)=0 \Longleftrightarrow P(x)=0$, \ie $P$ and $P/D$ have the same roots, disregarding multiplicity
+\end{enumerate}
+\end{lemma}
 
@@
 \begin{lemma}\label{thm:nonzero}
-Let $P_0,\ldots,P_{n-1}$ be a quasi-Sturm sequence. Assume $P_0(x_0)\neq 0$. Then $\sigma(P_0,\ldots,P_n; x_0)$ is constant in the neighbourhood of $x_0$.\\
+Let $P_0,\ldots,P_{n-1}$ be a quasi-Sturm sequence. Assume $P_0(x_0)\neq 0$. Then $\sigma(P_0,\ldots,P_{n-1}; x_0)$ is constant in the neighbourhood of $x_0$.\\
 \end{lemma}
@@
 Also, the sequence $P_1,\ldots,P_{n-1}$ is again a quasi-Sturm sequence with $P_1(x_0)\neq 0$, \ie the induction hypothesis can be applied and we can conclude that $\sigma(P_1,\ldots,P_{n-1};x_0)$ is also constant in the neighbourhood of $x_0$.
-\item if the sequence has length $\geq 3$ and $P_1(x_0)=0$, we have $P_0(x_0)P_2(x_0)<0$ by the definition of a quasi-Sturm chain and thus $P_0(x_0)\neq 0$, $P_2(x_0)\neq 0$, and $\mathit{sgn}(P_2(x_0))=-\mathit{sgn}(P_0(x_0))$.
+\item if the sequence has length $\geq 3$ and $P_1(x_0)=0$, we have $P_0(x_0)P_2(x_0)<0$ by the definition of a quasi-Sturm chain and thus $P_0(x_0)\neq 0$, $P_2(x_0)\neq 0$, and $\sgn(P_2(x_0))=-\sgn(P_0(x_0))$.
 With Lemma \ref{thm:sign_changes_distrib}, we have $\sigma(P_0,\ldots,P_{n-1}; x_0) = \sigma(P_0,P_1,P_2;x_0) + \sigma(P_2,\ldots,P_{n-1}; x_0)$.
@@
 \begin{align*}
-&&\mathit{where}\quad &P_i = \begin{cases}
-\hskip3mm P & \mathit{for}\ i = 0\\
-\hskip3mm P' & \mathit{for}\ i = 1\\
--P_{i-2}\ \mathit{mod}\ P_{i-1} &\mathit{otherwise}
+&&\mathrm{where}\quad &P_i = \begin{cases}
+\hskip3mm P & \mathrm{for}\ i = 0\\
+\hskip3mm P' & \mathrm{for}\ i = 1\\
+-P_{i-2}\ \mathrm{mod}\ P_{i-1} &\mathrm{otherwise}
 \end{cases}&\\[2mm]
@@
 
-$n$ is chosen such that the $n\geq 2$ and $P_{n-1}$ is constant.
+$n$ is chosen such that $n\geq 2$ and $P_{n-1}$ is constant.
 The details of this are not important, since choosing a higher value for $n$ only results in a number of zero polynomials at the end of the sequence, which do not change the result in any way.
@@
 
-Of course, one way to handle a polynomial with multiple roots is to \enquote{divide away} the \enquote{excess roots} by dividing the polynomial $P$ by $\gcd(P,P')$.
+Of course, one way to handle a polynomial with multiple roots is to use lemma \ref{thm:squarefree} and \enquote{divide away} the \enquote{excess roots} by dividing the polynomial $P$ by $\gcd(P,P')$.
 However, surprisingly, it turns out that as long as the interval bounds $a$ and $b$ are not multiple roots of $P$ themselves, i.e.
 at least one of $P(a)\neq 0$ and $P'(a)\neq 0$ and at least one of $P(b)\neq 0$ and $P'(b)\neq 0$ holds, one can simply use the canonical Sturm chain without any modification.
-To prove this, we will first prove the following auxiliary result:
+To show this, we will first prove the following auxiliary result:
 
-%Lemma: Let $P\in\RR[X]\setminus\{0\}$.
+\begin{lemma}\label{thm:sturm_multiple_aux}
+Let $P\in\RR[X]\setminus\RR$, i.e.
+a nonconstant real polynomial.
 Let $P_0,\ldots,P_{n-1}$ be the result of the canonical Sturm chain construction of $P$ and define $Q_i := P_i / \gcd(P,P')$.
-Then $Q_0,\ldots,Q_{n-1}$ is a Sturm chain and $Q_0$ has the same roots, disregarding multiplicity, as $P_0$.\\
-%Proof:
-%\begin{itemize}
-%\item since $Q_0,\ldots,Q_{n-1}$ has the same length as $P_0,\ldots,P_{n-1}$, which is the result of the canonical Sturm chain construction, the length of $Q_0,\ldots,Q_{n-1}$ is $\geq 2$.
+Then $Q_i$ is a Sturm chain and $Q_0$ has the same roots, disregarding multiplicity, as $P$.
+\end{lemma}
+\begin{proof}
+First, we note that $\gcd(P,P')\neq 0$, since $P'\neq 0$ by assumption.
+Furthermore, lemma \ref{thm:sturm_gcd} implies that $\gcd(P,P')\ |\ P_i$ for all $i$, thus $Q_i\in\RR[X]$, i.e.
+the $Q_i$ really are polynomials.
+This means that our definition of $Q_0,\ldots,Q_{n-1}$ is well-defined.\\
+Now we show that the five properties for a Sturm chain are satisfied:
+\begin{itemize}
+\item the sequence has at least length 2: obvious by construction of $Q_0, \ldots, Q_{n+1}$
+\item $Q_{n-1}$ does not change its sign: obvious, since $P_{n-1}$ is constant, so $Q_{n-1}$ is, too.
+\item $Q_0$ and $Q_1$ have no common roots:\\
+By construction, we have $Q_0 = P / \gcd(P,P')$ and $Q_1 = (P / \gcd(P, P'))'$.
+This then implies $\gcd(Q_0, Q_1) = 1$.
+However, the proof for this algebraic fact is nontrivial; it was, of course, proven in the \emph{Isabelle} version of this proof, but it is too complex and not important enough to be discussed in detail here.
+\item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $Q_{i+1}$, $Q_{i}(x_0)Q_{i+2}(x_0) < 0$ holds:\\
+By construction of the canonical Sturm chain, we have:
+$$P_{i+2} = -P_{i}\ \mathrm{mod}\ P_{i+1}$$
+Therefore, we have:
+$$(P_{i}\ \mathrm{div}\ P_{i+1}) \cdot P_{i+1} - P_{i+2} = P_{i}$$
+Let $D:=\gcd(P,P')$. Since $D$ divides all the $P_j$; we have $P_{i}=D\cdot Q_{i}$, $P_{i+1}=D\cdot Q_{i+1}$, and $P_{i+2}=D\cdot Q_{i+2}$ and thus:
+$$(D\cdot Q_{i}\ \mathrm{div}\ D\cdot Q_{i+1}) \cdot D\cdot Q_{i+1} - D\cdot Q_{i+2} = D\cdot Q_{i}$$
+Cancelling $D$ yields:
+$$(Q_{i}\ \mathrm{div}\ Q_{i+1}) \cdot Q_{i+1} - Q_{i+2} = Q_{i}$$
+Since $x_0$ is a root of $Q_{i+1}(x_0)$, we then have:
+$$-Q_{i+2}(x_0) = Q_{i}(x_0)$$
+Since $\gcd(Q_{i}, Q_{i+1}) = 1$ and $x_0$ is a root of $Q_{i+1}(x_0)$, $x_0$ cannot be a root of $Q_{i}$ and thus we have:
+$$Q_{i}(x_0)Q_{i+2}(x_0) = -Q_{i}(x_0)^2 < 0$$
 
-%\end{itemize}
+\item for any root $x_0$ of $Q_0$, $Q_0(x)Q_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
+Let, again, $D:=\gcd(P, P')$. Obtain some an $\epsilon>0$ such that the following two properties hold:
+\begin{enumerate}
+\item $D$ does not have a root in $(x_0-\epsilon; x_0+\epsilon)\setminus\{x_0\}$
+\item $P_0(x)P_1(x)<0$ for all $x\in (x_0-\epsilon; x_0)$ and $P_0(x)P_1(x)>0$ for all $x\in (x_0;x_0+\epsilon)$ % TODO
+\end{enumerate}
+Then we have, for any $x$ in $(x_0-\epsilon; x_0+\epsilon)\setminus\{x_0\}$:
+$$P_0(x)P_1(x) = Q_0(x)Q_1(x)\cdot \underbrace{D(x)^2}_{>0}$$
+This then obviously implies the property we want to show.
+\end{itemize}
+It remains to show that $Q_0$ has the same roots, disregarding multiplicity, as $P = P_0$. This is precisely the statement of lemma \ref{thm:squarefree}.
+\end{proof}
+
+Of course, there is no reason to use this construction in practice, since if one has computed $\gcd(P,P')$ already, it is easier to compute the canonial Sturm chain of $P/\gcd(P,P')$ directly than to compute the canonical Sturm chain of $P$ and then divide every polynomial in it by $\gcd(P,P')$.\\
+However, the following construction becomes useful when combined with the following insight:
+\begin{lemma}\label{thm:sturm_multiple_aux2}
+Let $P\in\RR[X]\setminus\RR$, i.e. a nonconstant real polynomial. Let $P_0, \ldots, P_{n-1}$ and $Q_0,\ldots, Q_{n-1}$ be as in the previous lemma. Then the following holds:
+\begin{enumerate}
+\item $\forall x\in\RR.\ P(x)\neq 0\vee P'(x)\neq 0 \Longrightarrow \sigma(Q_0, \ldots, Q_{n-1}; x) = \sigma(P_0, \ldots, P_{n-1}; x)$
+\item $\sigma_\infty(Q_0, \ldots, Q_{n-1}) = \sigma_\infty(P_0, \ldots, P_{n-1})$
+\item $\sigma_{-\infty}(Q_0, \ldots, Q_{n-1}) = \sigma_{-\infty}(P_0, \ldots, P_{n-1})$
+\end{enumerate}
+\end{lemma}
+\begin{proof}
+Let $D:=\gcd(P,P')$. Consider an arbitrary $x\in\RR$ for which $P(x)\neq 0$ or $P'(x)\neq 0$, \ie $D(x)\neq 0$. Then we have for all $i$:
+$$\mathrm{sgn}(Q_i(x)) = \mathrm{sgn}(P_i(x) / D(x)) = \mathrm{sgn}(P_i(x)) \cdot \mathrm{sgn}(D(x))$$
+It is now easy to see that in both of the two cases $D(x)>0$ and $D(x)<0$, the number of sign changes in the two sequences is the same, since the signs in $Q_0,\ldots, Q_{n-1}$ are all either the same as in $P_0,\ldots, P_{n-1}$ or all flipped \wrt the signs in $P_0,\ldots,P_{n-1}$.\\
+Furthermore, we can find a lower bound $l$ (resp.
+an upper bound $u$) so that $D$ has no roots below $l$ (resp.
+above $u$), which implies that $\sigma(P_0,\ldots,P_{n-1}; x) = \sigma(Q_0,\ldots,Q_{n-1};x)$ for all $x$ beyond that bound, which implies that the limit is equal, \ie the $\sigma_\infty$ and $\sigma_{-\infty}$.
+\end{proof}
+
+In conclusion: if we want to count the roots of some $P$ between bounds $a,b\in\RR\cup\{-\infty,\infty\}$ with $a\leq b$, we use $\sigma$, $\sigma_\infty$ and $\sigma_{-\infty}$ with a Sturm chain of $P$, and such a Sturm chain can be obtained by applying the canonical Sturm chain construction to $P$ if neither $a$ nor $b$ are roots of both $P$ and $P'$, or by applying it to $P/\gcd(P,P')$ if they are.
+
+\section{Applications}
+
+The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that Sturm chains can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b)$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
+However, by adding case distinctions on whether $a$ and $b$ are roots of $P$, this can easily be generalised to arbitrary open, half-closed, closed, bounded, and unbounded real intervals.
+
+Furthermore, statements such as $\forall x \in I.\ P(x) \neq 0$ can be decided by counting the number of roots with the above method and checking whether it is zero.
+This, in turn, can be used to decide statements of the form $\forall x\in I.\ P(x) > 0$ by checking that $\forall x\in I.\ P(x) > 0$ holds and $\lim_{x\to\infty} P(x) = \infty$.
+
+By observing that for any two polynomials $P$ and $Q$, we can express the properties 
+\begin{align*}
+P(x) = 0 \wedge Q(x) = 0\hskip5mm&\mathrm{resp.}\hskip5mm P(x) = 0 \vee Q(x) = 0$$
+\gcd(P,Q) = 0\hskip5mm&\mathrm{resp.}\hskip5mm (P\cdot Q)(x) = 0$$
+we can further generalise our method to arbitrary statements of the form
+$$\forall x \in I.\ P(x)$$
+where $I$ is an open, half-closed, or closed, bounded or unbounded real interval and $P$ is a property consisting of the operators $\wedge$, $\vee$, $<$, $>$, $+$, $-$, $\cdot$, $^$, $\neq$
 

### 2013-11-18 count.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 n_sturm=`cat Sturm.thy Sturm_Library.thy | wc -l`
-n_method=`cat Sturm_Method.thy | wc -l`
+n_method=`cat Sturm_Method.thy *.ML | wc -l`
 n_analysis=`cat *Analysis*.thy | wc -l`
 n_poly=`cat *Poly*.thy | wc -l`
 n_misc=`cat List_Group.thy Misc.thy | wc -l`

### 2013-11-18 guide/guide.tex
Added Rational_Root.thy
Cleanup

@@
 \begin{align*}
-&\lemma\ "\card\ \{x::\real. (x - 1)^2 * (x + 1) = 0\} = 2"\ \textrm{\textbf{by}\ sturm}\\
-&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}. -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
+&\lemma\ "\card\ \{x::\real.\ (x - 1)^2 * (x + 1) = 0\} = 2"\ \textrm{\textbf{by}\ sturm}\\
+&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}.\ -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
 &\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"\ \textrm{\textbf{by}\ sturm}\\
-&\lemma\ "\card\ \{x::\real. x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\} = 1"\ \textrm{\textbf{by}\ sturm}\\
-&\lemma\ "\card\ \{x::\real. x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\} = 4"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\} = 1"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\} = 4"\ \textrm{\textbf{by}\ sturm}\\
 &\lemma\ "(x::\real)^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
@@
 
-Using this reduction, the \sturm\ method can also decide single \enquote{less than}/\enquote{ŋreater than} inequalities of the form 
+Using this reduction, the \sturm\ method can also decide single \enquote{less than}/\enquote{greater than} inequalities of the form 
 $$\forall x::\real.\ x > a\ \wedge\ x < b\ \longrightarrow\ p\ x < q\ x$$
@@
 \begin{enumerate}
-\item ensure that your function is indeed a \emph{real} polynomial. Add a type annotation to \emph{real} if necessary.
+\item ensure that your function is indeed a \emph{real} polynomial. Add an appropriate type annotation if necessary.
 \item use a computer algebra system to ensure that the property is indeed correct

### 2013-11-18 paper/paper.tex
Cleanup

@@
 \usepackage{xspace}
+\usepackage{stmaryrd}
 
@@
 \newcommand{\shouldbe}{\stackrel{!}{=}}
-\newcommand{\sgn}{\sgn}
+\newcommand{\sgn}{\mathrm{sgn}}
 \renewcommand{\mod}{\ \ \textrm{mod}\ \ }
+
+\newcommand{\sturm}{\texttt{sturm}}
+\newcommand{\ilemma}{\textbf{lemma}}
+\newcommand{\card}{\textrm{card}}
+\newcommand{\real}{\textrm{real}}
 
@@
 \newcommand{\eg}{e.\,g.\xspace}
+\newcommand{\wrt}{w.\,r.\,t.\xspace}
 
@@
 \maketitle
-\vskip-3mm
+\tableofcontents
+\newpage
 \parindent0mm
@@
 
-This paper will give a detailed proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.
+This paper will give a detailed proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.\\
+
+Section \ref{sec:notation} introduces some basic notation for well-known concepts that are not directly related to Sturm sequence.
+Section \ref{sec:defs} then defines a number of basic concepts that are used in Sturm's theorem, in particular, the concept of a \emph{Sturm sequence}.
+Section \ref{sec:aux_lemmas} contains some simple auxiliary lemmas that will be required in the main proof.
+Section \ref{sec:roots_sturm} contains the main portion of the proof, as it shows that Sturm sequences can be used to count the real roots of polynomials.
+Section \ref{sec:construction} then shows how Sturm sequences can be constructed in different cases and how these constructions can be applied.
+Finally, section \ref{sec:applications} explains how further preprocessing can be used to automatically decide complex, interesting properties on real polynomials.
 
 \section{Notation}
-In this paper, $\RR[X]$ will be used to denote the set of all polynomials with real coefficients.
-For a polynomial $P\in\RR[X]$, we denote its derivative by $P'$.
-We say that some property holds in a neighbourhood of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.
+\label{sec:notation}
+In our the following proofs, we will use the following notation and terminology:
+\begin{itemize}
+\item $\RR[X]$ denotes the set of all polynomials with real coefficients.
+\item $P'$ denotes the derivative of a polynomial $P\in\RR[X]$.
+\item Some property holds in a neighbourhood of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.
+Note that it does not have to hold \emph{at} $x_0$ \emph{itself}.
+\item $\mathrm{lc}(P)$ denotes the leading coefficient of a polynomial $P$.
+\end{itemize}
 
+\newpage
 \section{Proof of Sturm's theorem}
@@
 \subsection{Definitions}
+\label{sec:defs}
 
@@
 \item $n \geq 2$, i.e. the sequence contains at least two polynomials
-\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$\footnote{
+\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval\\ \mbox{$(x_0-\epsilon;x_0)$} and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$\footnote{
 In \isabellehol, this is expressed as \enquote{eventually, the property $\sgn(P_0P_1(x))\ =\ \mathsf{if}\ x>x_0\ \mathsf{then}\ 1\ \mathsf{else}\ -1$ holds at $x_0$}.}
@@
 and count the sign changes in these sequences in the same way as before.
-The number of sign changes is called $\sigma_{-\infty}(P_0,\ldots,P_{n-1})$ and $\sigma_{\infty}(P_0,\ldots,P_{n-1})$, respectively.
-Note that the signs of these limits for a $P_i$ can be determined by taking the sign of the leading coefficient of $P_i$.
+The number of sign changes is called $\sigma(P_0,\ldots,P_{n-1};-\infty)$ and $\sigma(P_0,\ldots,P_{n-1};\infty)$, respectively.
+Note that the signs of these limits for a $P_i$ can be determined by taking the sign of the leading coefficient of $P_i$.\footnote{For $\infty$ in general and for $-\infty$ and even degree, it is simply the sign of the leading coefficient; for $-\infty$ and odd degree, it is the opposite of the sign of the leading coefficient.}
 
 \subsection{Important auxiliary lemmas}
+\label{sec:aux_lemmas}
 
@@
 Lastly, should $l<u$ not hold already, we can simply increase $u$ until it does.
-\end{proof}
+\end{proof}\vskip3mm
 
 \begin{lemma}\label{thm:sign_changes_distrib}
-Let $P_0, \ldots, P_{n-1}$ be a sequence of polynomials and $i\in\{0,\ldots,n-1\}$.
+Let $P_0, \ldots, P_{n-1}$ be a sequence of polynomials and $i\in\{0,\ldots,n-1\}$ and $x\in\RR$.
 Furthermore, assume $P_i(x) \neq 0$.
@@
 According to the informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{remdups\_adj} and \emph{filter} functions.
-\end{proof}
+\end{proof}\vskip3mm
 
 \begin{lemma}\label{thm:sign_changes_triple}
-Assume $P(x)\neq 0$ and $\sgn(R(x)) = -\sgn(P(x))$. Then $\sigma(P,R,Q; x) = 1$
+Let $P,Q,R\in\RR[X]$ and $x\in\RR$ with $P(x)\neq 0$ and $\sgn(R(x)) = -\sgn(P(x))$. Then $\sigma(P,Q,R; x) = 1$
 \end{lemma}
@@
 By case distinction
-\end{proof}
+\end{proof}\vskip3mm
 
 \begin{lemma}\label{thm:same_signs}
-Let $P_0,\ldots,P_{n-1}\in\RR[X]$, $Q_0\ldots,Q_{n-1}\in\RR[X]$ be two sequences of polynomials.
+Let $P_0,\ldots,P_{n-1}\in\RR[X]$, $Q_0\ldots,Q_{n-1}\in\RR[X]$ be two sequences of polynomials and $x\in\RR$.
 Assume $\sgn(P_i(x_1))=\sgn(Q_i(x_2))$ for all $i$.
@@
 Trivial.
-\end{proof}
+\end{proof}\vskip3mm
 
-Later, we will also require some algebraic properties of polynomials, such as this:
+Later, we will also require some algebraic properties of polynomials, such as squarefree decomposition of polynomials:
 \begin{lemma}\label{thm:squarefree}
@@
 \end{enumerate}
-\end{lemma}
+\end{lemma}\vskip3mm
 
@@
 
-\subsection{Sign changes of Sturm sequences}
+\subsection{Relating roots and Sturm sequences}
+\label{sec:roots_sturm}
+
+We will now show how Sturm sequences can be used to determine the number roots of polynomials in a given interval.
+To this end, we will first explore how the number of sign changes of a Sturm sequence behaves in the neighbourhood of non-roots and roots of the polynomial.\\
 
 \begin{lemma}\label{thm:nonzero}
-Let $P_0,\ldots,P_{n-1}$ be a quasi-Sturm sequence. Assume $P_0(x_0)\neq 0$. Then $\sigma(P_0,\ldots,P_{n-1}; x_0)$ is constant in the neighbourhood of $x_0$.\\
+Let $P_0,\ldots,P_{n-1}$ be a quasi-Sturm sequence and $x_0\in\RR$. Assume $P_0(x_0)\neq 0$. Then $\sigma(P_0,\ldots,P_{n-1}; x_0)$ is constant in the neighbourhood of $x_0$.
 \end{lemma}
@@
 \item if the sequence has length 1, the number of sign changes is, of course, $0$ at any position.
-\item if the sequence has length 2, we know that $P_1$ does not change its sign anywhere by definition of a quasi-Sturm chain.
+\item if the sequence has length 2, we know that $P_1$ does not change its sign anywhere by definition of a quasi-Sturm sequence.
 Furthermore, since $P_0(x_0)\neq 0$, the sign of $P_0$ does not change in the neighbourhood of $x_0$.
 By Lemma \ref{thm:same_signs}, the number of sign changes of the sequence $P_0,P_1$ also remains constant in the neighbourhood of $x_0$.
-\item if the sequence has length $\geq 3$ and $P_1(x_0)\neq 0$, we have $\sigma(P_0,\ldots,P_{n-1}; x_0)=\sigma(P_0,P_1;x_0)+\sigma(P_1,\ldots,P_{n-1};x_0)$ from Lemma \ref{thm:sign_changes_distrib}.
-Since $P_0(x_0)\neq 0$ and $P_1(x_0)\neq 0$, the signs of $P_0$ and $P_1$ do not change in the neighbourhood of $x_0$ and therefore $\sigma(P_1,\ldots,P_{n-1};x_0)$ is constant in the neighbourhood of $x_0$.
+\item if the sequence has length $\geq 3$ and $P_1(x_0)\neq 0$, Lemma \ref{thm:sign_changes_distrib} implies $$\sigma(P_0,\ldots,P_{n-1}; x_0)=\sigma(P_0,P_1;x_0)+\sigma(P_1,\ldots,P_{n-1};x_0)$$ Since $P_0(x_0)\neq 0$ and $P_1(x_0)\neq 0$, the signs of $P_0$ and $P_1$ do not change in the neighbourhood of $x_0$ and therefore $\sigma(P_0, P_1;x_0)$ is constant in the neighbourhood of $x_0$.
 Also, the sequence $P_1,\ldots,P_{n-1}$ is again a quasi-Sturm sequence with $P_1(x_0)\neq 0$, \ie the induction hypothesis can be applied and we can conclude that $\sigma(P_1,\ldots,P_{n-1};x_0)$ is also constant in the neighbourhood of $x_0$.
-\item if the sequence has length $\geq 3$ and $P_1(x_0)=0$, we have $P_0(x_0)P_2(x_0)<0$ by the definition of a quasi-Sturm chain and thus $P_0(x_0)\neq 0$, $P_2(x_0)\neq 0$, and $\sgn(P_2(x_0))=-\sgn(P_0(x_0))$.
-With Lemma \ref{thm:sign_changes_distrib}, we have $\sigma(P_0,\ldots,P_{n-1}; x_0) = \sigma(P_0,P_1,P_2;x_0) + \sigma(P_2,\ldots,P_{n-1}; x_0)$.
-Furthermore, since $P_0(x_0)\neq 0$ and $P_2(x_0)\neq 0$, $\sgn(P_2(x))=-\sgn(P_0(x))$ holds in the entire neighbourhood of $x_0$ as the signs of $P_0$ and $P_2$ do not change in the neighbourhood of $x_0$, and with Lemma \ref{thm:sign_changes_triple}, we then have $\sigma(P_0,P_1,P_2;x)=2$ in the entire neighbourhood of $x_0$.
+\item if the sequence has length $\geq 3$ and $P_1(x_0)=0$, we have $P_0(x_0)P_2(x_0)<0$ by the definition of a quasi-Sturm sequence and thus $P_0(x_0)\neq 0$, $P_2(x_0)\neq 0$, and $\sgn(P_2(x_0))=-\sgn(P_0(x_0))$.
+With Lemma \ref{thm:sign_changes_distrib}, we have: $$\sigma(P_0,\ldots,P_{n-1}; x_0) = \sigma(P_0,P_1,P_2;x_0) + \sigma(P_2,\ldots,P_{n-1}; x_0)$$ Furthermore, since $P_0(x_0)\neq 0$ and $P_2(x_0)\neq 0$, $\sgn(P_2(x))=-\sgn(P_0(x))$ holds in the entire neighbourhood of $x_0$ as the signs of $P_0$ and $P_2$ do not change in the neighbourhood of $x_0$, and with Lemma \ref{thm:sign_changes_triple}, we then have $\sigma(P_0,P_1,P_2;x)=1$ (\ie constant) in the entire neighbourhood of $x_0$.
 As for the second summand, we know that $P_2,\ldots,P_{n-1}$ is again a quasi-Sturm sequence and $P_2(x_0)\neq 0$, so we can apply the induction hypothesis and obtain that $\sigma(P_2,\ldots,P_{n-1};x_0)$, too, is constant in the neighbourhood of $x_0$.
@@
 
+\newpage
 \label{thm:zero}
 \begin{lemma}
-Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence.
+Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence and $x_0\in\RR$.
 Assume $P_0(x_0)=0$.
-Then for $x$ in the neighbourhood of $x_0$, we have $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)+1$ if $x<x_0$ and $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)$ if $x\geq x_0$.\\
+Then for $x$ in the neighbourhood of $x_0$, we have $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)+1$ if $x<x_0$ and $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,\ldots,P_{n-1};x_0)$ if $x\geq x_0$.
 \end{lemma}
 \begin{proof}
-From the definition of a Sturm chain, we know that $P_0(x_0)=0$ implies $P_1(x_0)\neq 0$ and that in the neighbourhood of $x_0$, we have $P_0(x)P_1(x)<0$ if $x<x_0$ and $P_0(x)P_1(x)>0$ if $x>x_0$.
-Therefore, in the neighbourhood of $x_0$, we have $P_1(x)\neq 0$ and, with Lemma \ref{thm:nonzero}, $\sigma(P_1,\ldots,P_{n-1};x) = \sigma(P_1,\ldots,P_{n-1};x_0)$.\\
-Fix some $x$ in that neighbourhood of $x_0$.
-By Lemma \ref{thm:sign_changes_distrib}, we then have $\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x)+\sigma(P_1,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x)+\sigma(P_1,\ldots,P_{n-1};x_0)$.
-What remains to be shown is that $\sigma(P_0,P_1;x)=1$ for $x<x_0$ and $\sigma(P_0,P_1;x)=0$ for $x>x_0$.
-This is a simple consequence of $P_0(x)P_1(x)<0$ for $x<0$ and $P_0(x)P_1(x)>0$ for $x>0$.
-\end{proof}
+From the definition of a Sturm sequence, we know that $P_0(x_0)=0$ implies $P_1(x_0)\neq 0$.
+Therefore, $P_1(x_0)$ has the same nonzero sign in a neighbourhood of $x_0$.
+Moreover, in a neighbourhood of $x_0$, we have $P_0(x)P_1(x)<0$ if $x<x_0$ and $P_0(x)P_1(x)>0$ if $x>x_0$.
+With Lemma \ref{thm:sign_changes_distrib}, we have:
+$$\hskip1mm\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x)+\sigma(P_1,\ldots,P_{n-1};x) \hskip10mm\mathrm{in\ a\ NH\ of\ }x_0$$
+Since $P_1(x_0)\neq 0$ in a neighbourhood of $x_0$, we can apply Lemma \ref{thm:nonzero} and obtain:
+$$\hskip12mm\sigma(P_1,\ldots,P_{n-1};x) = \sigma(P_1,\ldots,P_{n-1};x_0) \hskip21mm\mathrm{in\ a\ NH\ of\ }x_0$$
+Of course, since $P_0(x_0)=0$, we also have:
+$$\sigma(P_1,\ldots,P_{n-1};x_0) = \sigma(P_0,\ldots,P_{n-1};x_0)\hskip31mm$$
+In summary, we have shown so far that:
+$$\sigma(P_0,\ldots,P_{n-1};x)=\sigma(P_0,P_1;x) + \sigma(P_0,\ldots,P_{n-1};x_0) \hskip10mm\mathrm{in\ a\ NH\ of\ }x_0$$
+Therefore, what remains to be shown is that $\sigma(P_0,P_1;x)=1$ for $x<x_0$ and $\sigma(P_0,P_1;x)=0$ for $x>x_0$ for all $x$ in a neighbourhood of $x_0$.
+This is a simple of $P_0(x)P_1(x)<0$ for $x<0$ and $P_0(x)P_1(x)>0$ for $x>0$ and the fact that $P_1$ has the same sign in the entire neighbourhood of $x_0$.
+\end{proof}\vskip5mm
+
+To express it in a less formal way: when passing through the real numbers in increasing order and tracking the number of sign changes of a Sturm seuqnece of a real polynomial $P$, that number increases by $1$ every time one passes a root of $P$ and remains constant otherwise.
+Intuitively, it should now be clear that Sturm sequences can be used to count the number of roots of a polynomial in a given interval.
+We will now prove this formally.
 
@@
 \begin{lemma}
-Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$. Fix $a,b\in\RR$ with $a \leq b$. Then $\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = |\{x\in (a;b]\ |\ P_0(x)=0\}|$\\
+Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$ and $x_0\in\RR$. Fix $a,b\in\RR$ with $a \leq b$. Then the following holds:
+$$\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = |\{x\in (a;b]\ |\ P_0(x)=0\}|$$
 \end{lemma}
@@
 Lemma \ref{thm:nonzero} then implies that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of all $x\in[a;b]$ and must therefore be constant in the entire interval $[a;b]$.
-\item if $P_0(a)=0$, we know from Lemma \ref{thm:zero} that $\sigma(P_0,\ldots,P_{n-1}; x)=\sigma(P_0,\ldots,P_{n-1}; a)$ for all $x\in(a;a+\epsilon)$ for some $\epsilon > 0$, \ie $\sigma(P_0,\ldots,P_{n-1})$ is constant in the right neighbourhood of $a$.
+\item if $P_0(a)=0$, we know from Lemma \ref{thm:zero} that $$\sigma(P_0,\ldots,P_{n-1}; x)=\sigma(P_0,\ldots,P_{n-1}; a)$$ for all $x\in(a;a+\epsilon)$ for some $\epsilon > 0$, \ie $\sigma(P_0,\ldots,P_{n-1})$ is constant in the right neighbourhood of $a$.
 Furthermore, using $P_0(x)\neq 0$ for all $x\in(a;b]$ and Lemma \ref{thm:nonzero}, that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of every $x\in(a;b]$.
@@
 \end{itemize}
-Since $\sigma(P_0,\ldots,P_{n-1})$ is constant on the entire interval $[a;b]$, we have $\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = 0$, which is exactly the number of roots of $P_0$ in $(a;b]$.
+Since $\sigma(P_0,\ldots,P_{n-1})$ is constant on the entire interval $[a;b]$, we have $$\sigma(P_0,\ldots,P_{n-1}; a)-\sigma(P_0,\ldots,P_{n-1}; b) = 0$$ which is exactly the number of roots of $P_0$ in $(a;b]$.
 
@@
 \begin{lemma}\label{thm:count_roots}
-Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$. Then $\sigma_{-\infty}(P_0,\ldots,P_{n-1})-\sigma_\infty(P_0,\ldots,P_{n-1}) = |\{x\ |\ P_0(x)=0\}|$
+Let $P_0,\ldots,P_{n-1}$ be a Sturm sequence with $P_0\neq 0$. Then the following holds: $$\sigma(P_0,\ldots,P_{n-1}; -\infty)-\sigma(P_0,\ldots,P_{n-1}; \infty) = |\{x\ |\ P_0(x)=0\}|$$
 \end{lemma}
@@
 
-Note: similar statements for the usage of $\sigma_{-\infty}$ and $\sigma_\infty$ to compute the number of roots $>a$ or $\leq a$ for some fixed $a\in\RR$ can be proven analogously.
+Note: similar statements for the usage of $\sigma$ at $\pm\infty$ to compute the number of roots $>a$, $<a$, $\geq a$, or $\leq a$ for some fixed $a\in\RR$ can be proven analogously.
 
-\section{Constructing Sturm chains}
+\newpage
+\section{Constructing Sturm sequences}
+\label{sec:construction}
+\subsection{The canonical Sturm sequence}
 
-\subsection{The canonical Sturm chain}
-
-The canonical Sturm chain $P_0,\ldots,P_{n-1}$ of a polynomial $P\in\RR[X]$ is constructed as follows:
+The canonical Sturm sequence $P_0,\ldots,P_{n-1}$ of a polynomial $P\in\RR[X]$ is constructed as follows:
 \begin{align*}
@@
 -P_{i-2}\ \mathrm{mod}\ P_{i-1} &\mathrm{otherwise}
-\end{cases}&\\[2mm]
+\end{cases}&
 \end{align*}
 
-$n$ is chosen such that $n\geq 2$ and $P_{n-1}$ is constant.
-The details of this are not important, since choosing a higher value for $n$ only results in a number of zero polynomials at the end of the sequence, which do not change the result in any way.
+$n$ is chosen such that $n\geq 2$ and $P_{n-1}$ is constant, since this will simplify formalisation.
+Choosing a higher value for $n$ only results in a number of zero polynomials at the end of the sequence, which do not change the result in any way.
 Note that this construction always terminates as the degree of the polynomials involved strictly decreases with every step, so that one reaches a constant polynomial in finitely many steps.\\
@@
 \begin{lemma}\label{thm:sturm_gcd}
-For all $i\in\{0,n-1\}$, we have $\gcd(P_i, P_{i+1}) = \gcd(P,P')$\\
+For all $i\in\{0,n-1\}$, we have $\gcd(P_i, P_{i+1}) = \gcd(P,P')$
 \end{lemma}
 \begin{proof}
-By induction on $i$ using the fact that $\gcd(R,S)=\gcd(S,R\mod S)=\gcd(S,-R\mod S)$.
-\end{proof}
+By induction on $i$ using the fact that $$\gcd(R,S)=\gcd(S,R\mod S)=\gcd(S,-R\mod S)$$\vskip-3mm
+\end{proof}\vskip3mm
 
 \begin{lemma}\label{thm:sturm_seq_canonical}
-Let $P\in\RR[X]$ with no multiple roots. Then the canonical Sturm chain construction of $P$ yields an actual Sturm chain.
+Let $P\in\RR[X]$ with no multiple roots. Then the canonical Sturm sequence construction of $P$ yields an actual Sturm sequence.
 \end{lemma}
 \begin{proof}
-We simply prove the five conditions a Sturm chain has to satisfy:
+We simply prove the five conditions a Sturm sequence has to satisfy:
 \begin{itemize}
@@
 With $P_{i+1}(x_0)=0$, this implies that $P_{i+2}(x) = -P_i(x)$, and because of Lemma \ref{thm:sturm_gcd} and the fact that $P$ has no multiple roots, we also have $P_i(x)\neq 0$ and thus $P_i(x)P_{i+2}(x)<0$.
-\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
-since $P_0=P$ and $P_1=P'$ are polynomials, there exists some neighbourhood of $x_0$ that does not contain any roots of either $P_0$ or $P_1$.
+\item for any root $x_0$ of $P_0$, $P_0(x)P_1(x)$ is negative in some sufficiently small interval\\ \mbox{$(x_0-\epsilon;x_0)$} and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
+Since $P_0=P$ and $P_1=P'$ are polynomials, there exists some neighbourhood of $x_0$ that does not contain any roots of either $P_0$ or $P_1$.
 For any $x<x_0$ in that neighbourhood, we can then apply the mean value theorem to obtain some $\xi\in[x;x_0]$ with $P(x)-P(x_0) = P'(\xi)(x-x_0)$, and since $P(x_0)=0$ and $x-x_0<0$, $P(x)=P'(\xi)$.
@@
 However, surprisingly, it turns out that as long as the interval bounds $a$ and $b$ are not multiple roots of $P$ themselves, i.e.
-at least one of $P(a)\neq 0$ and $P'(a)\neq 0$ and at least one of $P(b)\neq 0$ and $P'(b)\neq 0$ holds, one can simply use the canonical Sturm chain without any modification.
+at least one of $P(a)\neq 0$ and $P'(a)\neq 0$ and at least one of $P(b)\neq 0$ and $P'(b)\neq 0$ holds, one can simply use the canonical Sturm sequence without any modification.
 To show this, we will first prove the following auxiliary result:
@@
 a nonconstant real polynomial.
-Let $P_0,\ldots,P_{n-1}$ be the result of the canonical Sturm chain construction of $P$ and define $Q_i := P_i / \gcd(P,P')$.
-Then $Q_i$ is a Sturm chain and $Q_0$ has the same roots, disregarding multiplicity, as $P$.
+Let $P_0,\ldots,P_{n-1}$ be the result of the canonical Sturm sequence construction of $P$ and define $Q_i := P_i / \gcd(P,P')$.
+Then $Q_i$ is a Sturm sequence and $Q_0$ has the same roots, disregarding multiplicity, as $P$.
 \end{lemma}
@@
 This means that our definition of $Q_0,\ldots,Q_{n-1}$ is well-defined.\\
-Now we show that the five properties for a Sturm chain are satisfied:
+Now we show that the five properties for a Sturm sequence are satisfied:
 \begin{itemize}
@@
 \item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $Q_{i+1}$, $Q_{i}(x_0)Q_{i+2}(x_0) < 0$ holds:\\
-By construction of the canonical Sturm chain, we have:
+By construction of the canonical Sturm sequence, we have:
 $$P_{i+2} = -P_{i}\ \mathrm{mod}\ P_{i+1}$$
@@
 
-\item for any root $x_0$ of $Q_0$, $Q_0(x)Q_1(x)$ is negative in some sufficiently small interval $(x_0-\epsilon;x_0)$ and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
+\newpage
+\item for any root $x_0$ of $Q_0$, $Q_0(x)Q_1(x)$ is negative in some sufficiently small interval\\ \mbox{$(x_0-\epsilon;x_0)$} and positive in some sufficiently small interval $(x_0;x_0+\epsilon)$:\\
 Let, again, $D:=\gcd(P, P')$. Obtain some an $\epsilon>0$ such that the following two properties hold:
@@
 \end{itemize}
+\vskip3mm
 It remains to show that $Q_0$ has the same roots, disregarding multiplicity, as $P = P_0$. This is precisely the statement of lemma \ref{thm:squarefree}.
-\end{proof}
+\end{proof}\vskip10mm
 
-Of course, there is no reason to use this construction in practice, since if one has computed $\gcd(P,P')$ already, it is easier to compute the canonial Sturm chain of $P/\gcd(P,P')$ directly than to compute the canonical Sturm chain of $P$ and then divide every polynomial in it by $\gcd(P,P')$.\\
-However, the following construction becomes useful when combined with the following insight:
+Of course, there is no reason to use this construction in practice, since if one has computed $\gcd(P,P')$ already, it is easier to compute the canonial Sturm sequence of $P/\gcd(P,P')$ directly than to compute the canonical Sturm sequence of $P$ and then divide every polynomial in it by $\gcd(P,P')$.
+However, it becomes useful when combined with the following insight:
 \begin{lemma}\label{thm:sturm_multiple_aux2}
@@
 \item $\forall x\in\RR.\ P(x)\neq 0\vee P'(x)\neq 0 \Longrightarrow \sigma(Q_0, \ldots, Q_{n-1}; x) = \sigma(P_0, \ldots, P_{n-1}; x)$
-\item $\sigma_\infty(Q_0, \ldots, Q_{n-1}) = \sigma_\infty(P_0, \ldots, P_{n-1})$
-\item $\sigma_{-\infty}(Q_0, \ldots, Q_{n-1}) = \sigma_{-\infty}(P_0, \ldots, P_{n-1})$
+\item $P\neq 0 \Longrightarrow \sigma(Q_0,\ldots,Q_{n-1}; \pm\infty) = \sigma(P_9,\ldots,P_{n-1};\pm\infty)$
 \end{enumerate}
 \end{lemma}
-\begin{proof}
-Let $D:=\gcd(P,P')$. Consider an arbitrary $x\in\RR$ for which $P(x)\neq 0$ or $P'(x)\neq 0$, \ie $D(x)\neq 0$. Then we have for all $i$:
+\begin{proof}\ 
+\begin{enumerate}
+\item Let $D:=\gcd(P,P')$. Consider an arbitrary $x\in\RR$ for which $P(x)\neq 0$ or $P'(x)\neq 0$, \ie $D(x)\neq 0$. Then we have for all $i$:
 $$\mathrm{sgn}(Q_i(x)) = \mathrm{sgn}(P_i(x) / D(x)) = \mathrm{sgn}(P_i(x)) \cdot \mathrm{sgn}(D(x))$$
-It is now easy to see that in both of the two cases $D(x)>0$ and $D(x)<0$, the number of sign changes in the two sequences is the same, since the signs in $Q_0,\ldots, Q_{n-1}$ are all either the same as in $P_0,\ldots, P_{n-1}$ or all flipped \wrt the signs in $P_0,\ldots,P_{n-1}$.\\
-Furthermore, we can find a lower bound $l$ (resp.
-an upper bound $u$) so that $D$ has no roots below $l$ (resp.
-above $u$), which implies that $\sigma(P_0,\ldots,P_{n-1}; x) = \sigma(Q_0,\ldots,Q_{n-1};x)$ for all $x$ beyond that bound, which implies that the limit is equal, \ie the $\sigma_\infty$ and $\sigma_{-\infty}$.
+It is now easy to see that in both of the two cases $D(x)>0$ and $D(x)<0$, the number of sign changes in the two sequences is the same, since the signs in $Q_0,\ldots, Q_{n-1}$ are all either the same as in $P_0,\ldots, P_{n-1}$ or all flipped \wrt the signs in $P_0,\ldots,P_{n-1}$.
+\item We now consider the case of $\pm\infty$.
+Since $P_i = Q_i\cdot D$, we obviously have $\mathrm{lc}(P_i) = \mathrm{lc}(D)\cdot \mathrm{lc}(Q_i)$.
+When considering the way in which $\sigma$ at $\pm\infty$ can be computed using the leading coefficients, it is then again obvious that the two sign sequences are either the same or one is \enquote{flipped} \wrt the other; in either way, we have $\sigma(Q_0,\ldots,Q_{n-1};\pm\infty) = \sigma(P_0,\ldots,P_{n-1};\pm\infty)$.
+\end{enumerate}
 \end{proof}
 
-In conclusion: if we want to count the roots of some $P$ between bounds $a,b\in\RR\cup\{-\infty,\infty\}$ with $a\leq b$, we use $\sigma$, $\sigma_\infty$ and $\sigma_{-\infty}$ with a Sturm chain of $P$, and such a Sturm chain can be obtained by applying the canonical Sturm chain construction to $P$ if neither $a$ nor $b$ are roots of both $P$ and $P'$, or by applying it to $P/\gcd(P,P')$ if they are.
+In conclusion: if we want to count the roots of some $P$ between bounds $a,b\in\RR\cup\{-\infty,\infty\}$ with $a\leq b$, we use $\sigma$, $\sigma_\infty$ and $\sigma_{-\infty}$ with a Sturm sequence of $P$, and such a Sturm sequence can be obtained by applying the canonical Sturm sequence construction to $P$ if neither $a$ nor $b$ are roots of both $P$ and $P'$, or by applying it to $P/\gcd(P,P')$ if they are.
 
+\newpage
 \section{Applications}
+\label{sec:applications}
 
-The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that Sturm chains can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b)$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
+The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that Sturm sequences can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b)$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
 However, by adding case distinctions on whether $a$ and $b$ are roots of $P$, this can easily be generalised to arbitrary open, half-closed, closed, bounded, and unbounded real intervals.
@@
 
-By observing that for any two polynomials $P$ and $Q$, we can express the properties 
+By observing that for any two polynomials $P$ and $Q$, we have 
 \begin{align*}
-P(x) = 0 \wedge Q(x) = 0\hskip5mm&\mathrm{resp.}\hskip5mm P(x) = 0 \vee Q(x) = 0$$
-\gcd(P,Q) = 0\hskip5mm&\mathrm{resp.}\hskip5mm (P\cdot Q)(x) = 0$$
+P(x) = 0 \wedge Q(x) = 0\hskip5mm&\Longleftrightarrow\hskip5mm \gcd(P,Q) = 0\\
+P(x) = 0 \vee Q(x) = 0\hskip5mm&\Longleftrightarrow\hskip5mm (P\cdot Q)(x) = 0
+\end{align*}
 we can further generalise our method to arbitrary statements of the form
-$$\forall x \in I.\ P(x)$$
-where $I$ is an open, half-closed, or closed, bounded or unbounded real interval and $P$ is a property consisting of the operators $\wedge$, $\vee$, $<$, $>$, $+$, $-$, $\cdot$, $^$, $\neq$
+$$(\forall x \in I.\ P(x))\hskip5mm\mathrm{and}\hskip5mm |\{x\ |\ Q(x)\}| = n$$
+where $I$ is an open, half-closed, or closed, bounded or unbounded real interval and $P$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $\hat{\ }$, $\neq$ and $Q$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $\hat{\ }$, $=$.\\
+
+Independently from this, the observation that $\forall x\in I.\ P(x)>0$ holds iff $P$ has no roots in $I$ and $P(x)>0$ for some arbitrary $x\in I$ also allows us to decide statements of the form
+$$(\forall x\in I.\ P(x) > 0)\hskip5mm\mathrm{and}\hskip5mm(\forall x\in I.\ P(x) < 0)$$\vskip5mm
+
+Concrete examples of statements that can be proved in \isabellehol by the \emph{sturm} method that implements the decision procedure we just explained are:
+\footnote{For readers that are not familiar with Isabelle, \enquote{card} denotes the cardinality of a set, and \enquote{poly} constructs a polynomial from a list of coefficients, the first list element being the constant coefficient, the second one being the linear coefficient and so on.
+Moreover, in Isabelle, free variables in statements are implicitly universally quantified by convention.}
+\begin{align*}
+&\ilemma\ "\card\ \{x::\real.\ (x - 1)^2 * (x + 1) = 0\} = 2"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "\mathrm{card}\ \{x::\mathrm{real}.\ -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
+&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\} = 1"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\} = 4"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "(x::\real)^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "(x::\real) > 0\ \Longrightarrow\ x^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "\llbracket (x::\real) > 0; x \leq 2/3\rrbracket\ \Longrightarrow\ x*x \neq\ x"\ \textrm{\textbf{by}\ sturm}\\
+&\ilemma\ "(x::\real) > 1\ \Longrightarrow\ x*x > x"\ \textrm{\textbf{by}\ sturm}\\
+\end{align*}
 

### 2013-11-27 paper/paper.tex
Tuned paper

@@
 \maketitle
+
+\begin{abstract}
+Sturm sequences are a sequence of polynomials with certain properties that, as shown by Sturm's theorem, can be used to determine the number of roots of a real polynomial in a given interval.
+In this project, Sturm's theorem and a number of me\-thods to construct Sturm sequences have been formalised in Isabelle\slash HOL.
+Buil\-ding upon this, an Isabelle\slash HOL proof method was then implemented to prove statements about the number of roots of a real polynomial and related properties.
+\end{abstract}\vskip10mm
+
 \tableofcontents
@@
 
-This paper will give a detailed proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.\\
+This paper will give a detailed account of the formalisation, including a very explicit proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.\\
 
@@
 \item Some property holds in a neighbourhood of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.
-Note that it does not have to hold \emph{at} $x_0$ \emph{itself}.
+Note that it does \emph{not} have to hold \emph{at} $x_0$ \emph{itself}.
 \item $\mathrm{lc}(P)$ denotes the leading coefficient of a polynomial $P$.

### 2013-12-02 paper/paper.tex
Tuned paper

--- 2013-11-18 paper/paper.tex
@@
 
-\usepackage[utf8x]{inputenc} 
+\usepackage[utf8]{inputenc} 
 \usepackage[T1]{fontenc}
@@
 \usepackage{stmaryrd}
+\usepackage{hyperref}
+
+\usepackage[style=alphabetic]{biblatex}
 
 \usepackage{tgpagella}
-
-\DeclareFontFamily{OT1}{pzc}{}
-\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
-\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
-\DeclareUnicodeCharacter{10230}{$\rightarrow$}
 
@@
 \geometry{a4paper,left=30mm,right=30mm, top=25mm, bottom=30mm} 
+\bibliography{sturm}
 
@@
 \maketitle
+
+\begin{abstract}
+\emph{Sturm sequences} are a method of efficiently computing the number of real roots of a real polynomial inside a given interval.
+In this project, this fact and a number of me\-thods to construct Sturm sequences efficiently have been formalised with the interactive theorem prover Isabelle\slash HOL.
+Buil\-ding upon this, an Isabelle\slash HOL proof method was then implemented to prove statements about the number of roots of a real polynomial and related properties.
+\end{abstract}\vskip10mm
+
 \tableofcontents
@@
 
-\section{Outline}
+\section{Introduction}
+\emph{Sturm sequences} are finite sequences of polynomials with certain properties that, as shown by \emph{Sturm's Theorem}, can be used to determine the number of roots of a real polynomial in a given interval alorithmically.
+This is used to prove properties about the roots of specific polynomials directly, but can also be part of other algorithms such as approximation of roots through bisection, and it forms the basis of \textsc{Tarski}'s decision procedure for real arithmetic.
+Implementations of Sturm sequences for these purposes exist in a number of computer algebra systems, %TODO check this
+\\
 
-This paper will give a detailed proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about analysis and polynomials.\\
+There are different ways to construct Sturm sequences for a fixed polynomial $P$.
+The \enquote{canonical} Sturm sequence construction works only if $P$ has no multiple roots, but it can be adapter to the more general case of any nonzero polynomial.\\
+
+Both Sturm's Theorem and these constructions were formally proven in \isabellehol in this project.
+This paper will give a detailed account of the formalisation, including a very explicit proof that closely follows the structure of the formal proof in \isabellehol, but omits the proofs for basic lemmas about real analysis and polynomials.\\
 
@@
 Section \ref{sec:construction} then shows how Sturm sequences can be constructed in different cases and how these constructions can be applied.
-Finally, section \ref{sec:applications} explains how further preprocessing can be used to automatically decide complex, interesting properties on real polynomials.
+Finally, section \ref{sec:applications} explains how further preprocessing can be used to automatically decide more complex, interesting properties on real polynomials.
 
@@
 \item $P'$ denotes the derivative of a polynomial $P\in\RR[X]$.
-\item Some property holds in a neighbourhood of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.
-Note that it does not have to hold \emph{at} $x_0$ \emph{itself}.
+\item Some property holds \emph{in a neighbourhood} of some $x_0\in\RR$ if there exists an $\epsilon > 0$ such that the property holds for all $x\in (x_0-\epsilon;x_0+\epsilon)\setminus\{x_0\}$.
+Note that it does \emph{not} have to hold \emph{at} $x_0$ \emph{itself}.
 \item $\mathrm{lc}(P)$ denotes the leading coefficient of a polynomial $P$.
@@
 
-First, we define three notions that will be important in the proof of Sturm's theorem:
+First, we define two notions that will be important in the proof of Sturm's theorem:
 
-A \emph{quasi-Sturm sequence} is a list of polynomials $P_0,P_1,\ldots,P_{n-1} \in \RR[X]$ that fulfils the following properties:
+A \emph{quasi-Sturm sequence}\footnote{This term is, to our knowledge, not used in any of the literature.
+It was \enquote{invented} solely for the purpose of this formal proof, since we need to perform explicit induction on the sequence, which does not preserve the Sturm sequence property, but does preserve the weaker notion of quasi-Sturm sequence} is a list of polynomials $P_0,P_1,\ldots,P_{n-1} \in \RR[X]$ that fulfils the following properties:
 \begin{itemize}
@@
 \item $P_{n-1}$ does not change its sign, i.e. $\sgn(P_{n-1}(x))=\sgn(P_{n-1}(y))$ for all $x,y\in\RR$
-\item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds
+\item for any $i\in\{0,\ldots,n-3\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds
 \end{itemize}
@@
 
-A \emph{Sturm sequence} is a quasi-Sturm sequence that fulfils the following additional properties:
+A \emph{Sturm sequence} is a quasi-Sturm sequence that fulfils the following additional properties:\footnote{This definition was adapted from \cite{sag} as it has the virtue of being very general, allowing us to easily generalise the canonical construction later.}
 \begin{itemize}
@@
 
-Since Sturm's theorem is based on counting sign changes, a crucial part in proving it is to formalise the counting of sign changes in a way that is convenient for subsequent proofs.
-In this proof, this was done by proving that under certain conditions, the computation of the number of sign changes of a sequence can be \enquote{decomposed}.
+The first auxiliary lemma concerns the behaviour of polynomials at infinity and will be important for counting roots in intervals of infinite length.
 
@@
 a $u$) such that for all $x\leq l$ (resp.
-$x\geq u$), $P(x)$ increases above every bound in the case of the limit $\infty$ or falls below every bound in case of the limit $-\infty$.
-In particular, if we choose 0 as the bound, we see that $P(x)$ will be positive for all values beyond some point in the case of $\infty$ or negative in the case of $-\infty$, which implies $\sgn(P(x)) = \sgn(\lim_{x\to -\infty} P(x))$ (resp.
+$x\geq u$), $P(x)$ increases above \mbox{every} bound in the case of the limit $\infty$ or falls below every bound in case of the limit $-\infty$.
+In particular, if we choose 0 as the bound, we see that $P(x)$ will be positive for all \mbox{values} beyond some point in the case of $\infty$ or negative in the case of $-\infty$, which implies $\sgn(P(x)) = \sgn(\lim_{x\to -\infty} P(x))$ (resp.
 $\sgn(P(x)) = \sgn(\lim_{x\to \infty} P(x))$) for all $x$ beyond that point.
@@
 
+Since Sturm's theorem is based on counting sign changes, a crucial part in proving it is to formalise the counting of sign changes in a way that is convenient for subsequent proofs.
+In this proof, this was done by proving that under certain conditions, the computation of the number of sign changes of a sequence can be \enquote{decomposed}.
+
 \begin{lemma}\label{thm:sign_changes_distrib}
@@
 \begin{proof}
-According to the informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{remdups\_adj} and \emph{filter} functions.
+For our informal definition of sign changes, this is obvious; for the Isabelle definition using list operation, it can be proven using simple properties of the \emph{remdups\_adj} and \emph{filter} functions.
 \end{proof}\vskip3mm
@@
 
-Later, we will also require some algebraic properties of polynomials, such as squarefree decomposition of polynomials:
+Later, we will also require some algebraic properties of polynomials, such as squarefree decomposition of polynomials. 
 \begin{lemma}\label{thm:squarefree}
@@
 \begin{enumerate}
-\item $\gcd(P/D, P'/D) = 1$
+\item $\gcd(P/D, P'/D) = 1$, \ie $P/D$ and $P'/D$ are coprime
 \item $\forall x\in\RR.\ (P/D)(x)=0 \Longleftrightarrow P(x)=0$, \ie $P$ and $P/D$ have the same roots, disregarding multiplicity
@@
 Therefore, what remains to be shown is that $\sigma(P_0,P_1;x)=1$ for $x<x_0$ and $\sigma(P_0,P_1;x)=0$ for $x>x_0$ for all $x$ in a neighbourhood of $x_0$.
-This is a simple of $P_0(x)P_1(x)<0$ for $x<0$ and $P_0(x)P_1(x)>0$ for $x>0$ and the fact that $P_1$ has the same sign in the entire neighbourhood of $x_0$.
+This is a simple consequence of $P_0(x)P_1(x)<0$ for $x<0$ and $P_0(x)P_1(x)>0$ for $x>0$ and the fact that $P_1$ has the same sign in the entire neighbourhood of $x_0$.
 \end{proof}\vskip5mm
 
-To express it in a less formal way: when passing through the real numbers in increasing order and tracking the number of sign changes of a Sturm seuqnece of a real polynomial $P$, that number increases by $1$ every time one passes a root of $P$ and remains constant otherwise.
+To express it in a less formal way: when passing through the real numbers in increasing order and tracking the number of sign changes of a Sturm sequence of a real polynomial $P$, that number increases by $1$ every time one passes a root of $P$ and remains constant otherwise.
 Intuitively, it should now be clear that Sturm sequences can be used to count the number of roots of a polynomial in a given interval.
@@
 \begin{proof}
-By induction over $|\{x\in (a;b].\ P_0(x)=0\}|$ for arbitrary $a, b$.
+By induction over $k:=|\{x\in (a;b].\ P_0(x)=0\}|$ for arbitrary $a, b$.
 \begin{itemize}
-\item if $|\{x\in (a;b].\ P_0(x)=0\}|=0$, we have $P_0(x)\neq 0$ for all $x\in (a;b]$. We now distinguish two cases:
+\item if $k=0\}|=0$, we have $P_0(x)\neq 0$ for all $x\in (a;b]$. We now distinguish two cases:
 \begin{itemize}
 \item if $P_0(a)\neq 0$, we have $P_0(x)\neq 0$ for all $x\in[a;b]$.
-Lemma \ref{thm:nonzero} then implies that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of all $x\in[a;b]$ and must therefore be constant in the entire interval $[a;b]$.
+Lemma \ref{thm:nonzero} then implies that $\sigma(P_0,\ldots,P_{n-1})$ is constant in the neighbourhood of all $x\in[a;b]$ and must therefore be constant in the entire interval $[a;b]$ as polynomials are continuous.
 \item if $P_0(a)=0$, we know from Lemma \ref{thm:zero} that $$\sigma(P_0,\ldots,P_{n-1}; x)=\sigma(P_0,\ldots,P_{n-1}; a)$$ for all $x\in(a;a+\epsilon)$ for some $\epsilon > 0$, \ie $\sigma(P_0,\ldots,P_{n-1})$ is constant in the right neighbourhood of $a$.
@@
 
-\item if $|\{x\in (a;b].\ P_0(x)=0\}|=n>0$, we take the smallest root of $P_0$ in the interval $(a;b]$ and call it $x_2$.
-With Lemma \ref{thm:zero}, we know that $\sigma(P_0,\ldots,P_{n-1};x) = \sigma(P_0,\ldots,P_{n-1};x_2)+1$ for all $x$ in a sufficiently small left neighbourhood of $x_2$.
-Let $x_1$ then be some value in $(a;x_2)$ that is in this neighbourhood.
-We know that $P_0$ has no roots in the interval $(a;x_2]$ and thus, by induction hypothesis, $\sigma(P_0,\ldots,P_{n-1};a)=\sigma(P_0,\ldots,P_{n-1};x_1)=\sigma(P_0,\ldots,P_{n-1};x_2)+1$.
-Furthermore, the number of roots of $P_0$ in the interval $(x_2;b]$ is exactly $n-1$, so by induction hypothesis, $\sigma(P_0,\ldots,P_{n-1};x_2)-\sigma(P_0,\ldots,P_{n-1};b) = n-1$.
-Adding these two equations yields $\sigma(P_0,\ldots,P_{n-1};a)-\sigma(P_0,\ldots,P_{n-1};b) = n$.
+\item if $k>0$, we take the smallest root of $P_0$ in the interval $(a;b]$ and call it $x_2$.
+With Lemma \ref{thm:zero}, we know that for all $x$ in a sufficiently small left neighbourhood of $x_2$: $$\sigma(P_0,\ldots,P_{n-1};x) = \sigma(P_0,\ldots,P_{n-1};x_2)+1$$  Let $x_1$ then be some value in $(a;x_2)$ that is in this neighbourhood.
+We know that $P_0$ has no roots in the interval $(a;x_2]$ and thus, by induction hypothesis, $$\sigma(P_0,\ldots,P_{n-1};a)=\sigma(P_0,\ldots,P_{n-1};x_1)=\sigma(P_0,\ldots,P_{n-1};x_2)+1$$ Furthermore, the number of roots of $P_0$ in the interval $(x_2;b]$ is exactly $k-1$, so by induction hypothesis, $$\sigma(P_0,\ldots,P_{n-1};x_2)-\sigma(P_0,\ldots,P_{n-1};b) = k-1$$ Adding these two equations yields $$\sigma(P_0,\ldots,P_{n-1};a)-\sigma(P_0,\ldots,P_{n-1};b) = k$$
 \end{itemize}
@@
 \begin{proof}
-Using Lemma \ref{thm:limits} for every $P_i$, we obtain $l,u\in\RR$ with $l<u$ such that $P_0$ has no roots outside the interval $(l;u)$ and for all $P_i$, $\sgn(P_i(l))=\sgn(\lim_{x\to -\infty} P_i(x))$ and $\sgn(P_i(u))=\sgn(\lim_{x\to \infty} P_i(x))$.
-It is then easy to see that the number of sign changes of the sequence $P_0,\ldots,P_{n-1}$ can be determined by considering the signs of the limits of the $P_i$ at $\pm\infty$ instead of $P_i(u)$ and $P_i(l)$.
+Using Lemma \ref{thm:limits} for every $P_i$ and taking the smallest lower bound $l$ and largest upper bound $u$, we obtain $l,u\in\RR$ with $l<u$ such that $P_0$ has no roots outside the interval $(l;u)$ and for all $P_i$: $$\sgn(P_i(l))=\sgn\left(\lim_{x\to -\infty} P_i(x)\right)\hskip5mm\mathrm{and}\hskip5mm\sgn(P_i(u))=\sgn\left(\lim_{x\to \infty} P_i(x)\right)$$ It is then easy to see that the number of sign changes of the sequence $P_0,\ldots,P_{n-1}$ at $l$ resp.
+$u$ can be determined by considering the signs of the limits of the $P_i$ at $\pm\infty$ instead of $P_i(u)$ and $P_i(l)$.
 Since all roots of $P_0$ lie in the interval $(l;u)$, we have thus shown that the total number of roots of $P_0$ can be determined in the way described above.
@@
 \begin{align*}
-&&\mathrm{where}\quad &P_i = \begin{cases}
+P_i = \begin{cases}
 \hskip3mm P & \mathrm{for}\ i = 0\\
@@
 \begin{lemma}\label{thm:sturm_gcd}
-For all $i\in\{0,n-1\}$, we have $\gcd(P_i, P_{i+1}) = \gcd(P,P')$
+For all $i\in\{0,n-2\}$, we have $\gcd(P_i, P_{i+1}) = \gcd(P,P')$
 \end{lemma}
@@
 \item $P_{n-1}$ does not change its sign: obvious, since $P_{n-1}$ is constant
-\item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds:\\
-By construction, we have $P_{i+2} = -P_{i}\ \textit{mod}\ P_{i+1}$.
+\item for any $i\in\{0,\ldots,n-2\}$ and any root $x_0$ of $P_{i+1}$, $P_{i}(x_0)P_{i+2}(x_0) < 0$ holds:\\
+By construction, we have $P_{i+2} = -P_{i}\ \textit{mod}\ P_{i+1}$ and thus $P_i = P_{i+1}Q - P_{i+2}$ for some $Q\in\RR[X]$.
 With $P_{i+1}(x_0)=0$, this implies that $P_{i+2}(x) = -P_i(x)$, and because of Lemma \ref{thm:sturm_gcd} and the fact that $P$ has no multiple roots, we also have $P_i(x)\neq 0$ and thus $P_i(x)P_{i+2}(x)<0$.
@@
 Since $P_0=P$ and $P_1=P'$ are polynomials, there exists some neighbourhood of $x_0$ that does not contain any roots of either $P_0$ or $P_1$.
-For any $x<x_0$ in that neighbourhood, we can then apply the mean value theorem to obtain some $\xi\in[x;x_0]$ with $P(x)-P(x_0) = P'(\xi)(x-x_0)$, and since $P(x_0)=0$ and $x-x_0<0$, $P(x)=P'(\xi)$.
-Furthermore, $P(x)\neq 0$ and $P'(\xi)$ has the same sign as $P'(x)$, since we chose a neighbourhood without roots; therefore we have $P(x)P'(x)<0$.
-For any $x>x_0$, we can use the same argument to find that $P(x)P'(x)>0$.
+For any $x<x_0$ in that neighbourhood, we can then apply the mean value theorem to obtain some $\xi\in[x;x_0]$ with 
+$$P'(\xi)(x-x_0) = P(x)-P(x_0) = P(x)$$
+Due to $x < x_0$ $P(x)$, this implies $\sgn(P(x))=-\sgn(P'(\xi))$.
+Furthermore, recall that we chose a neighbourhood of $x_0$ that contains no roots of $P$ or $P'$; therefore, $0\neq \sgn(P'(\xi))=\sgn(P'(x))=-\sgn(P(x))$.
+This directly implies $P(x)P'(x)<0$.
+Similarly, for any $x>x_0$ in that neighbourhood, we can use the same argument to find that $P(x)P'(x)>0$.
 \item $P_0$ and $P_1$ have no common roots:\\
@@
 \begin{itemize}
-\item the sequence has at least length 2: obvious by construction of $Q_0, \ldots, Q_{n+1}$
+\item the sequence has at least length 2: obvious by construction of $Q_0, \ldots, Q_{n-1}$
 \item $Q_{n-1}$ does not change its sign: obvious, since $P_{n-1}$ is constant, so $Q_{n-1}$ is, too.
@@
 This then implies $\gcd(Q_0, Q_1) = 1$.
-However, the proof for this algebraic fact is nontrivial; it was, of course, proven in the \emph{Isabelle} version of this proof, but it is too complex and not important enough to be discussed in detail here.
+However, the proof for this algebraic fact is nontrivial; it was, of course, proven in the \emph{Isabelle} version of this proof, but it is too complex and at the same time not important enough to be discussed in detail here.
 \item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $Q_{i+1}$, $Q_{i}(x_0)Q_{i+2}(x_0) < 0$ holds:\\
@@
 $$(P_{i}\ \mathrm{div}\ P_{i+1}) \cdot P_{i+1} - P_{i+2} = P_{i}$$
-Let $D:=\gcd(P,P')$. Since $D$ divides all the $P_j$; we have $P_{i}=D\cdot Q_{i}$, $P_{i+1}=D\cdot Q_{i+1}$, and $P_{i+2}=D\cdot Q_{i+2}$ and thus:
+Let $D:=\gcd(P,P')$. Since $D$ divides all the $P_j$, we have $P_{i}=D\cdot Q_{i}$, $P_{i+1}=D\cdot Q_{i+1}$, and $P_{i+2}=D\cdot Q_{i+2}$ and thus:
 $$(D\cdot Q_{i}\ \mathrm{div}\ D\cdot Q_{i+1}) \cdot D\cdot Q_{i+1} - D\cdot Q_{i+2} = D\cdot Q_{i}$$
-Cancelling $D$ yields:
+Cancelling $D$ (allowed since $\RR[X]$ is an integral domain and $D\neq 0$) yields:
 $$(Q_{i}\ \mathrm{div}\ Q_{i+1}) \cdot Q_{i+1} - Q_{i+2} = Q_{i}$$
@@
 
-In conclusion: if we want to count the roots of some $P$ between bounds $a,b\in\RR\cup\{-\infty,\infty\}$ with $a\leq b$, we use $\sigma$, $\sigma_\infty$ and $\sigma_{-\infty}$ with a Sturm sequence of $P$, and such a Sturm sequence can be obtained by applying the canonical Sturm sequence construction to $P$ if neither $a$ nor $b$ are roots of both $P$ and $P'$, or by applying it to $P/\gcd(P,P')$ if they are.
+In conclusion: if we want to count the roots of some $P$ between bounds $a,b\in\RR\cup\{-\infty,\infty\}$ with $a\leq b$, we use $\sigma$ with a Sturm sequence of $P$, and such a Sturm sequence can be obtained by applying the canonical Sturm sequence construction to $P$ if neither $a$ nor $b$ are roots of both $P$ and $P'$, or by applying it to $P/\gcd(P,P')$ if they are.
 
@@
 
-The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that Sturm sequences can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b)$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
+The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that\ Sturm se\-quen\-ces can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b]$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
 However, by adding case distinctions on whether $a$ and $b$ are roots of $P$, this can easily be generalised to arbitrary open, half-closed, closed, bounded, and unbounded real intervals.
@@
 Furthermore, statements such as $\forall x \in I.\ P(x) \neq 0$ can be decided by counting the number of roots with the above method and checking whether it is zero.
-This, in turn, can be used to decide statements of the form $\forall x\in I.\ P(x) > 0$ by checking that $\forall x\in I.\ P(x) > 0$ holds and $\lim_{x\to\infty} P(x) = \infty$.
 
@@
 we can further generalise our method to arbitrary statements of the form
-$$(\forall x \in I.\ P(x))\hskip5mm\mathrm{and}\hskip5mm |\{x\ |\ Q(x)\}| = n$$
-where $I$ is an open, half-closed, or closed, bounded or unbounded real interval and $P$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $\hat{\ }$, $\neq$ and $Q$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $\hat{\ }$, $=$.\\
+$$(\forall x \in I.\ P(x))\hskip5mm\mathrm{and}\hskip5mm |\{x\in I\ |\ Q(x)\}| = n$$
+where $I$ is an open, half-closed, or closed, bounded or unbounded real interval and $P$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $^c$, $\neq$ and $Q$ is a property consisting of the operators $\wedge$, $\vee$, $+$, $-$, $\cdot$, $^c$, $=$.\\
 
-Independently from this, the observation that $\forall x\in I.\ P(x)>0$ holds iff $P$ has no roots in $I$ and $P(x)>0$ for some arbitrary $x\in I$ also allows us to decide statements of the form
+Independently from this, the observation that $\forall x\in I.\ P(x)>0$ holds iff $P$ has no roots in $I$ and $\lim_{x\to\infty} P(x) = \infty$ also allows us to decide statements of the form
 $$(\forall x\in I.\ P(x) > 0)\hskip5mm\mathrm{and}\hskip5mm(\forall x\in I.\ P(x) < 0)$$\vskip5mm
@@
 
+For more information on the proof method, refer to the user guide \cite{sturm_userguide}.
+
+\printbibliography
+
 \end{document}

### 2013-12-05 paper/paper.tex
Added content to slides, tuned paper and guide

@@
 \item $Q_0$ and $Q_1$ have no common roots:\\
-By construction, we have $Q_0 = P / \gcd(P,P')$ and $Q_1 = (P / \gcd(P, P'))'$.
-This then implies $\gcd(Q_0, Q_1) = 1$.
-However, the proof for this algebraic fact is nontrivial; it was, of course, proven in the \emph{Isabelle} version of this proof, but it is too complex and at the same time not important enough to be discussed in detail here.
+By construction, we have $Q_0 = P / \gcd(P,P')$ and $Q_1 = P' / \gcd(P, P')$. Obviously, $Q_0$ and $Q_1$ are then coprime and cannot have any common roots.
 \item for any $i\in\{0,\ldots,n-1\}$ and any root $x_0$ of $Q_{i+1}$, $Q_{i}(x_0)Q_{i+2}(x_0) < 0$ holds:\\

### 2013-12-05 slides/slides.tex
Added content to slides, tuned paper and guide

\documentclass[%handout,
	sans,			% use sans-serif font
	%serif,			% use serif-font
	mathsans,		% set mathtext to sans-serif
	%mathserif,		% set mathtext to serif
	%10pt,
	10pt,
	%12pt,
	t		% add text at the top border of slide
	slidescentered,% center text on slide
	%draft,			% compile as draft version
	%notes,			% include nodes in slides
	%compress		% compress navigation bar
]{beamer}

\usetheme{lmtslides}
\setlang{en}
\usepackage{eso-pic}
\usepackage{graphicx}
%\usepackage[pdftex]{color}
\usepackage{times}
\usepackage[utf8x]{inputenc}
\usepackage[amssymb]{SIunits}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{url}
\usepackage{csquotes}

\usepackage{default}
\usepackage{hyperref}

\usepackage{mathptmx}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{}{n}

\renewcommand*\sfdefault{uop}
\usepackage[T1]{fontenc}

\graphicspath{{figures/}}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\sturm}{\texttt{sturm}}
\newcommand{\lem}{\textbf{lemma}}
\newcommand{\card}{\textrm{card}}
\newcommand{\real}{\textrm{real}}

% MODIFY THESE ACCORDINGLY! ---
\title[asdf]{A Formalisation of Sturm's Theorem\\in Isabelle/HOL}
\subtitle{}
\type{IDP Talk}
\author{Manuel Eberl}
\email{eberlm@in.tum.de}
\date{12 December 2013} % TODO
%------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\AddToShipoutPicture{\TitlePicture}
\maketitle
\ClearShipoutPicture
\AddToShipoutPicture{\BackgroundPicture}

\begin{frame}{Motivation}
\begin{description}
\item[We have:] a polynomial with real coefficients
\pause\item[We want:] the number of real roots in a specific interval
\end{description}
\end{frame}

\begin{frame}{Motivation}
\vspace*{1.65mm}
\begin{description}
\item[We have:] a polynomial with $\overbrace{\text{\color{red}real\ coefficients}}^{\text{not\ really}}$
\vspace*{-0.05mm}
\item[We want:] the number of real roots in a specific interval
\end{description}
\vspace{5mm}
\pause
For \enquote{real} computations: restricted to appropriate subset of $\RR$, such as $\QQ$.
\end{frame}

\begin{frame}{Motivation}
The solution: \emph{Sturm's Theorem}\\[2mm]

Provides a method for counting real roots \emph{algorithmically}.\\[10mm]

$\Longrightarrow$ Let's formalise it in Isabelle/HOL
\end{frame}

\begin{frame}{Notation}
\emph{Sign changes:} $\sigma(P_0,\ldots,P_{n-1}; x)$ denotes denotes the number of sign changes in the sequence $P_0(x), \ldots, P_{n-1}(x)$\\[2mm]
For the functionally inclined:\vskip-6mm
$$\sigma(ps; x) = (\mathrm{length} \circ \mathrm{remdups\_adj} \circ \mathrm{filter}\ (\neq 0) \circ \mathrm{map}\ (\lambda p.\ p(x)))\ ps\ -\ 1$$
\end{frame}

\begin{frame}{Sturm's Theorem}
Sturm's Theorem: \textit{Let $P$ be a real polynomial and $P_0, \ldots, P_{n-1}$ a Sturm sequence of $P$. Then
$$\sigma(P_0, \ldots, P_{n-1}; a) - \sigma(P_0, \ldots, P_{n-1}; b)$$
is the number of real roots of $P$ in the interval $(a; b]$.}
\end{frame}

\begin{frame}{Sturm sequence}
But what is a Sturm sequence of $P$?\\
A (reasonably) general definition:\\
\begin{itemize}
\item $P_0 = P$
\item $P_0$ and $P_1$ have no common roots
\item $P_{n-1}$ (last element) does not change its sign
\item if $x_0$ is root of $P_0$: $P_0P_1(x)<0$ in some left-NH and $P_0P_1(x)>0$ in some right-NH of $x_0$
\item if $x_0$ is root of another $P_i$: $P_{i-1}P_{i+1}(x_0) < 0$
\end{itemize}
\end{frame}

\begin{frame}{Assessment}
\begin{description}
\item[The good news:]\ \\formalisation of real analysis, polynomials, algebra already exists
\pause\item[The bad news:]\ \\no formalisation of limits of polynomials, very little on divisibility
\pause\item[The ugly news:]\ \\textbook proofs of \emph{Sturm's theorem} are \emph{extremely} informal proof sketchs at best
\end{description}
\end{frame}

\begin{frame}{Proving Sturm's Theorem}
Assume we already have a Sturm chain. Why does it count roots?\\
Follow $x \mapsto \sigma(P_0, \ldots, P_{n-1}; x)$ passing over $\RR$. Obviously, it can only change at $x_0$ if one of the $P_i$ has a root at $x_0$
\begin{itemize}
\item if $P_i \neq P_0$ has $x_0$ as a root, $P_{i-1}P_{i+1}(x_0) < 0$\\
$\Rightarrow$ signs of $P_{i-1}$, $P_{i+1}$ are $\neq 0$, opposite, constant in NH of $x_0$\\
$\Rightarrow$ signs $[1,\, \_\,, -1]$ or $[-1,\, \_\,, 1]$ in the entire NH, i.e. one sign change\\
$\Rightarrow$ total number of sign changes not influenced\\[1mm]
\item if $P_0$ has $x_0$ as root, $P_0P_1(x_0) < 0$ in left-NH of $x_0$, $>0$ in right-NH\\
$\Rightarrow$ signs are different left of $x_0$ and the same right of $x_0$\\
$\Rightarrow$ total number of sign changes decreases by one
\end{itemize}
\end{frame}

\begin{frame}{Proving Sturm's Theorem}
Formal proof: a lot of induction on the sequences and number of roots\\[2mm]
$\Longrightarrow$ messy and not terribly interesting
\end{frame}

\begin{frame}{Construction Sturm sequences}
Canonical construction for $P$ with no multiple roots (i.e. $\mathrm{gcd}(P,P') = 1$):
$$P_i = \begin{cases}
P & \mathrm{for}\ i = 0\\
P' & \mathrm{for}\ i = 1\\
-(P_{i-2}\ \mathrm{mod}\ P_{i-1}) & \mathrm{otherwise}
\end{cases}$$
\end{frame}

\begin{frame}{Construction Sturm sequences}
Why does it work? Nonobvious parts:\\[3mm]

If $x_0$ is root of $P_0 = P$: $PP'(x_0) < 0$ in left-NH and $PP'(x_0) > 0$ in right-NH\\
-- pick neighbourhood without roots of $P_0$ and $P_1$ (except for $x_0$),\\
\hskip3mm apply mean value theorem\\[3mm]

If $x_0$ is root of another $P_i$: $P_{i-1}P_{i+1}(x_0) < 0$ in some NH of $x_0$\\
-- by construction, $P_{i-1} = Q\cdot P_i - P_{i+1}$ for some $Q\in\RR[X]$\\
\hskip3mm$\Longrightarrow$ $P_{i-1}(x_0) = -P_{i+1}(x_0)$\\
\hskip3mm also: $P_{i-1}(x_0)\neq 0$ since $\mathrm{gcd}(P_{i-1}, P_i) = \mathrm{gcd}(P_0,P_1) = 1$
\end{frame}

\begin{frame}{Construction Sturm sequences}
In case of multiple roots: Let $D := \mathrm{gcd}(P,P')$. Then:
\begin{description}
\item[The obvious way:]\ \\
\begin{itemize}
\item compute canonical Sturm chain of $P / D$\\ (\enquote{divide out} multiple roots)
\end{itemize}
\item[The clever way:]\ \\
\begin{itemize}
\item we can compute the canonical Sturm chain of $P$ and divide by $D$ afterwards
\item but: if $D(x)\neq 0$, dividing by $D$ does not change the number of sign changes at $x$
\item $\Longrightarrow$ unless the interval bounds are multiple roots, we can use the canonical construction without changes
\end{itemize}
\end{description}
\end{frame}

\begin{frame}{Making a decision procedure}
\textit{count\_roots\_between\ p\ a\ b}: picks the most efficient Sturm chain construction and:
$$\textit{count\_roots\_between}\ p\ a\ b\ \ =\ \ |\{x.\ a < x\ \wedge\ x \leq b\ \wedge\ p(x) = 0\}|$$\vskip2mm

Some fluff:
\begin{itemize}
\item case distinctions allow arbitrary combination og $\leq$ and $<$ in bounds
\item \enquote{limit signs} allow infinite bounds
\end{itemize}
In summary: we can count roots in any open/halfopen/closed, bounded/unbounded real interval
\end{frame}

\begin{frame}{Making a decision procedure}
Some more fluff:
\begin{itemize}
\item and/or: count $x$ with $$P(x)=0\ \wedge\ Q(x)=0\hskip5mm\mathrm{or}\hskip5mm P(x)=0\ \vee\ Q(x)=0$$
\item $\forall$-inequalities: $$\forall x.\ P(x) \neq Q(x)\ \wedge\ R(x) \neq S(x)\ \vee\ T(x)\neq U(x)$$
\item $\forall$ with $<$ and $>$: $$\forall x.\ P(x) < Q(x)\ \wedge\ R(x) > S(x)\ \vee\ T(x)\neq U(x)$$
\end{itemize}
\end{frame}

\begin{frame}{\hskip-2mm Making a decision procedure}
\hskip-2mm Examples:
\begin{align*}
&\hskip-2mm\scriptstyle
\lem\ "\card\ \{x::\real.\ (x\, -\, 1)^2\, *\, (x\, +\, 1)\ =\ 0\}\ =\ 2"\ \textrm{\textbf{by}\ sturm}\\[2mm]
&\hskip-2mm\scriptstyle\textbf{lemma}\ "\mathrm{card}\ \{x::\mathrm{real}.\ -0.010831\, <\, x\ \wedge\ x\, <\, 0.010831\ \wedge\\
&\hskip-2mm\scriptstyle\hskip20mm \mathrm{poly}\ [:\,0,\  -17/2097152,\  -49/16777216,\  1/6,\  1/24,\  1/120\,:]\ x\ =\ 0\}\  =\  3"\ \textrm{\textbf{by}\ sturm}\\[2mm]
&\hskip-2mm\scriptstyle\lem\ "\card\ \{x::\real.\ x^3\, +\, x\ =\ 2\,*\,x^2\ \wedge\ x^3\,-\,6*x^2\,+\,11*x\ =\ 6\}\  =\  1"\ \textrm{\textbf{by}\ sturm}\\[2mm]
&\hskip-2mm\scriptstyle\lem\ "(x::\real)^2\,+\,1\ >\ 0"\ \textrm{\textbf{by}\ sturm}
\end{align*}
\end{frame}

\begin{frame}{Size of the formalisation}
3725 LOC in total, 185 of that ML, the rest Isabelle
\begin{center}
\includegraphics[width=8cm]{plot.pdf}
\end{center}
\end{frame}

\end{document}

### 2013-12-17 slides/slides.tex
Updated slides, tuned polynomials

@@
 \begin{itemize}
-\item $P_0 = P$
-\item $P_0$ and $P_1$ have no common roots
-\item $P_{n-1}$ (last element) does not change its sign
-\item if $x_0$ is root of $P_0$: $P_0P_1(x)<0$ in some left-NH and $P_0P_1(x)>0$ in some right-NH of $x_0$
+\item $P_0 = P$\pause
+\item $P_0$ and $P_1$ have no common roots\pause
+\item $P_{n-1}$ (last element) does not change its sign\pause
+\item if $x_0$ is root of $P_0$: $P_0P_1(x)<0$ in some left-NH and $P_0P_1(x)>0$ in some right-NH of $x_0$\pause
 \item if $x_0$ is root of another $P_i$: $P_{i-1}P_{i+1}(x_0) < 0$
@@
 Assume we already have a Sturm chain. Why does it count roots?\\
-Follow $x \mapsto \sigma(P_0, \ldots, P_{n-1}; x)$ passing over $\RR$. Obviously, it can only change at $x_0$ if one of the $P_i$ has a root at $x_0$
+Follow $x \mapsto \sigma(P_0, \ldots, P_{n-1}; x)$ passing over $\RR$. Obviously, it can only change at $x_0$ if one of the $P_i$ has a root at $x_0$\pause
 \begin{itemize}
-\item if $P_i \neq P_0$ has $x_0$ as a root, $P_{i-1}P_{i+1}(x_0) < 0$\\
-$\Rightarrow$ signs of $P_{i-1}$, $P_{i+1}$ are $\neq 0$, opposite, constant in NH of $x_0$\\
-$\Rightarrow$ signs $[1,\, \_\,, -1]$ or $[-1,\, \_\,, 1]$ in the entire NH, i.e. one sign change\\
-$\Rightarrow$ total number of sign changes not influenced\\[1mm]
-\item if $P_0$ has $x_0$ as root, $P_0P_1(x_0) < 0$ in left-NH of $x_0$, $>0$ in right-NH\\
-$\Rightarrow$ signs are different left of $x_0$ and the same right of $x_0$\\
+\item if $P_i \neq P_0$ has $x_0$ as a root, $P_{i-1}P_{i+1}(x_0) < 0$\\\pause
+$\Rightarrow$ signs of $P_{i-1}$, $P_{i+1}$ are $\neq 0$, opposite, constant in NH of $x_0$\\\pause
+$\Rightarrow$ signs $[1,\, \_\,, -1]$ or $[-1,\, \_\,, 1]$ in the entire NH, i.e. one sign change\\\pause
+$\Rightarrow$ total number of sign changes not influenced\\[1mm]\pause
+\item if $P_0$ has $x_0$ as root, $P_0P_1(x_0) < 0$ in left-NH of $x_0$, $>0$ in right-NH\\\pause
+$\Rightarrow$ signs are different left of $x_0$ and the same right of $x_0$\\\pause
 $\Rightarrow$ total number of sign changes decreases by one
@@
 Formal proof: a lot of induction on the sequences and number of roots\\[2mm]
-$\Longrightarrow$ messy and not terribly interesting
+$\Longrightarrow$ messy and not terribly interesting, I'll spare you the details
+\end{frame}
+
+\begin{frame}{Proving Sturm's Theorem}
+We now know that Sturm sequences can count roots.\\
+But how do we construct one?
 \end{frame}
@@
 \begin{frame}{Construction Sturm sequences}
-Why does it work? Nonobvious parts:\\[3mm]
+Why does it work? Nonobvious parts:\\[3mm]\pause
 
@@
 -- pick neighbourhood without roots of $P_0$ and $P_1$ (except for $x_0$),\\
-\hskip3mm apply mean value theorem\\[3mm]
+\hskip3mm apply mean value theorem\\[3mm]\pause
 
@@
 \begin{frame}{Construction Sturm sequences}
-In case of multiple roots: Let $D := \mathrm{gcd}(P,P')$. Then:
+This construction assumed no multiple roots.\\
+What do we do if there are multiple roots?
+\end{frame}
+
+\begin{frame}{Construction Sturm sequences}
+In case of multiple roots: Let $D := \mathrm{gcd}(P,P')$. Then:\pause
 \begin{description}
@@
 \item compute canonical Sturm chain of $P / D$\\ (\enquote{divide out} multiple roots)
-\end{itemize}
+\end{itemize}\pause
 \item[The clever way:]\ \\
 \begin{itemize}
-\item we can compute the canonical Sturm chain of $P$ and divide by $D$ afterwards
-\item but: if $D(x)\neq 0$, dividing by $D$ does not change the number of sign changes at $x$
+\item we can compute the canonical Sturm chain of $P$ and divide by $D$ afterwards\pause
+\item but: if $D(x)\neq 0$, dividing by $D$ does not change the number of sign changes at $x$\pause
 \item $\Longrightarrow$ unless the interval bounds are multiple roots, we can use the canonical construction without changes
@@
 $$\textit{count\_roots\_between}\ p\ a\ b\ \ =\ \ |\{x.\ a < x\ \wedge\ x \leq b\ \wedge\ p(x) = 0\}|$$\vskip2mm
+\pause
 
@@
 \begin{itemize}
-\item and/or: count $x$ with $$P(x)=0\ \wedge\ Q(x)=0\hskip5mm\mathrm{or}\hskip5mm P(x)=0\ \vee\ Q(x)=0$$
-\item $\forall$-inequalities: $$\forall x.\ P(x) \neq Q(x)\ \wedge\ R(x) \neq S(x)\ \vee\ T(x)\neq U(x)$$
+\item and/or: count $x$ with $$P(x)=0\ \wedge\ Q(x)=0\hskip5mm\mathrm{or}\hskip5mm P(x)=0\ \vee\ Q(x)=0$$\pause
+\item $\forall$-inequalities: $$\forall x.\ P(x) \neq Q(x)\ \wedge\ R(x) \neq S(x)\ \vee\ T(x)\neq U(x)$$\pause
 \item $\forall$ with $<$ and $>$: $$\forall x.\ P(x) < Q(x)\ \wedge\ R(x) > S(x)\ \vee\ T(x)\neq U(x)$$

### 2013-12-18 slides/slides.tex
Tuned slides

@@
 \begin{description}
-\item[We have:] a polynomial with real coefficients
+\item[We have:] a polynomial with {\text{\only<1,2>{real coefficients}\only<3->{\color{red}real coefficients}}}
 \pause\item[We want:] the number of real roots in a specific interval
-\end{description}
-\end{frame}
-
-\begin{frame}{Motivation}
-\vspace*{1.65mm}
-\begin{description}
-\item[We have:] a polynomial with $\overbrace{\text{\color{red}real\ coefficients}}^{\text{not\ really}}$
-\vspace*{-0.05mm}
-\item[We want:] the number of real roots in a specific interval
-\end{description}
-\vspace{5mm}
+\end{description}\vskip5mm
 \pause
-For \enquote{real} computations: restricted to appropriate subset of $\RR$, such as $\QQ$.
+For reals?\\\pause
+No. Only for rationals. (or any other sufficiently nice subset of $\RR$)
 \end{frame}
@@
 \begin{frame}{Notation}
-\emph{Sign changes:} $\sigma(P_0,\ldots,P_{n-1}; x)$ denotes denotes the number of sign changes in the sequence $P_0(x), \ldots, P_{n-1}(x)$\\[2mm]
+\emph{Sign changes:} $\sigma(P_0,\ldots,P_{n-1}; x)$ denotes the number of sign changes in the sequence $P_0(x), \ldots, P_{n-1}(x)$\\[2mm]
 For the functionally inclined:\vskip-6mm

### 2014-01-08 paper/paper.tex
Tuned paper

@@
 \begin{abstract}
-\emph{Sturm sequences} are a method of efficiently computing the number of real roots of a real polynomial inside a given interval.
+\emph{Sturm sequences} are a method for computing the number of real roots of a real polynomial inside a given interval efficiently.
 In this project, this fact and a number of me\-thods to construct Sturm sequences efficiently have been formalised with the interactive theorem prover Isabelle\slash HOL.
@@
 
-Of course, one way to handle a polynomial with multiple roots is to use lemma \ref{thm:squarefree} and \enquote{divide away} the \enquote{excess roots} by dividing the polynomial $P$ by $\gcd(P,P')$.
+Of course, one way to handle a polynomial with multiple roots is to use Lemma \ref{thm:squarefree} and \enquote{divide away} the \enquote{excess roots} by dividing the polynomial $P$ by $\gcd(P,P')$.
 However, surprisingly, it turns out that as long as the interval bounds $a$ and $b$ are not multiple roots of $P$ themselves, i.e.
@@
 First, we note that $\gcd(P,P')\neq 0$, since $P'\neq 0$ by assumption.
-Furthermore, lemma \ref{thm:sturm_gcd} implies that $\gcd(P,P')\ |\ P_i$ for all $i$, thus $Q_i\in\RR[X]$, i.e.
+Furthermore, Lemma \ref{thm:sturm_gcd} implies that $\gcd(P,P')\ |\ P_i$ for all $i$, thus $Q_i\in\RR[X]$, i.e.
 the $Q_i$ really are polynomials.
@@
 \vskip3mm
-It remains to show that $Q_0$ has the same roots, disregarding multiplicity, as $P = P_0$. This is precisely the statement of lemma \ref{thm:squarefree}.
+It remains to show that $Q_0$ has the same roots, disregarding multiplicity, as $P = P_0$. This is precisely the statement of Lemma \ref{thm:squarefree}.
 \end{proof}\vskip10mm
@@
 
-The statements that were proven in lemma \ref{thm:count_roots_between} and lemma \ref{thm:count_roots} were that\ Sturm se\-quen\-ces can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b]$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
+The statements that were proven in Lemma \ref{thm:count_roots_between} and Lemma \ref{thm:count_roots} were that\ Sturm se\-quen\-ces can be used to count the roots of a polynomial $P$ in an interval of the form $(a;b]$ for $a,b\in\RR$, $a < b$ or in the interval $(-\infty;\infty)$.
 However, by adding case distinctions on whether $a$ and $b$ are roots of $P$, this can easily be generalised to arbitrary open, half-closed, closed, bounded, and unbounded real intervals.

### 2014-01-11 document/root.tex
Finalised AFP submission; tuned

\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
\usepackage{amsmath}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}

\begin{document}

\title{A Formalisation of Sturm's Theorem}
\author{Manuel Eberl}
\maketitle

\begin{abstract}
\emph{Sturm sequences} are a method for computing the number of real roots of a real polynomial inside a given interval efficiently. In this project, this fact and a number of me\-thods to construct Sturm sequences efficiently have been formalised with the interactive theorem prover Isabelle\slash HOL. Buil\-ding upon this, an Isabelle\slash HOL proof method was then implemented to prove statements about the number of roots of a real polynomial and related properties.
\end{abstract}
\vskip10mm

\tableofcontents

\newpage

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
%\bibliographystyle{abbrv}
%\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-08-19 guide/guide.tex
Stuff

@@
 It also provides some preprocessing to decide a number of statements that can be reduced to real roots of polynomials, such as simple polynomial inequalities and logical combinations of polynomial equations.
-\vspace*{10mm}
 
@@
 \begin{align*}
-&\lemma\ "\card\ \{x::\real.\ (x - 1)^2 * (x + 1) = 0\} = 2"\ \textrm{\textbf{by}\ sturm}\\
-&\textbf{lemma} "\mathrm{card}\ \{x::\mathrm{real}.\ -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
-&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\} = 3"\ \textrm{\textbf{by}\ sturm}\\
-&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\} = 1"\ \textrm{\textbf{by}\ sturm}\\
-&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\} = 4"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\card\ \{x::\real.\ (x - 1)^2 * (x + 1) = 0\}\  =\  2"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\mathrm{card}\ \{x::\mathrm{real}.\ -0.010831 < x\ \wedge\ x < 0.010831\ \wedge\\
+&\hskip20mm \mathrm{poly}\ [:0, -17/2097152, -49/16777216, 1/6, 1/24, 1/120:]\ \ x\ =\ 0\}\  =\  3"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \wedge\ x^3-6*x^2+11*x=6\}\  =\  1"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\card\ \{x::\real.\ x^3 + x = 2*x^2\ \vee\ x^3-6*x^2+11*x=6\}\  =\  4"\ \textrm{\textbf{by}\ sturm}\\
 &\lemma\ "(x::\real)^2+1 > 0"\ \textrm{\textbf{by}\ sturm}\\
@@
 &\lemma\ "(x::\real) > 1\ \Longrightarrow\ x*x > x"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "x \geq -0.5\ \Longrightarrow\ x \leq 1\ \Longrightarrow\ x^3 - 5*x^2 + x + 3 \geq (0::\mathrm{real})"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\textrm{mono}\ (\lambda x::\mathrm{real}.\ x^3)"\ \textrm{\textbf{by}\ sturm}\\
+&\lemma\ "\textrm{strict\_mono}\ (\lambda x::real.\\\
+&\hskip20mm\ x^7/7 - 14/5*x^5 + 3*x^4 + 49/3*x^3 - 42*x^2 + 36*x - 13)"\ \textrm{\textbf{by}\ sturm}\\
 \end{align*}
@@
 
-\newpage
 \subsection{Inequalities}
@@
 By using some simple preprocessing, the \sturm\ method can also decide more complex statements:
-$$\card\ \{x::\real.\ x > a\ \wedge\ x < b\ \wedge\ P\ x\}$$
+$$\card\ \{x::\real.\ x > a\ \wedge\ x < b\ \wedge\ P\ x\}\ =\ n$$
 where $P\ x$ is a \enquote{polynomial expression}, which is defined as:
@@
 
-\subsection{Simple ordered inequalities}
+\subsection{Simple strict linear inequalities}
 For any polynomial $p\in\RR[X]$, the question whether $p(x) > 0$ for all $x\in I$ for a non-empty real interval $I$ can obviously be reduced to the question of whether $p(x) \neq 0$ for all $x\in I$, \ie $p$ has no roots in $I$, and $p(x) > 0$ for some arbitrary fixed $x\in I$, the first of which can be decided using Sturm's theorem and the second by choosing an arbitrary $x\in I$ and evaluating $p(x)$.\\
@@
 $$\forall x::\real.\ x > a\ \wedge\ x < b\ \longrightarrow\ p\ x < q\ x$$
+
+\subsection{Simple nonstrict linear inequalities and monotonicity}
+
+The \sturm\ method can also decide nonstrict inequalities, e.g.
+$p(x) \geq 0$ for all $x\in I$ where $I$ is a non-empty real interval.
+In contrast to the previous cases, proving nonstrict inequalities is done with help from an ML function that computes a witness, which is then used in the proof by the actual proof method.
+The witness is a decomposition of the interval in question into disjoint intervals such that every interval contains exactly one root, which makes it possible to check nonnegativity by verifying it on each interval.
+
+Note that while this does not compromise soundness in any way (as witnesses still have to go through checking in Isabelle), it may compromise completeness if the ML code, due to a bug, fails to provide an adequate witness.\newpage
+
+Additionally, the \sturm\ method can also prove monotonicity and strict monotonicity of polynomials (over the entire real line) by reducing the problem to nonnegativity of the derivative.
 
@@
 \item use a computer algebra system to ensure that the property is indeed correct
-\item if this did not help, send the statement in question to \texttt{eberlm@in.tum.de}; it may be a bug in the preprocessing of the proof method.
+\item if this did not help, send the statement in question to \texttt{eberlm@in.tum.de}; it may be a bug in the preprocessing of the proof method or
 \end{enumerate}
 
+Note: in case of monotonicity and nonnegativity, the proof method will output a counterexample if the statement is false.
+Should proving nonnegativity or monotonicity fail without a counterexample being given, please also report this as a bug.
+
 \end{document}
